### 2020-06-01 dsaa_application_manuscript/dsaa_application_manuscript.Rmd
submission to dsaa

---
title: Urban and socio-economic correlates of property price evolution\:\ Application to Dublin's costal area
affiliation:
  institution-columnar: true

  institution:
    - name: Dublin City University
      department: Business School
      location: Glasnevin, Dublin 09
      email: damien.dupre@dcu.ie
      mark: 1
      author:
        - name: Damien Dupré

keywords: ["Property price", "Housing Market", "Feature Analysis", "Machine Learning", "Geocoding"]
abstract: |
  Understanding the characteristics of the housing market is essential for both sellers and buyers. However, the housing market is influenced by multiple factors. In this paper, the urban and socio-economic structure of an area is used to predict the price of 10389 houses sold in 2018 in the city of Dublin. More precisely, the direct distance from each house to 160 urban features taken from OpenStreetMap is calculated, and an extreme gradient boosting linear regression performed. Using these features, the model explains 47\% of the housing price variance. The most important features in this model are the proximity to an embassy and to a grassland. In addition, the results of a population census from 2016 are also used to correlate with the price of houses. From this census, 48 features are used as the input of a gradient boost linear regression model. In all, the socio-economic features are explaining 42\% of the housing price variance as well. The most important socio-economic features are; the density of houses having more than eight or more rooms in the area, the density of young children, and the density of individuals reporting that they have no religion. By taking into account either urban or socio-economic features, it is possible to accurately estimate housing prices and to predict their evolution.

bibliography: [references.bib, r-references.bib]
output:
  bookdown::pdf_book:
    base_format: rticles::ieee_article

# output: word_document

header-includes:
  - \usepackage{graphicx}
---

(ref:distrib-plot) Overall (A) and geographic (B) distribution of the houses prices density in the city of Dublin in 2018. The geographic distribution was obtained by using a generalized additive model with soap film smooth parameter to take into account the influence of coastal boundaries. Geographic estimates outside of the 95\% Confidence Interval were removed.
(ref:osm-features-xgb) Property price prediction accuracy (A) and Property price prediction error (B) using urban features with XGBoost.
(ref:osm-features-table) Urban features importance (higher than 1\%).
(ref:census-features-xgb) Property price prediction accuracy (A) and Property price prediction error (B) using socio-economic features with XGBoost.
(ref:census-features-table) Socio-economic features importance (higher than 1\%).

```{r setup, include = FALSE}
# options ----------------------------------------------------------------------
set.seed(123)
options(scipen=999)
knitr::opts_chunk$set(
  cache.extra = knitr::rand_seed, 
  message = FALSE, 
  warning = FALSE, 
  error = FALSE, 
  echo = FALSE
  )

# libraries --------------------------------------------------------------------
library(tidyverse)
library(magrittr)
library(here)
library(lubridate)
library(papaja)
library(patchwork)
library(kableExtra)
library(latex2exp)
library(yardstick)
library(sf)
library(tmap)
library(lwgeom)
library(osmdata)
library(OpenStreetMap)
library(mgcv)
library(xgboost)

# R package references ---------------------------------------------------------
#papaja::r_refs(file = "dsaa_application_manuscript/r-references.bib")

# functions --------------------------------------------------------------------
source(here::here("script/functions_project.R"))
source(here::here("script/osm_dublin_features.R"))
# data path --------------------------------------------------------------------
data_path <- here::here() %>% 
  dirname() %>% 
  dirname() %>% 
  paste0("/data/property_price_ireland_data")

# all price registry------------------------------------------------------------
data_dublin_geocoded <- file.path(data_path, "data_dublin_geocoded_clean.rds") %>% 
  readr::read_rds() %>%
  janitor::clean_names() %>%
  dplyr::mutate(
    year_fact = as.factor(as.character(year))
  )

mean_dat <- mean(data_dublin_geocoded$price)
sd_dat <- sd(data_dublin_geocoded$price)

data_dublin_geocoded <- data_dublin_geocoded %>%
  dplyr::filter(price > mean_dat - sd_dat)

# all dublin features ----------------------------------------------------------
data_dublin_2018_features <- data_path %>%
  file.path("data_dublin_2018_osm_features.rds") %>%
  readr::read_rds() %>% 
  janitor::remove_empty(which = "cols") %>% 
  janitor::clean_names() %>%
  dplyr::mutate(geog_id = stringr::str_remove(geogid, "[A]"))

list_geogid <- unique(data_dublin_2018_features$geog_id)

data_dublin_2018_features <- c(
  "carers.csv",
  "disabilty_age_group.csv", 
  "general_health.csv",
  "housing_occupancy.csv",
  "housing_rooms.csv",
  "housing_tenure.csv",
  "housing_type.csv",
  "population.csv",
  "religion.csv"
  ) %>%
  purrr::map(read_join) %>% 
  purrr::reduce(dplyr::left_join) %>% 
  dplyr::inner_join(data_dublin_2018_features, by = c("GEOG_ID" = "geog_id"))

# train test split -------------------------------------------------------------
train_index <- sample(1:nrow(data_dublin_2018_features), 0.8 * nrow(data_dublin_2018_features))
test_index <- setdiff(1:nrow(data_dublin_2018_features), train_index)

train_dublin_2018_features <- data_dublin_2018_features[train_index, ]
test_dublin_2018_features <- data_dublin_2018_features[test_index, ]

# dublin map -------------------------------------------------------------------
# boundaries_dublin <-
#   osmdata::opq(bbox = 'Dublin, Ireland') %>%
#   osmdata::add_osm_feature(key = 'admin_level', value = '7') %>%
#   osmdata::osmdata_sf() %>%
#   osmdata::unique_osmdata()
# 
# bb <- st_bbox(boundaries_dublin$osm_multipolygons$geometry)
# # Defining a buffer
# buffer <- 0.1
# p_big <- rbind(
#   c(bb[1] - buffer, bb[2] - buffer),
#   c(bb[1] - buffer, bb[4] + buffer),
#   c(bb[3] + buffer, bb[4] + buffer),
#   c(bb[3] + buffer, bb[2] - buffer),
#   c(bb[1] - buffer, bb[2] - buffer)
#   )
# # Putting the coordinates into a squared polygon object
# pol_dublin <- st_polygon(list(p_big)) %>% st_geometry
# # Providing the SRID (here, unprojected lon/lat)
# st_crs(pol_dublin) <- 4326
# 
# # get Dublin coast line ------------------------------------------------------
# coast_dublin <- opq(bbox = st_bbox(pol_dublin)) %>%
#   add_osm_feature(key = 'natural', value = 'coastline') %>%
#   osmdata_sf %>% unique_osmdata
# 
# blade_dublin <- coast_dublin$osm_lines$geometry %>% st_union %>% st_line_merge
# 
# ls <- st_linestring(blade_dublin[[1]][[1]])
# 
# dublin_multipol <- st_split(st_geometry(pol_dublin), st_geometry(ls))
# 
# readr::write_rds(dublin_multipol, here::here("data/dublin_multipol.rds"))
dublin_multipol <- readr::read_rds(here::here("data/dublin_multipol.rds"))

dublin_sea <- dublin_multipol[[1]][[2]] %>% 
  st_polygon %>% 
  st_geometry
st_crs(dublin_sea) <- 4326
```

# Introduction

First time property acquisition is an important achievement in individuals' lifetime. It provides not only housing security but also the feeling of being a landowner. However, access to the status of landowner is a challenge for low to middle income earners as it corresponds to significant spending which will potentially have budget impacts for years [@savage2014property]. For this reason, understanding property price markets as well as factors driving price is essential to identify investment opportunities.

Beside their inherent characteristics such as size, design and materials [@bourassa2007spatial; @liu2013spatial], property valuation is dictated by multiple external factors explaining their spatial autocorrelation [@basu1998analysis]. Some of these factors are macroeconomic [@antonakakis2016dynamic], related to the evolution of population general wealth while others are related to the characteristics of the neighbourhood [@dubin1998predicting]. These characteristics include the urban and socio economic structure of the neighbourhood [@goodman1998housing]. However, it is difficult to evaluate the influence of such external factors on property valuation [@clapp2002predicting]. By using machine learning algorithms, prediction of house prices has become more and more accurate. Techniques used usually include artificial neural networks due to the volatility of the market [@limsombunchai2004house; @feng2015comparing; @selim2009determinants].

In this paper, an alternative machine learning algorithm is used to identify urban and socio-economic correlates of property price evolution in the area of Dublin, Ireland in 2018. Due to the high number of potential features, an eXtreme Gradient Boosting (XGBoost) regression model is believed to obtain the best results with low parameter customisation and high speed.

# Method

Since 2010, under the Irish Property Services (Regulation) Act, all individuals acquiring a property in Ireland have to declare it to Property Services Regulatory Authority (PSRA). They must provide details such as the date of sale, the price and the address of all residential properties purchased in Ireland as declared to the Revenue Commissioners for stamp duty purposes [@rppr2020]. Data is filed electronically by persons doing the conveyancing of the property on behalf of the purchaser, and it must be noted that errors may occur when the data is being filed.

Because of the evolution of urban planning and the evolution of socio-economic features measured by the 2016 census, only houses sold in 2018 were processed to avoid potential temporal incoherences. In 2018, a total of 10395 property in the area of Dublin were sold, and these constitute the database used to identify the urban and socio-economic correlates of property prices.

In order to evaluate the spacial distribution of the property sold, property addresses were geocoded (i.e., converted to latitude and longitude) using the OpenStreetMap API. OpenStreetMap is a collaborative project which aims to create and provide access to free editable maps of the world [@haklay2008openstreetmap].

To estimate the geographical distribution of the price density, multiple methods such as multiple regression [@mccluskey2000application] or Bayesian smoothing [@clapp2002predicting] have been employed. While these methods are efficient in a non-restricted space, they have limitations when they are used in coastal areas. To deal with the influence of boundaries on price estimation, a generalised additive model using soap film smoothing has been used [@wood2008soap].

## Distance to urban features

OpenStreetMap combines information about more than 400 urban features including road information and building information to categorize features such as amenities, leisure or tourism structures [@openstreetmap2020]. Among these features, only 160 are available or relevant to the Dublin Area. The distance between each property and the closest point corresponds to each of the 160 urban features.

## Density of socio-economic features

In addition to the distance to every urban feature, each house sold in 2018 was related to 48 socio-economic features of the small area including the house, as measured by the Irish census 2016 [@cso2020]. The results of Irish 2016 census consultation is accessible through both the Central Statistics Office and All-Island Research Observatory. The data obtained can be mapped over small area boundaries which are fractions of Irish Electoral Division map. The social features extracted are corresponding to population information, religion, carers and health. Economic features correspond to the characteristics of each small area; including the proportion of housing types, number of rooms, occupancy and tenure per small area. Each property is then associated to the value corresponding to its small area. For anonymisation purposes, the results from small areas having less than five respondents to the census were converted to a proportion of five respondents.

## Gradient Boosting Regressions

Lastly, an eXtreme Gradient Boosting (XGBoost) regression model was calculated to examine the extent to which house prices rely on urban and on socio-economic features. The XGBoost algorithm optimizes the prediction accuracy by performing iterative least-squares regressions (set to 100 iterations as the best trade-off between accuracy and speed), thereby minimizing the root mean squared error [@friedman2001greedy; @chen2016xgboost]. The original dataset has been randomly split to 80% for training and 20% for testing the models. Models accuracy is estimated using the coefficient of determination ($R^2$), the Root Mean Square Error ($RMSE$) and the Mean Absolute Error ($MAE$) of the predicted property price values of the test dataset.

# Results

```{r gam-model}
# bound ------------------------------------------------------------------------
gam_boundaries <- dublin_multipol[[1]][[1]][[1]]

bound <- list(
  list(
    x = gam_boundaries[,1], 
    y = gam_boundaries[,2], 
    f = rep(0, nrow(gam_boundaries))
  )
)

# knots ------------------------------------------------------------------------
# boundaries_dublin <- 
#   osmdata::opq(bbox = 'Dublin, Ireland') %>%
#   osmdata::add_osm_feature(key = 'admin_level', value = '7') %>%
#   osmdata::osmdata_sf() %>% 
#   osmdata::unique_osmdata()
# 
# bb <- st_bbox(boundaries_dublin$osm_multipolygons$geometry)
# 
# readr::write_rds(bb, here::here("data/bb.rds"))

bb <- readr::read_rds(here::here("data/bb.rds"))

bound.grid.lng <- seq(from = bb[1], to = bb[3], len = 10)
bound.grid.lat<- seq(from = bb[2], to = bb[4], len = 10)

bound.grid.map <- expand.grid(
  lng = bound.grid.lng, 
  lat = bound.grid.lat
  )

names(bound.grid.map) <- c("x","y")

knots <- bound.grid.map[with(bound.grid.map, inSide(bound, x, y)), ]

names(knots) <- c("lng", "lat")
names(bound[[1]]) <- c("lng", "lat", "f")

dublin_sea_frame <- dublin_multipol[[1]][[2]]

knots$pb <- apply(knots, 1, function(row) {  
  sf::st_is_within_distance(
    dublin_sea_frame,
    sf::st_point(c(row["lng"],row["lat"])),
    dist = 0.001
  ) %>% 
    as.numeric()
})

knots_pb <- knots %>%
  dplyr::filter(pb == 1)

knots <- dplyr::anti_join(knots, knots_pb, by = c("lng", "lat"))

# model ------------------------------------------------------------------------
gam_dublin_gps <- mgcv::gam(
  price ~ s(lng, lat, bs = "so", xt = list(bnd = bound)), 
  data = dplyr::filter(data_dublin_geocoded, year == 2018), 
  method = "REML", 
  knots = knots,
  family = gaussian(link = "identity")
)

gam_pred <- gam_density(gam_dublin_gps, too.far = 0.05, n.grid = 30)
```

In 2018, the average price of a sold property in Dublin was `r scales::dollar(round(mean_dat, 0), suffix = "", prefix = "€")` (SD = `r scales::dollar(round(sd_dat, 0), suffix = "", prefix = "€")`). In order to remove potential human errors and outliers, prices higher or lower than 1 SD were removed from the original dataset.

```{r distrib-plot, fig.env='figure*', fig.height=3, out.width='0.98\\textwidth', fig.cap="(ref:distrib-plot)"}
price_density <- data_dublin_geocoded %>%
  dplyr::filter(year == 2018) %>%
  magrittr::use_series(price) %>%
  stats::density() %$%
  data.frame(x, y) %>%
  ggplot() +
  geom_segment(aes(
    x = x,
    xend = x,
    y = 0,
    yend = y,
    color = x
  )) +
  scale_x_continuous("Price", labels = scales::dollar_format(suffix = "", prefix = "€")) +
  scale_y_continuous("Density") +
  scale_color_gradient(
    low = '#3a53a4', 
    high = '#ed1c24', 
    ) +
  theme_apa() +
  theme(
    text = element_text(family = "serif", size = 6),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  ) +
  ggtitle("A")

map_density <- gam_pred %>%
  dplyr::filter(value > 200000) %>%
  ggplot() +
  geom_raster(aes(x = lng, y = lat, fill = value)) +
  geom_sf(data = dublin_sea, fill = "light blue") +
  scale_fill_gradient(
    "Price Density", 
    low = '#3a53a4', 
    high = '#ed1c24', 
    trans = "log",
    breaks = c(250000, 1000000), 
    labels = scales::dollar_format(suffix = "", prefix = "€")) +
  scale_x_continuous(
    "Longitude", 
    limits = c(-6.5, -6), 
    breaks = seq(-6.5, -6, 0.2)) +
  scale_y_continuous(
    "Latitude", 
    limits = c(53.2, 53.5), 
    breaks = seq(53.2, 53.5, 0.1)) +
  theme_apa() +
  theme(
    text = element_text(family = "serif", size = 6),
    legend.position = "none"
  ) +
  ggtitle("B")

price_density + map_density
```

The density of housing prices reveals a unimodal distribution of property prices (Figure \@ref(fig:distrib-plot)A). In addition, the geographical distribution of property prices indicates a higher valuation in the South-West of the city as well as on the coast line (Figure \@ref(fig:distrib-plot)B).

## Correlates with urban features

```{r osm-features}
train_tbl_y <- train_dublin_2018_features %>%
  dplyr::select(price) %>%
  as.matrix()
train_tbl_x <- train_dublin_2018_features %>%
  dplyr::select(any_of(osm_dublin_features)) %>% 
  as.matrix()

test_tbl_y <- test_dublin_2018_features %>%
  dplyr::select(price) %>%
  as.matrix()
test_tbl_x <- test_dublin_2018_features %>%
  dplyr::select(any_of(osm_dublin_features)) %>% 
  as.matrix()

xgmodel <- xgboost::xgboost(
  data = train_tbl_x,
  label = train_tbl_y,
  objective = "reg:linear",
  eval_metric = "rmse",
  verbose = 0,
  nround = 100
)

pred <- predict(xgmodel, test_tbl_x)

result <- data.frame(pred = pred, raw = test_tbl_y)

metrics_result <- result %>% 
  yardstick::metrics(price, pred) %>% 
  select(-.estimator) %>% 
  mutate(.estimate = round(.estimate, 2))

lm_result <- lm(pred ~ price, result) %>%
  papaja::apa_print.lm()
```

Using the XGBoost algorithm, the 160 urban features are explaining `r scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"])` of the property price variance (`r lm_result$statistic$modelfit$r2`) with a $RMSE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "rmse"], suffix = "", prefix = "€")` and a $MAE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "mae"], suffix = "", prefix = "€")`.

```{r osm-features-xgb, fig.cap="(ref:osm-features-xgb)", fig.height=8, out.width='0.98\\columnwidth'}
result_accuracy <- result %>%
  ggplot(aes(x = price, y = pred)) +
  geom_point(alpha = 0.1) +
  geom_smooth(
    method = "lm",
    fullrange = TRUE,
    se = TRUE,
    color = "black",
    size = 0.1
  ) +
  scale_x_continuous(
    "Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  scale_y_continuous(
    "Predicted Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  theme_apa() +
  theme(
    text = element_text(family = "serif"),
    legend.position = "none"
  ) +
  labs(
    title = "A",
    subtitle = latex2exp::TeX(paste(
      "$R^2 =$", 
      scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"]),
      ";",
      "$RMSE =$",
      scales::dollar(
        metrics_result$.estimate[metrics_result$.metric == "rmse"], 
        suffix = "", 
        prefix = "€"
      ),
      ";",
      "$MAE =$",
      scales::dollar(
        metrics_result$.estimate[metrics_result$.metric == "mae"], 
        suffix = "", 
        prefix = "€"
      )
      )
    )
  )

result_error <- result %>%
  dplyr::mutate(diff_percent = (pred - price) / price) %>%
  ggplot(aes(x = price, y = diff_percent)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_smooth(
    color = "black",
    se = TRUE,
    fullrange = TRUE,
    size = 0.1
  ) +
  scale_x_continuous(
    "Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  scale_y_continuous(
    "Predicted Property Price Error",
    limits = c(-0.5, 0.5),
    labels = scales::percent
  ) +
  theme_apa() +
  theme(
    text = element_text(family = "serif")
  ) +
  ggtitle("B")

result_accuracy / result_error
```

The overall correlation of the predicted property prices with urban features is shown in Figure \@ref(fig:osm-features-xgb)A. It can be noticed that property prices situated at the low and the high end of the range are the most difficult to predict (Figure \@ref(fig:osm-features-xgb)B). A possible explanation is the absence of distinctive and recurrent patterns in urban features for these houses. However, the prices higher than €300,000 are potentially driving down the prediction accuracy due to outliers.

```{r osm-features-table, results="asis"}
table_osm <- xgb.importance(model = xgmodel) %>%
  dplyr::filter(Gain > 0.01) %>%
  dplyr::arrange(desc(Gain)) %>%
  dplyr::mutate(across(where(is.numeric), scales::percent, accuracy = 0.1)) %>%
  dplyr::select(Feature, Importance = Gain) %>%
  tidyr::separate(Feature, c("Feature Category", "Feature Type"), sep = "_")

table_osm %>% 
  knitr::kable(
    "latex",
    caption = "(ref:osm-features-table)",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kableExtra::kable_styling(font_size = 8, full_width = TRUE)
```

By analysing their importance (Table \@ref(tab:osm-features-table)), the most relevant geographical features to predict housing prices are the distance to an embassy (`r table_osm$Importance[1]`) and the distance to natural grasslands such as parks and gardens (`r table_osm$Importance[2]`).

## Correlates with socio-economic features

```{r census-features}
# census dublin features
source(here::here("script/census_dublin_features.R"))

train_tbl_y <- train_dublin_2018_features %>%
  dplyr::select(price) %>%
  as.matrix()
train_tbl_x <- train_dublin_2018_features %>%
  dplyr::select(any_of(census_dublin_features)) %>% 
  as.matrix()

test_tbl_y <- test_dublin_2018_features %>%
  dplyr::select(price) %>%
  as.matrix()
test_tbl_x <- test_dublin_2018_features %>%
  dplyr::select(any_of(census_dublin_features)) %>% 
  as.matrix()

xgmodel <- xgboost::xgboost(
  data = train_tbl_x,
  label = train_tbl_y,
  objective = "reg:linear",
  eval_metric = "rmse",
  verbose = 0,
  nround = 100
)

pred <- predict(xgmodel, test_tbl_x)

result <- data.frame(pred = pred, raw = test_tbl_y)

metrics_result <- result %>% 
  yardstick::metrics(price, pred) %>% 
  select(-.estimator) %>% 
  mutate(.estimate = round(.estimate, 2))

lm_result <- lm(pred ~ price, result) %>%
  papaja::apa_print.lm()
```

Using the XGBoost algorithm, the 48 socio-economic features explain `r scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"])` of the property price variance (`r lm_result$statistic$modelfit$r2`) with a $RMSE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "rmse"], suffix = "", prefix = "€")` and a $MAE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "mae"], suffix = "", prefix = "€")`.

```{r census-features-xgb, fig.cap="(ref:census-features-xgb)", fig.height=8, out.width='0.98\\columnwidth'}
result_accuracy <- result %>%
  ggplot(aes(x = price, y = pred)) +
  geom_point(alpha = 0.1) +
  geom_smooth(
    method = "lm",
    fullrange = TRUE,
    se = TRUE,
    color = "black",
    size = 0.1
  ) +
  scale_x_continuous(
    "Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  scale_y_continuous(
    "Predicted Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  theme_apa() +
  theme(
    text = element_text(family = "serif"),
    legend.position = "none"
  ) +
    labs(
    title = "A",
    subtitle = latex2exp::TeX(paste(
      "$R^2 =$", 
      scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"]),
      ";",
      "$RMSE =$",
      scales::dollar(
        metrics_result$.estimate[metrics_result$.metric == "rmse"], 
        suffix = "", 
        prefix = "€"
      ),
      ";",
      "$MAE =$",
      scales::dollar(
        metrics_result$.estimate[metrics_result$.metric == "mae"], 
        suffix = "", 
        prefix = "€"
      )
      )
    )
  )

result_error <- result %>%
  dplyr::mutate(diff_percent = (pred - price) / price) %>%
  ggplot(aes(x = price, y = diff_percent)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_smooth(
    color = "black",
    se = TRUE,
    fullrange = TRUE,
    size = 0.1
  ) +
  scale_x_continuous(
    "Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  scale_y_continuous(
    "Predicted Property Price Error",
    limits = c(-0.5, 0.5),
    labels = scales::percent
  ) +
  theme_apa() +
  theme(text = element_text(family = "serif")) +
  ggtitle("B")

result_accuracy / result_error
```

The overall correlation of the predicted property prices with socio-economic features is shown in Figure \@ref(fig:census-features-xgb)A. Similar to the model used for urban features, the model based on socio-economic features reveals that prices lower than €300,000 lead to the highest errors in over-valuation while prices higher than €300,000 lead to systematic under-valuation errors (Figure \@ref(fig:census-features-xgb)B).

```{r census-features-table, results="asis"}
table_census <- xgb.importance(model = xgmodel) %>%
  dplyr::filter(Gain > 0.01) %>%
  dplyr::arrange(desc(Gain)) %>%
  dplyr::mutate_if(is.numeric, scales::percent, accuracy = 0.1) %>%
  dplyr::mutate(
    `Feature Category` = case_when(
      Feature %in% carers ~ "carers",
      Feature %in% disabilty_age_group ~ "disabilty age group",
      Feature %in% general_health ~ "general health",
      Feature %in% housing_occupancy ~ "housing occupancy",
      Feature %in% housing_rooms ~ "housing rooms",
      Feature %in% housing_tenure ~ "housing tenure",
      Feature %in% housing_type ~ "housing type",
      Feature %in% population ~ "population",
      Feature %in% religion ~ "religion"
    )
  )  %>% 
  dplyr::mutate(
    Feature = str_remove(Feature, "\\(Households\\)"),
    Feature = str_remove(Feature, "% "),
  ) %>%
  dplyr::select(
    `Feature Category`,
    `Feature Type` = Feature,
    Importance = Gain
  )

table_census %>%
  knitr::kable(
    "latex",
    caption = "(ref:census-features-table)",
    booktabs = T,
    linesep = ""
  ) %>%
  kableExtra::kable_styling(font_size = 8)
```

According to the analysis of the relative importance of socio-economic features (Table \@ref(tab:census-features-table)), the most important socio-economic features are the proportion of large houses (i.e., houses with eight or more rooms) in the small area containing the property (`r table_census$Importance[1]`). It also appears that areas having a high proportion of young children (`r table_census$Importance[3]`), as well as the proportion of people reporting that they have no religion (`r table_census$Importance[2]`), influence the model.

# Discussion

While Dublin is a very specific city due to its low population density spread on a large surface, its housing market is one of the most expensive in Europe. The main reason is the structure of the market, mainly made of residential properties and few apartment buildings. In this study, the prices of 10387 properties sold in 2018 in the area of Dublin were analysed from the Property Services Regulatory Authority database in order to identify the potential factors correlating with these prices. The enforcement of a housing price registry has been shown as being an essential tool to regulate the housing market [@tomson2016property]. However, a major limitation of this study is the absence of characteristics for the property itself. Indeed, prices are mainly determined by factors such as the size of the property or the number of rooms.

Despite the absence of property characteristics, urban and socio-economic features were used to identify spatial correlates to the housing prices. Results revealed that proximity to embassies and grasslands is a driver of house prices. Embassies are in general located in the most expensive areas of cities which are also the most aesthetically pleasing and the most secure part of cities. In addition, the density of embassies and properties allocated to embassy staff reduces the density of property available on the market and drives their prices up. The distance to a park or a natural area is also a very important factor for house prices. Results also revealed that the density of large houses in the area and the proportion of individuals reporting having no religion are very important. Again, the size of the property sold was not included in the database; however, the higher the density of large houses in the area, the higher the probability for the property to be a large house. The second most important feature is the density of young children. While the literature is not conclusive of the link between wealth and family size, future research could investigate the mediation effect of house size between family size and property price. Finally, the relation with the expression of religious belief, or more precisely its absence, is a very interesting feature. Again, the relationship between religion and income is not confirmed by the literature. Whereas some studies reveal a positive correlation between religion and income [@guiso2003people; @elgin2013religion] others reveals no evidence of such relationship [@de2008does]. Here, it appears that prices of houses are higher in areas with a high density of people indicating that they have no religion.

# Conclusion

The evolution of housing prices is a pressing issue in most European capitals and specially in Dublin. Given their significant increase, houses are less and less affordable for individuals. By preforming a feature analysis with urban and socio-economic features, it is possible to evaluate and predict the potential price of a house. Indeed features such as the presence of embassies or parks are criteria that influence significantly the price of houses. Similarly, the characteristics of inhabitants in the area such as religion, health and age is correlated to the evolution of housing prices. These results allow an understanding of why some areas have higher prices than others.

# Acknowledgement

The authors would like to thank the developers of the following R packages used to process, analyse, display and report data: `r papaja::cite_r(file = "r-references.bib")`.

# Data availability

The R code and relevant data for statistical computing are available at https://github.com/damien-dupre/DSAA_2020.

# References

### 2020-06-01 dsaa_application_manuscript/dsaa_application_manuscript.tex
submission to dsaa

%% template.tex
%% from
%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference,final,]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at:
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%

%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

%% BEGIN MY ADDITIONS %%

\usepackage{longtable,booktabs}

\usepackage[unicode=true]{hyperref}

\hypersetup{
            pdftitle={Urban and socio-economic correlates of property price evolution:~Application to Dublin's costal area},
            pdfkeywords={Property price, Housing Market, Feature Analysis, Machine Learning, Geocoding},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{5}

% Pandoc syntax highlighting

% Pandoc header
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%% END MY ADDITIONS %%

\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Urban and socio-economic correlates of property price evolution:~Application to Dublin's costal area}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations

\author{

%% ---- classic IEEETrans wide authors' list ----------------
 % -- end affiliation.wide
%% ----------------------------------------------------------

%% ---- classic IEEETrans one column per institution --------
 %% -- beg if/affiliation.institution-columnar
\IEEEauthorblockN{
  %% -- beg for/affiliation.institution.author
Damien Dupré %% -- end for/affiliation.institution.author
}
\IEEEauthorblockA{Business School\\
Dublin City University\\
Glasnevin, Dublin 09
\\damien.dupre@dcu.ie
}
 %% -- end for/affiliation.institution
 %% -- end if/affiliation.institution-columnar
%% ----------------------------------------------------------

%% ---- one column per author, classic/default IEEETrans ----
 %% -- end if/affiliation.institution-columnar
%% ----------------------------------------------------------

}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
%
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3},
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Understanding the characteristics of the housing market is essential for both sellers and buyers. However, the housing market is influenced by multiple factors. In this paper, the urban and socio-economic structure of an area is used to predict the price of 10389 houses sold in 2018 in the city of Dublin. More precisely, the direct distance from each house to 160 urban features taken from OpenStreetMap is calculated, and an extreme gradient boosting linear regression performed. Using these features, the model explains 47\% of the housing price variance. The most important features in this model are the proximity to an embassy and to a grassland. In addition, the results of a population census from 2016 are also used to correlate with the price of houses. From this census, 48 features are used as the input of a gradient boost linear regression model. In all, the socio-economic features are explaining 42\% of the housing price variance as well. The most important socio-economic features are; the density of houses having more than eight or more rooms in the area, the density of young children, and the density of individuals reporting that they have no religion. By taking into account either urban or socio-economic features, it is possible to accurately estimate housing prices and to predict their evolution.
\end{abstract}

% keywords
\begin{IEEEkeywords}
Property price; Housing Market; Feature Analysis; Machine Learning; Geocoding
\end{IEEEkeywords}

% use for special paper notices

% make the title area
\maketitle

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

First time property acquisition is an important achievement in individuals' lifetime. It provides not only housing security but also the feeling of being a landowner. However, access to the status of landowner is a challenge for low to middle income earners as it corresponds to significant spending which will potentially have budget impacts for years (Savage et al. 2014). For this reason, understanding property price markets as well as factors driving price is essential to identify investment opportunities.

Beside their inherent characteristics such as size, design and materials (Bourassa, Cantoni, and Hoesli 2007; Liu 2013), property valuation is dictated by multiple external factors explaining their spatial autocorrelation (Basu and Thibodeau 1998). Some of these factors are macroeconomic (Antonakakis and Floros 2016), related to the evolution of population general wealth while others are related to the characteristics of the neighbourhood (Dubin 1998). These characteristics include the urban and socio economic structure of the neighbourhood (Goodman and Thibodeau 1998). However, it is difficult to evaluate the influence of such external factors on property valuation (Clapp, Kim, and Gelfand 2002). By using machine learning algorithms, prediction of house prices has become more and more accurate. Techniques used usually include artificial neural networks due to the volatility of the market (Limsombunchai 2004; Feng and Jones 2015; Selim 2009).

In this paper, an alternative machine learning algorithm is used to identify urban and socio-economic correlates of property price evolution in the area of Dublin, Ireland in 2018. Due to the high number of potential features, an eXtreme Gradient Boosting (XGBoost) regression model is believed to obtain the best results with low parameter customisation and high speed.

\hypertarget{method}{%
\section{Method}\label{method}}

Since 2010, under the Irish Property Services (Regulation) Act, all individuals acquiring a property in Ireland have to declare it to Property Services Regulatory Authority (PSRA). They must provide details such as the date of sale, the price and the address of all residential properties purchased in Ireland as declared to the Revenue Commissioners for stamp duty purposes (Authority 2020). Data is filed electronically by persons doing the conveyancing of the property on behalf of the purchaser, and it must be noted that errors may occur when the data is being filed.

Because of the evolution of urban planning and the evolution of socio-economic features measured by the 2016 census, only houses sold in 2018 were processed to avoid potential temporal incoherences. In 2018, a total of 10395 property in the area of Dublin were sold, and these constitute the database used to identify the urban and socio-economic correlates of property prices.

In order to evaluate the spacial distribution of the property sold, property addresses were geocoded (i.e., converted to latitude and longitude) using the OpenStreetMap API. OpenStreetMap is a collaborative project which aims to create and provide access to free editable maps of the world (Haklay and Weber 2008).

To estimate the geographical distribution of the price density, multiple methods such as multiple regression (McCluskey et al. 2000) or Bayesian smoothing (Clapp, Kim, and Gelfand 2002) have been employed. While these methods are efficient in a non-restricted space, they have limitations when they are used in coastal areas. To deal with the influence of boundaries on price estimation, a generalised additive model using soap film smoothing has been used (Wood, Bravington, and Hedley 2008).

\hypertarget{distance-to-urban-features}{%
\subsection{Distance to urban features}\label{distance-to-urban-features}}

OpenStreetMap combines information about more than 400 urban features including road information and building information to categorize features such as amenities, leisure or tourism structures (OpenStreetMap 2020). Among these features, only 160 are available or relevant to the Dublin Area. The distance between each property and the closest point corresponds to each of the 160 urban features.

\hypertarget{density-of-socio-economic-features}{%
\subsection{Density of socio-economic features}\label{density-of-socio-economic-features}}

In addition to the distance to every urban feature, each house sold in 2018 was related to 48 socio-economic features of the small area including the house, as measured by the Irish census 2016 (Office 2020). The results of Irish 2016 census consultation is accessible through both the Central Statistics Office and All-Island Research Observatory. The data obtained can be mapped over small area boundaries which are fractions of Irish Electoral Division map. The social features extracted are corresponding to population information, religion, carers and health. Economic features correspond to the characteristics of each small area; including the proportion of housing types, number of rooms, occupancy and tenure per small area. Each property is then associated to the value corresponding to its small area. For anonymisation purposes, the results from small areas having less than five respondents to the census were converted to a proportion of five respondents.

\hypertarget{gradient-boosting-regressions}{%
\subsection{Gradient Boosting Regressions}\label{gradient-boosting-regressions}}

Lastly, an eXtreme Gradient Boosting (XGBoost) regression model was calculated to examine the extent to which house prices rely on urban and on socio-economic features. The XGBoost algorithm optimizes the prediction accuracy by performing iterative least-squares regressions (set to 100 iterations as the best trade-off between accuracy and speed), thereby minimizing the root mean squared error (Friedman 2001; Chen and Guestrin 2016). The original dataset has been randomly split to 80\% for training and 20\% for testing the models. Models accuracy is estimated using the coefficient of determination (\(R^2\)), the Root Mean Square Error (\(RMSE\)) and the Mean Absolute Error (\(MAE\)) of the predicted property price values of the test dataset.

\hypertarget{results}{%
\section{Results}\label{results}}

In 2018, the average price of a sold property in Dublin was €330,364 (SD = €180,448). In order to remove potential human errors and outliers, prices higher or lower than 1 SD were removed from the original dataset.

\begin{figure*}
\includegraphics[width=0.98\textwidth]{dsaa_application_manuscript_files/figure-latex/distrib-plot-1} \caption{Overall (A) and geographic (B) distribution of the houses prices density in the city of Dublin in 2018. The geographic distribution was obtained by using a generalized additive model with soap film smooth parameter to take into account the influence of coastal boundaries. Geographic estimates outside of the 95\% Confidence Interval were removed.}\label{fig:distrib-plot}
\end{figure*}

The density of housing prices reveals a unimodal distribution of property prices (Figure \ref{fig:distrib-plot}A). In addition, the geographical distribution of property prices indicates a higher valuation in the South-West of the city as well as on the coast line (Figure \ref{fig:distrib-plot}B).

\hypertarget{correlates-with-urban-features}{%
\subsection{Correlates with urban features}\label{correlates-with-urban-features}}

Using the XGBoost algorithm, the 160 urban features are explaining 47\% of the property price variance (\(F(1, 2076) = 1,840.91\), \(p < .001\)) with a \(RMSE\) of €124,364 and a \(MAE\) of €83,963.61.

\begin{figure}
\includegraphics[width=0.98\columnwidth]{dsaa_application_manuscript_files/figure-latex/osm-features-xgb-1} \caption{Property price prediction accuracy (A) and Property price prediction error (B) using urban features with XGBoost.}\label{fig:osm-features-xgb}
\end{figure}

The overall correlation of the predicted property prices with urban features is shown in Figure \ref{fig:osm-features-xgb}A. It can be noticed that property prices situated at the low and the high end of the range are the most difficult to predict (Figure \ref{fig:osm-features-xgb}B). A possible explanation is the absence of distinctive and recurrent patterns in urban features for these houses. However, the prices higher than €300,000 are potentially driving down the prediction accuracy due to outliers.

\begin{table}

\caption{\label{tab:osm-features-table}Urban features importance (higher than 1\%).}
\centering
\fontsize{8}{10}\selectfont
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\toprule
Feature Category & Feature Type & Importance\\
\midrule
amenity & embassy & 18.2\%\\
natural & grassland & 6.7\%\\
power & line & 2.3\%\\
route & bus & 2.2\%\\
boundary & administrative & 2.1\%\\
boundary & political & 1.7\%\\
route & ferry & 1.7\%\\
highway & residential & 1.3\%\\
highway & secondary & 1.2\%\\
historic & yes & 1.2\%\\
route & power & 1.1\%\\
cycleway & lane & 1.1\%\\
service & crossover & 1.1\%\\
barrier & wall & 1.1\%\\
place & locality & 1.1\%\\
area & yes & 1.0\%\\
\bottomrule
\end{tabu}
\end{table}

By analysing their importance (Table \ref{tab:osm-features-table}), the most relevant geographical features to predict housing prices are the distance to an embassy (18.2\%) and the distance to natural grasslands such as parks and gardens (6.7\%).

\hypertarget{correlates-with-socio-economic-features}{%
\subsection{Correlates with socio-economic features}\label{correlates-with-socio-economic-features}}

Using the XGBoost algorithm, the 48 socio-economic features explain 42\% of the property price variance (\(F(1, 2076) = 1,508.10\), \(p < .001\)) with a \(RMSE\) of €131,322 and a \(MAE\) of €87,328.79.

\begin{figure}
\includegraphics[width=0.98\columnwidth]{dsaa_application_manuscript_files/figure-latex/census-features-xgb-1} \caption{Property price prediction accuracy (A) and Property price prediction error (B) using socio-economic features with XGBoost.}\label{fig:census-features-xgb}
\end{figure}

The overall correlation of the predicted property prices with socio-economic features is shown in Figure \ref{fig:census-features-xgb}A. Similar to the model used for urban features, the model based on socio-economic features reveals that prices lower than €300,000 lead to the highest errors in over-valuation while prices higher than €300,000 lead to systematic under-valuation errors (Figure \ref{fig:census-features-xgb}B).

\begin{table}

\caption{\label{tab:census-features-table}Socio-economic features importance (higher than 1\%).}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{lll}
\toprule
Feature Category & Feature Type & Importance\\
\midrule
housing rooms & 8 or more Rooms & 29.0\%\\
religion & No Religion & 5.0\%\\
population & Age 0 - 14 & 3.8\%\\
housing rooms & 6 Rooms & 3.3\%\\
general health & Very Good & 3.1\%\\
population & Age 80 Plus & 2.8\%\\
housing tenure & Owner Occupier with Mortgage & 2.5\%\\
general health & Good & 2.4\%\\
religion & Other Catholic & 2.4\%\\
housing rooms & 5 Rooms & 2.3\%\\
disabilty age group & Persons with a disability aged 25 - 44 & 2.3\%\\
religion & No Answer & 2.2\%\\
housing rooms & 7 Rooms & 2.2\%\\
disabilty age group & Persons with a disability aged 0 - 14 & 2.0\%\\
population & Age 45 - 64 & 1.8\%\\
population & Age 15 - 24 & 1.6\%\\
disabilty age group & Persons with a disability aged 65 Plus & 1.6\%\\
housing rooms & 3 Rooms & 1.6\%\\
housing tenure & Social Rented & 1.6\%\\
disabilty age group & Persons with a disability aged 45 - 64 & 1.5\%\\
housing tenure & Rented Free of Rent & 1.4\%\\
disabilty age group & Persons with a disability aged 15 - 44 & 1.4\%\\
general health & Bad & 1.4\%\\
housing rooms & 1 Room & 1.4\%\\
housing tenure & Owner Occupier No Mortgage & 1.4\%\\
disabilty age group & Total persons with a disability & 1.2\%\\
housing rooms & 4 Rooms & 1.2\%\\
population & Female & 1.2\%\\
religion & Roman Catholic & 1.2\%\\
population & Age 25 - 44 & 1.1\%\\
carers & Provides No Care & 1.1\%\\
housing tenure & Private Rented & 1.0\%\\
\bottomrule
\end{tabular}
\end{table}

According to the analysis of the relative importance of socio-economic features (Table \ref{tab:census-features-table}), the most important socio-economic features are the proportion of large houses (i.e., houses with eight or more rooms) in the small area containing the property (29.0\%). It also appears that areas having a high proportion of young children (3.8\%), as well as the proportion of people reporting that they have no religion (5.0\%), influence the model.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

While Dublin is a very specific city due to its low population density spread on a large surface, its housing market is one of the most expensive in Europe. The main reason is the structure of the market, mainly made of residential properties and few apartment buildings. In this study, the prices of 10387 properties sold in 2018 in the area of Dublin were analysed from the Property Services Regulatory Authority database in order to identify the potential factors correlating with these prices. The enforcement of a housing price registry has been shown as being an essential tool to regulate the housing market (Tomson 2016). However, a major limitation of this study is the absence of characteristics for the property itself. Indeed, prices are mainly determined by factors such as the size of the property or the number of rooms.

Despite the absence of property characteristics, urban and socio-economic features were used to identify spatial correlates to the housing prices. Results revealed that proximity to embassies and grasslands is a driver of house prices. Embassies are in general located in the most expensive areas of cities which are also the most aesthetically pleasing and the most secure part of cities. In addition, the density of embassies and properties allocated to embassy staff reduces the density of property available on the market and drives their prices up. The distance to a park or a natural area is also a very important factor for house prices. Results also revealed that the density of large houses in the area and the proportion of individuals reporting having no religion are very important. Again, the size of the property sold was not included in the database; however, the higher the density of large houses in the area, the higher the probability for the property to be a large house. The second most important feature is the density of young children. While the literature is not conclusive of the link between wealth and family size, future research could investigate the mediation effect of house size between family size and property price. Finally, the relation with the expression of religious belief, or more precisely its absence, is a very interesting feature. Again, the relationship between religion and income is not confirmed by the literature. Whereas some studies reveal a positive correlation between religion and income (Guiso, Sapienza, and Zingales 2003; Elgin et al. 2013) others reveals no evidence of such relationship (De La O and Rodden 2008). Here, it appears that prices of houses are higher in areas with a high density of people indicating that they have no religion.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

The evolution of housing prices is a pressing issue in most European capitals and specially in Dublin. Given their significant increase, houses are less and less affordable for individuals. By preforming a feature analysis with urban and socio-economic features, it is possible to evaluate and predict the potential price of a house. Indeed features such as the presence of embassies or parks are criteria that influence significantly the price of houses. Similarly, the characteristics of inhabitants in the area such as religion, health and age is correlated to the evolution of housing prices. These results allow an understanding of why some areas have higher prices than others.

\hypertarget{acknowledgement}{%
\section{Acknowledgement}\label{acknowledgement}}

The authors would like to thank the developers of the following R packages used to process, analyse, display and report data: R (Version 4.0.0; R Core Team 2020) and the R-packages \emph{dplyr} (Version 1.0.0; Wickham, François, et al. 2020), \emph{forcats} (Version 0.5.0; Wickham 2020), \emph{ggplot2} (Version 3.3.1; Wickham 2016), \emph{here} (Version 0.1; Müller 2017), \emph{kableExtra} (Version 1.1.0; Zhu 2019), \emph{latex2exp} (Version 0.4.0; Meschiari 2015), \emph{lubridate} (Version 1.7.8; Grolemund and Wickham 2011), \emph{lwgeom} (Version 0.2.4; Pebesma 2020), \emph{magrittr} (Version 1.5; Bache and Wickham 2014), \emph{mgcv} (Version 1.8.31; Wood 2011, 2004, 2003; Wood et al. 2016), \emph{nlme} (Version 3.1.147; Pinheiro et al. 2020), \emph{OpenStreetMap} (Version 0.3.4; Fellows and JMapViewer library by Jan Peter Stotz 2019), \emph{osmdata} (Version 0.1.3; Padgham et al. 2017), \emph{papaja} (Version 0.1.0.9942; Aust and Barth 2020), \emph{patchwork} (Version 1.0.0; Pedersen 2019), \emph{purrr} (Version 0.3.4; Henry and Wickham 2020), \emph{readr} (Version 1.3.1; Wickham, Hester, and Francois 2018), \emph{sf} (Version 0.9.3; Pebesma 2018), \emph{stringr} (Version 1.4.0; Wickham 2019), \emph{tibble} (Version 3.0.1; Müller and Wickham 2020), \emph{tidyr} (Version 1.1.0; Wickham and Henry 2020), \emph{tidyverse} (Version 1.3.0; Wickham, Averick, et al. 2019), \emph{tmap} (Version 3.0; Tennekes 2018), \emph{xgboost} (Version 1.0.0.2; Chen et al. 2020), and \emph{yardstick} (Version 0.0.6; Kuhn and Vaughan 2020).

\hypertarget{data-availability}{%
\section{Data availability}\label{data-availability}}

The R code and relevant data for statistical computing are available at https://github.com/damien-dupre/DSAA\_2020.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-antonakakis2016dynamic}{}%
Antonakakis, Nikolaos, and Christos Floros. 2016. ``Dynamic Interdependencies Among the Housing Market, Stock Market, Policy Uncertainty and the Macroeconomy in the United Kingdom.'' \emph{International Review of Financial Analysis} 44: 111--22.

\leavevmode\hypertarget{ref-R-papaja}{}%
Aust, Frederik, and Marius Barth. 2020. \emph{papaja: Create APA Manuscripts with R Markdown}. \url{https://github.com/crsh/papaja}.

\leavevmode\hypertarget{ref-rppr2020}{}%
Authority, Property Services Regulatory. 2020. ``Residential Property Price Register.'' June 1, 2020. \url{https://propertypriceregister.ie}.

\leavevmode\hypertarget{ref-R-magrittr}{}%
Bache, Stefan Milton, and Hadley Wickham. 2014. \emph{Magrittr: A Forward-Pipe Operator for R}. \url{https://CRAN.R-project.org/package=magrittr}.

\leavevmode\hypertarget{ref-basu1998analysis}{}%
Basu, Sabyasachi, and Thomas G Thibodeau. 1998. ``Analysis of Spatial Autocorrelation in House Prices.'' \emph{The Journal of Real Estate Finance and Economics} 17 (1): 61--85.

\leavevmode\hypertarget{ref-bourassa2007spatial}{}%
Bourassa, Steven C, Eva Cantoni, and Martin Hoesli. 2007. ``Spatial Dependence, Housing Submarkets, and House Price Prediction.'' \emph{The Journal of Real Estate Finance and Economics} 35 (2): 143--60.

\leavevmode\hypertarget{ref-chen2016xgboost}{}%
Chen, Tianqi, and Carlos Guestrin. 2016. ``Xgboost: A Scalable Tree Boosting System.'' In \emph{Proceedings of the 22nd Acm Sigkdd International Conference on Knowledge Discovery and Data Mining}, 785--94.

\leavevmode\hypertarget{ref-R-xgboost}{}%
Chen, Tianqi, Tong He, Michael Benesty, Vadim Khotilovich, Yuan Tang, Hyunsu Cho, Kailong Chen, et al. 2020. \emph{Xgboost: Extreme Gradient Boosting}. \url{https://CRAN.R-project.org/package=xgboost}.

\leavevmode\hypertarget{ref-clapp2002predicting}{}%
Clapp, John M, Hyon--Jung Kim, and Alan E Gelfand. 2002. ``Predicting Spatial Patterns of House Prices Using Lpr and Bayesian Smoothing.'' \emph{Real Estate Economics} 30 (4): 505--32.

\leavevmode\hypertarget{ref-de2008does}{}%
De La O, Ana L, and Jonathan A Rodden. 2008. ``Does Religion Distract the Poor? Income and Issue Voting Around the World.'' \emph{Comparative Political Studies} 41 (4-5): 437--76.

\leavevmode\hypertarget{ref-dubin1998predicting}{}%
Dubin, Robin A. 1998. ``Predicting House Prices Using Multiple Listings Data.'' \emph{The Journal of Real Estate Finance and Economics} 17 (1): 35--59.

\leavevmode\hypertarget{ref-elgin2013religion}{}%
Elgin, Ceyhun, Turkmen Goksel, Mehmet Y Gurdal, and Cuneyt Orman. 2013. ``Religion, Income Inequality, and the Size of the Government.'' \emph{Economic Modelling} 30: 225--34.

\leavevmode\hypertarget{ref-R-OpenStreetMap}{}%
Fellows, Ian, and using the JMapViewer library by Jan Peter Stotz. 2019. \emph{OpenStreetMap: Access to Open Street Map Raster Images}. \url{https://CRAN.R-project.org/package=OpenStreetMap}.

\leavevmode\hypertarget{ref-feng2015comparing}{}%
Feng, Yingyu, and Kelvyn Jones. 2015. ``Comparing Multilevel Modelling and Artificial Neural Networks in House Price Prediction.'' In \emph{2015 2nd Ieee International Conference on Spatial Data Mining and Geographical Knowledge Services (Icsdm)}, 108--14. IEEE.

\leavevmode\hypertarget{ref-friedman2001greedy}{}%
Friedman, Jerome H. 2001. ``Greedy Function Approximation: A Gradient Boosting Machine.'' \emph{Annals of Statistics}, 1189--1232.

\leavevmode\hypertarget{ref-goodman1998housing}{}%
Goodman, Allen C, and Thomas G Thibodeau. 1998. ``Housing Market Segmentation.'' \emph{Journal of Housing Economics} 7 (2): 121--43.

\leavevmode\hypertarget{ref-R-lubridate}{}%
Grolemund, Garrett, and Hadley Wickham. 2011. ``Dates and Times Made Easy with lubridate.'' \emph{Journal of Statistical Software} 40 (3): 1--25. \url{http://www.jstatsoft.org/v40/i03/}.

\leavevmode\hypertarget{ref-guiso2003people}{}%
Guiso, Luigi, Paola Sapienza, and Luigi Zingales. 2003. ``People's Opium? Religion and Economic Attitudes.'' \emph{Journal of Monetary Economics} 50 (1): 225--82.

\leavevmode\hypertarget{ref-haklay2008openstreetmap}{}%
Haklay, Mordechai, and Patrick Weber. 2008. ``Openstreetmap: User-Generated Street Maps.'' \emph{IEEE Pervasive Computing} 7 (4): 12--18.

\leavevmode\hypertarget{ref-R-purrr}{}%
Henry, Lionel, and Hadley Wickham. 2020. \emph{Purrr: Functional Programming Tools}. \url{https://CRAN.R-project.org/package=purrr}.

\leavevmode\hypertarget{ref-R-yardstick}{}%
Kuhn, Max, and Davis Vaughan. 2020. \emph{Yardstick: Tidy Characterizations of Model Performance}. \url{https://CRAN.R-project.org/package=yardstick}.

\leavevmode\hypertarget{ref-limsombunchai2004house}{}%
Limsombunchai, Visit. 2004. ``House Price Prediction: Hedonic Price Model Vs. Artificial Neural Network.'' In \emph{New Zealand Agricultural and Resource Economics Society Conference}, 25--26.

\leavevmode\hypertarget{ref-liu2013spatial}{}%
Liu, Xiaolong. 2013. ``Spatial and Temporal Dependence in House Price Prediction.'' \emph{The Journal of Real Estate Finance and Economics} 47 (2): 341--69.

\leavevmode\hypertarget{ref-mccluskey2000application}{}%
McCluskey, William J, William G Deddis, Ian G Lamont, and Richard A Borst. 2000. ``The Application of Surface Generated Interpolation Models for the Prediction of Residential Property Values.'' \emph{Journal of Property Investment \& Finance}.

\leavevmode\hypertarget{ref-R-latex2exp}{}%
Meschiari, Stefano. 2015. \emph{Latex2exp: Use Latex Expressions in Plots}. \url{https://CRAN.R-project.org/package=latex2exp}.

\leavevmode\hypertarget{ref-R-here}{}%
Müller, Kirill. 2017. \emph{Here: A Simpler Way to Find Your Files}. \url{https://CRAN.R-project.org/package=here}.

\leavevmode\hypertarget{ref-R-tibble}{}%
Müller, Kirill, and Hadley Wickham. 2020. \emph{Tibble: Simple Data Frames}. \url{https://CRAN.R-project.org/package=tibble}.

\leavevmode\hypertarget{ref-cso2020}{}%
Office, Central Statistics. 2020. ``Census 2016 Small Area Population Statistics.'' June 1, 2020. \url{http://census.cso.ie/sapmap}.

\leavevmode\hypertarget{ref-openstreetmap2020}{}%
OpenStreetMap. 2020. ``Map Features (Wiki).'' June 1, 2020. \url{https://wiki.openstreetmap.org/wiki/Map_Features}.

\leavevmode\hypertarget{ref-R-osmdata}{}%
Padgham, Mark, Bob Rudis, Robin Lovelace, and Maëlle Salmon. 2017. ``Osmdata.'' \emph{The Journal of Open Source Software} 2 (14). \url{https://doi.org/10.21105/joss.00305}.

\leavevmode\hypertarget{ref-R-sf}{}%
Pebesma, Edzer. 2018. ``Simple Features for R: Standardized Support for Spatial Vector Data.'' \emph{The R Journal} 10 (1): 439--46. \url{https://doi.org/10.32614/RJ-2018-009}.

\leavevmode\hypertarget{ref-R-lwgeom}{}%
---------. 2020. \emph{Lwgeom: Bindings to Selected 'Liblwgeom' Functions for Simple Features}. \url{https://CRAN.R-project.org/package=lwgeom}.

\leavevmode\hypertarget{ref-R-patchwork}{}%
Pedersen, Thomas Lin. 2019. \emph{Patchwork: The Composer of Plots}. \url{https://CRAN.R-project.org/package=patchwork}.

\leavevmode\hypertarget{ref-R-nlme}{}%
Pinheiro, Jose, Douglas Bates, Saikat DebRoy, Deepayan Sarkar, and R Core Team. 2020. \emph{nlme: Linear and Nonlinear Mixed Effects Models}. \url{https://CRAN.R-project.org/package=nlme}.

\leavevmode\hypertarget{ref-R-base}{}%
R Core Team. 2020. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.

\leavevmode\hypertarget{ref-savage2014property}{}%
Savage, Michael, James Barlow, Peter Dickens, and Tom Fielding. 2014. \emph{Property Bureaucracy \& Culture}. Routledge.

\leavevmode\hypertarget{ref-selim2009determinants}{}%
Selim, Hasan. 2009. ``Determinants of House Prices in Turkey: Hedonic Regression Versus Artificial Neural Network.'' \emph{Expert Systems with Applications} 36 (2): 2843--52.

\leavevmode\hypertarget{ref-R-tmap}{}%
Tennekes, Martijn. 2018. ``tmap: Thematic Maps in R.'' \emph{Journal of Statistical Software} 84 (6): 1--39. \url{https://doi.org/10.18637/jss.v084.i06}.

\leavevmode\hypertarget{ref-tomson2016property}{}%
Tomson, AIVAR. 2016. ``Property Sales Register as a Tool to Improve the Quality of Valuation and Market Research: Lessons Learned from Selected Countries.'' In \emph{World Bank Conference on Land and Poverty, Washington, Dc, March}, 14--18.

\leavevmode\hypertarget{ref-R-ggplot2}{}%
Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.

\leavevmode\hypertarget{ref-R-stringr}{}%
---------. 2019. \emph{Stringr: Simple, Consistent Wrappers for Common String Operations}. \url{https://CRAN.R-project.org/package=stringr}.

\leavevmode\hypertarget{ref-R-forcats}{}%
---------. 2020. \emph{Forcats: Tools for Working with Categorical Variables (Factors)}. \url{https://CRAN.R-project.org/package=forcats}.

\leavevmode\hypertarget{ref-R-tidyverse}{}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019. ``Welcome to the tidyverse.'' \emph{Journal of Open Source Software} 4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\leavevmode\hypertarget{ref-R-dplyr}{}%
Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{Dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.

\leavevmode\hypertarget{ref-R-tidyr}{}%
Wickham, Hadley, and Lionel Henry. 2020. \emph{Tidyr: Tidy Messy Data}. \url{https://CRAN.R-project.org/package=tidyr}.

\leavevmode\hypertarget{ref-R-readr}{}%
Wickham, Hadley, Jim Hester, and Romain Francois. 2018. \emph{Readr: Read Rectangular Text Data}. \url{https://CRAN.R-project.org/package=readr}.

\leavevmode\hypertarget{ref-wood2008soap}{}%
Wood, Simon N, Mark V Bravington, and Sharon L Hedley. 2008. ``Soap Film Smoothing.'' \emph{Journal of the Royal Statistical Society: Series B (Statistical Methodology)} 70 (5): 931--55.

\leavevmode\hypertarget{ref-R-mgcv_d}{}%
Wood, S. N. 2003. ``Thin-Plate Regression Splines.'' \emph{Journal of the Royal Statistical Society (B)} 65 (1): 95--114.

\leavevmode\hypertarget{ref-R-mgcv_c}{}%
---------. 2004. ``Stable and Efficient Multiple Smoothing Parameter Estimation for Generalized Additive Models.'' \emph{Journal of the American Statistical Association} 99 (467): 673--86.

\leavevmode\hypertarget{ref-R-mgcv_a}{}%
---------. 2011. ``Fast Stable Restricted Maximum Likelihood and Marginal Likelihood Estimation of Semiparametric Generalized Linear Models.'' \emph{Journal of the Royal Statistical Society (B)} 73 (1): 3--36.

\leavevmode\hypertarget{ref-R-mgcv_b}{}%
Wood, S. N., N., Pya, and B. S"afken. 2016. ``Smoothing Parameter and Model Selection for General Smooth Models (with Discussion).'' \emph{Journal of the American Statistical Association} 111: 1548--75.

\leavevmode\hypertarget{ref-R-kableExtra}{}%
Zhu, Hao. 2019. \emph{KableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.

\end{document}

### 2020-06-01 script/census_dublin_features.R

@@ -1,0 +1,68 @@
+census_dublin_features <- c(
+carers <- c(              
+  "% Provides No Care",                          
+  "% 1-19 hours unpaid PW",                     
+  "% 20-49 hours unpaid PW",                     
+  "% 50+ hours unpaid PW",                      
+  "% Total Care Providers"                      
+),
+disabilty_age_group <- c(
+  "% Persons with a disability aged 0 - 14",      
+  "% Persons with a disability aged 15 - 44",    
+  "% Persons with a disability aged 25 - 44",   
+  "% Persons with a disability aged 45 - 64",    
+  "% Persons with a disability aged 65 Plus",   
+  "% Total persons with a disability" 
+),
+general_health <- c(
+  "% Very Good",                                
+  "% Good",                                     
+  "% Fair",                                     
+  "% Bad",                                       
+  "% Very Bad"
+),
+housing_occupancy <- c(
+  "% Occupied/HS With ususal Residents",         
+  "% Unoccupied/HS Without Ususal Residents" 
+),
+housing_rooms <- c(
+  "% 1 Room (Households)",                      
+  "% 2 Rooms (Households)",                      
+  "% 3 Rooms (Households)",                     
+  "% 4 Rooms (Households)",                      
+  "% 5 Rooms (Households)",                     
+  "% 6 Rooms (Households)",                      
+  "% 7 Rooms (Households)",                     
+  "% 8 or more Rooms (Households)"
+),
+housing_tenure <- c(
+  "% Owner Occupier with Mortgage (Households)",
+  "% Owner Occupier No Mortgage (Households)",   
+  "% Private Rented",                           
+  "% Social Rented",                             
+  "% Rented Free of Rent (Households)" 
+),                          
+housing_type <- c(
+  "% House/Bungalow",                           
+  "% Flat/Apartment/BedSit",                     
+  "% Caravan/Mobile home/Temperory" 
+),          
+population <- c(
+  "% Male",                                     
+  "% Female",                                  
+  "% Age 0 - 14",                                   
+  "% Age 15 - 24",                                 
+  "% Age 25 - 44",                                  
+  "% Age 45 - 64",                                  
+  "% Age 65 Plus",                                    
+  "% Age 15- 64",                                  
+  "% Age 80 Plus"
+),
+religion <- c(
+  "% Roman Catholic",                                 
+  "% Other Catholic",                                   
+  "% Other Religion",                                  
+  "% No Answer",                                       
+  "% No Religion"
+)
+)

### 2020-06-01 script/functions_project.R

@@ -1,0 +1,141 @@
+gam_density <- function (x, view = NULL, cond = list(), n.grid = 40, too.far = 0, 
+                    col = NA, color = "heat", contour.col = NULL, se = -1, type = "link", 
+                    plot.type = "persp", zlim = NULL, nCol = 50, ...) 
+{
+  fac.seq <- function(fac, n.grid) {
+    fn <- length(levels(fac))
+    gn <- n.grid
+    if (fn > gn) 
+      mf <- factor(levels(fac))[1:gn]
+    else {
+      ln <- floor(gn/fn)
+      mf <- rep(levels(fac)[fn], gn)
+      mf[1:(ln * fn)] <- rep(levels(fac), rep(ln, fn))
+      mf <- factor(mf, levels = levels(fac))
+    }
+    mf
+  }
+  dnm <- names(list(...))
+  v.names <- names(x$var.summary)
+  if (is.null(view)) {
+    k <- 0
+    view <- rep("", 2)
+    for (i in 1:length(v.names)) {
+      ok <- TRUE
+      if (is.matrix(x$var.summary[[i]])) 
+        ok <- FALSE
+      else if (is.factor(x$var.summary[[i]])) {
+        if (length(levels(x$var.summary[[i]])) <= 1) 
+          ok <- FALSE
+      }
+      else {
+        if (length(unique(x$var.summary[[i]])) == 1) 
+          ok <- FALSE
+      }
+      if (ok) {
+        k <- k + 1
+        view[k] <- v.names[i]
+      }
+      if (k == 2) 
+        break
+    }
+    if (k < 2) 
+      stop("Model does not seem to have enough terms to do anything useful")
+  }
+  ok <- TRUE
+  for (i in 1:2) if (is.factor(x$var.summary[[view[i]]])) {
+    if (length(levels(x$var.summary[[view[i]]])) <= 1) 
+      ok <- FALSE
+  }
+  else {
+    if (length(unique(x$var.summary[[view[i]]])) <= 1) 
+      ok <- FALSE
+  }
+  if (!ok) 
+    stop(gettextf("View variables must contain more than one value. view = c(%s,%s).", 
+                  view[1], view[2]))
+  if (is.factor(x$var.summary[[view[1]]])) 
[…]

### 2020-06-01 script/osm_dublin_features.R

@@ -1,0 +1,233 @@
+osm_dublin_features <- c(
+amenity <- c(
+  "amenity_bar",
+  "amenity_college",                             
+  "amenity_school",
+  "amenity_university",                          
+  "amenity_bicycle_parking",
+  "amenity_fuel",                               
+  "amenity_parking",
+  "amenity_community_centre",                    
+  "amenity_bench",
+  "amenity_embassy",                            
+  "amenity_police",
+  "amenity_prison",                             
+  "amenity_recycling"
+),
+barrier <- c(
+  "barrier_city_wall",                           
+  "barrier_ditch",                              
+  "barrier_fence",                               
+  "barrier_guard_rail",                         
+  "barrier_hedge",                              
+  "barrier_kerb",                               
+  "barrier_retaining_wall",                      
+  "barrier_wall",                               
+  "barrier_block",                               
+  "barrier_bollard",                            
+  "barrier_chain",                               
+  "barrier_full-height_turnstile",              
+  "barrier_gate",                                
+  "barrier_jersey_barrier",                     
+  "barrier_yes"
+),
+boundary <- c(
+  "boundary_administrative",                    
+  "boundary_historic",                           
+  "boundary_political",                         
+  "boundary_postal_code",                        
+  "boundary_protected_area"
+),
+building <- c(
+  "building_apartments",                         
+  "building_house",                             
+  "building_residential",                        
+  "building_commercial",                        
+  "building_industrial",                         
+  "building_retail",                           
+  "building_hospital",                           
+  "building_university",                        
+  "building_yes"
+),
+highway <- c(
+  "highway_motorway",                           
+  "highway_trunk",                              
+  "highway_secondary",                          
+  "highway_tertiary",                            
+  "highway_unclassified",                       
+  "highway_residential",                         
+  "highway_service",                            
+  "highway_motorway_link",                       
+  "highway_trunk_link",                         
+  "highway_secondary_link",                      
+  "highway_tertiary_link",                      
+  "highway_pedestrian",                          
+  "highway_track",                              
+  "highway_road",                                
+  "highway_footway",                            
+  "highway_steps",                               
+  "highway_path",                               
+  "highway_cycleway"
+),
+cycleway <- c(
+  "cycleway_lane",                              
+  "cycleway_opposite",                          
+  "cycleway_opposite_lane",                     
[…]

### 2020-08-06 dsaa_application_manuscript/dsaa_application_manuscript.Rmd
renv

@@
 # libraries --------------------------------------------------------------------
+library(rticles)
 library(tidyverse)
@@
   label = train_tbl_y,
-  objective = "reg:linear",
+  objective = "reg:squarederror",
   eval_metric = "rmse",
@@
 lm_result <- lm(pred ~ price, result) %>%
-  papaja::apa_print.lm()
+  papaja::apa_print()
 ```

### 2020-08-06 dsaa_application_manuscript/dsaa_application_manuscript.tex
renv

@@
 
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {\par}
+
 % Pandoc header
@@
 
-Using the XGBoost algorithm, the 160 urban features are explaining 47\% of the property price variance (\(F(1, 2076) = 1,840.91\), \(p < .001\)) with a \(RMSE\) of €124,364 and a \(MAE\) of €83,963.61.
+Using the XGBoost algorithm, the 160 urban features are explaining 44\% of the property price variance (\(F(1, 2076) = 1,629.56\), \(p < .001\)) with a \(RMSE\) of €127,783 and a \(MAE\) of €85,115.25.
 
@@
 \midrule
-amenity & embassy & 18.2\%\\
+amenity & embassy & 17.7\%\\
 natural & grassland & 6.7\%\\
-power & line & 2.3\%\\
-route & bus & 2.2\%\\
-boundary & administrative & 2.1\%\\
-boundary & political & 1.7\%\\
-route & ferry & 1.7\%\\
-highway & residential & 1.3\%\\
-highway & secondary & 1.2\%\\
-historic & yes & 1.2\%\\
-route & power & 1.1\%\\
-cycleway & lane & 1.1\%\\
-service & crossover & 1.1\%\\
-barrier & wall & 1.1\%\\
+route & bus & 2.3\%\\
+power & line & 2.2\%\\
+boundary & administrative & 1.5\%\\
+place & island & 1.5\%\\
+barrier & wall & 1.5\%\\
+cycleway & share & 1.4\%\\
+route & ferry & 1.3\%\\
+cutting & yes & 1.1\%\\
+amenity & bar & 1.1\%\\
+barrier & yes & 1.1\%\\
 place & locality & 1.1\%\\
-area & yes & 1.0\%\\
+boundary & political & 1.1\%\\
+route & road & 1.1\%\\
+highway & path & 1.1\%\\
+tunnel & yes & 1.0\%\\
+boundary & postal & 1.0\%\\
 \bottomrule
@@
 
-By analysing their importance (Table \ref{tab:osm-features-table}), the most relevant geographical features to predict housing prices are the distance to an embassy (18.2\%) and the distance to natural grasslands such as parks and gardens (6.7\%).
+By analysing their importance (Table \ref{tab:osm-features-table}), the most relevant geographical features to predict housing prices are the distance to an embassy (17.7\%) and the distance to natural grasslands such as parks and gardens (6.7\%).
 
@@
 
-Using the XGBoost algorithm, the 48 socio-economic features explain 42\% of the property price variance (\(F(1, 2076) = 1,508.10\), \(p < .001\)) with a \(RMSE\) of €131,322 and a \(MAE\) of €87,328.79.
+Using the XGBoost algorithm, the 48 socio-economic features explain 42\% of the property price variance (\(F(1, 2076) = 1,500.87\), \(p < .001\)) with a \(RMSE\) of €131,270 and a \(MAE\) of €86,891.85.
 
@@
 \midrule
-housing rooms & 8 or more Rooms & 29.0\%\\
-religion & No Religion & 5.0\%\\
-population & Age 0 - 14 & 3.8\%\\
-housing rooms & 6 Rooms & 3.3\%\\
-general health & Very Good & 3.1\%\\
-population & Age 80 Plus & 2.8\%\\
+housing rooms & 8 or more Rooms & 28.2\%\\
+population & Age 0 - 14 & 4.5\%\\
+religion & No Religion & 4.4\%\\
+population & Age 80 Plus & 3.6\%\\
+general health & Very Good & 3.5\%\\
+housing rooms & 6 Rooms & 3.2\%\\
 housing tenure & Owner Occupier with Mortgage & 2.5\%\\
-general health & Good & 2.4\%\\
-religion & Other Catholic & 2.4\%\\
-housing rooms & 5 Rooms & 2.3\%\\
-disabilty age group & Persons with a disability aged 25 - 44 & 2.3\%\\
-religion & No Answer & 2.2\%\\
-housing rooms & 7 Rooms & 2.2\%\\
-disabilty age group & Persons with a disability aged 0 - 14 & 2.0\%\\
-population & Age 45 - 64 & 1.8\%\\
-population & Age 15 - 24 & 1.6\%\\
-disabilty age group & Persons with a disability aged 65 Plus & 1.6\%\\
-housing rooms & 3 Rooms & 1.6\%\\
-housing tenure & Social Rented & 1.6\%\\
-disabilty age group & Persons with a disability aged 45 - 64 & 1.5\%\\
-housing tenure & Rented Free of Rent & 1.4\%\\
-disabilty age group & Persons with a disability aged 15 - 44 & 1.4\%\\
-general health & Bad & 1.4\%\\
+disabilty age group & Persons with a disability aged 65 Plus & 2.3\%\\
+housing rooms & 5 Rooms & 2.2\%\\
+housing tenure & Social Rented & 2.2\%\\
+housing rooms & 4 Rooms & 2.1\%\\
+religion & Other Catholic & 2.0\%\\
+housing rooms & 7 Rooms & 1.9\%\\
+housing rooms & 3 Rooms & 1.8\%\\
+religion & No Answer & 1.8\%\\
+housing tenure & Private Rented & 1.6\%\\
+population & Age 45 - 64 & 1.5\%\\
+population & Age 25 - 44 & 1.5\%\\
+housing tenure & Owner Occupier No Mortgage & 1.5\%\\
+disabilty age group & Persons with a disability aged 0 - 14 & 1.5\%\\
+disabilty age group & Persons with a disability aged 25 - 44 & 1.4\%\\
 housing rooms & 1 Room & 1.4\%\\
-housing tenure & Owner Occupier No Mortgage & 1.4\%\\
+carers & Provides No Care & 1.3\%\\
+religion & Roman Catholic & 1.3\%\\
+general health & Good & 1.3\%\\
+population & Female & 1.3\%\\
+general health & Fair & 1.3\%\\
 disabilty age group & Total persons with a disability & 1.2\%\\
-housing rooms & 4 Rooms & 1.2\%\\
-population & Female & 1.2\%\\
-religion & Roman Catholic & 1.2\%\\
-population & Age 25 - 44 & 1.1\%\\
-carers & Provides No Care & 1.1\%\\
-housing tenure & Private Rented & 1.0\%\\
+general health & Bad & 1.2\%\\
+religion & Other Religion & 1.2\%\\
+disabilty age group & Persons with a disability aged 45 - 64 & 1.2\%\\
+carers & Total Care Providers & 1.2\%\\
+population & Age 15 - 24 & 1.1\%\\
+carers & 20-49 hours unpaid PW & 1.1\%\\
+housing tenure & Rented Free of Rent & 1.0\%\\
 \bottomrule
@@
 
-According to the analysis of the relative importance of socio-economic features (Table \ref{tab:census-features-table}), the most important socio-economic features are the proportion of large houses (i.e., houses with eight or more rooms) in the small area containing the property (29.0\%).
-It also appears that areas having a high proportion of young children (3.8\%), as well as the proportion of people reporting that they have no religion (5.0\%), influence the model.
+According to the analysis of the relative importance of socio-economic features (Table \ref{tab:census-features-table}), the most important socio-economic features are the proportion of large houses (i.e., houses with eight or more rooms) in the small area containing the property (28.2\%).
+It also appears that areas having a high proportion of young children (4.4\%), as well as the proportion of people reporting that they have no religion (4.5\%), influence the model.
 
@@
 
-The authors would like to thank the developers of the following R packages used to process, analyse, display and report data: R (Version 4.0.0; R Core Team 2020) and the R-packages \emph{dplyr} (Version 1.0.0; Wickham, François, et al.
-2020), \emph{forcats} (Version 0.5.0; Wickham 2020), \emph{ggplot2} (Version 3.3.1; Wickham 2016), \emph{here} (Version 0.1; Müller 2017), \emph{kableExtra} (Version 1.1.0; Zhu 2019), \emph{latex2exp} (Version 0.4.0; Meschiari 2015), \emph{lubridate} (Version 1.7.8; Grolemund and Wickham 2011), \emph{lwgeom} (Version 0.2.4; Pebesma 2020), \emph{magrittr} (Version 1.5; Bache and Wickham 2014), \emph{mgcv} (Version 1.8.31; Wood 2011, 2004, 2003; Wood et al.
-2016), \emph{nlme} (Version 3.1.147; Pinheiro et al.
+The authors would like to thank the developers of the following R packages used to process, analyse, display and report data: R (Version 4.0.2; R Core Team 2020) and the R-packages \emph{dplyr} (Version 1.0.1; Wickham, François, et al.
+2020), \emph{forcats} (Version 0.5.0; Wickham 2020), \emph{ggplot2} (Version 3.3.2; Wickham 2016), \emph{here} (Version 0.1; Müller 2017), \emph{kableExtra} (Version 1.1.0; Zhu 2019), \emph{latex2exp} (Version 0.4.0; Meschiari 2015), \emph{lubridate} (Version 1.7.9; Grolemund and Wickham 2011), \emph{lwgeom} (Version 0.2.5; Pebesma 2020), \emph{magrittr} (Version 1.5; Bache and Wickham 2014), \emph{mgcv} (Version 1.8.31; Wood 2011, 2004, 2003; Wood et al.
+2016), \emph{nlme} (Version 3.1.148; Pinheiro et al.
 2020), \emph{OpenStreetMap} (Version 0.3.4; Fellows and JMapViewer library by Jan Peter Stotz 2019), \emph{osmdata} (Version 0.1.3; Padgham et al.
-2017), \emph{papaja} (Version 0.1.0.9942; Aust and Barth 2020), \emph{patchwork} (Version 1.0.0; Pedersen 2019), \emph{purrr} (Version 0.3.4; Henry and Wickham 2020), \emph{readr} (Version 1.3.1; Wickham, Hester, and Francois 2018), \emph{sf} (Version 0.9.3; Pebesma 2018), \emph{stringr} (Version 1.4.0; Wickham 2019), \emph{tibble} (Version 3.0.1; Müller and Wickham 2020), \emph{tidyr} (Version 1.1.0; Wickham and Henry 2020), \emph{tidyverse} (Version 1.3.0; Wickham, Averick, et al.
-2019), \emph{tmap} (Version 3.0; Tennekes 2018), \emph{xgboost} (Version 1.0.0.2; Chen et al.
-2020), and \emph{yardstick} (Version 0.0.6; Kuhn and Vaughan 2020).
+2017), \emph{papaja} (Version 0.1.0.9997; Aust and Barth 2020), \emph{patchwork} (Version 1.0.1; Pedersen 2019), \emph{purrr} (Version 0.3.4; Henry and Wickham 2020), \emph{readr} (Version 1.3.1; Wickham, Hester, and Francois 2018), \emph{sf} (Version 0.9.5; Pebesma 2018), \emph{stringr} (Version 1.4.0; Wickham 2019), \emph{tibble} (Version 3.0.3; Müller and Wickham 2020), \emph{tidyr} (Version 1.1.0; Wickham and Henry 2020), \emph{tidyverse} (Version 1.3.0; Wickham, Averick, et al.
+2019), \emph{tmap} (Version 3.1; Tennekes 2018), \emph{xgboost} (Version 1.1.1.1; Chen et al.
+2020), and \emph{yardstick} (Version 0.0.7; Kuhn and Vaughan 2020).
 
@@
 \hypertarget{refs}{}
+\begin{cslreferences}
 \leavevmode\hypertarget{ref-antonakakis2016dynamic}{}%
@@
 Zhu, Hao. 2019. \emph{KableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
+\end{cslreferences}
 

### 2020-08-14 manuscript/manuscript.Rmd
reviewer answer startup

--- 2020-08-06 dsaa_application_manuscript/dsaa_application_manuscript.Rmd
@@
 
-bibliography: [references.bib, r-references.bib]
+bibliography: ["bib/references.bib", "bib/r-references.bib"]
 output:
@@
 source(here::here("script/osm_dublin_features.R"))
+
 # data path --------------------------------------------------------------------

### 2020-08-14 review/revision_letter.Rmd
reviewer answer startup

---
title       : "Urban and socio-economic correlates of property price evolution: Application to Dublin's costal area"
# authors     : "First Author & Second Author"
journal     : "Special Session-Environmental and Geo-spatial Data Analytics (EnGeoData) of the 7th IEEE International Conference on Data Science and Advanced Analytics (DSAA'2020)."
manuscript  : "Paper 147"

output      :
  bookdown::pdf_book:
    base_format: papaja::revision_letter_pdf
---

```{r setup, include = FALSE}
# options ----------------------------------------------------------------------
options(scipen = 999)
set.seed(999)
knitr::opts_chunk$set(
  cache.extra = knitr::rand_seed,
  cache = TRUE, 
  message = FALSE, 
  warning = FALSE, 
  error = FALSE, 
  echo = FALSE, 
  fig.retina = 3
  )

# libraries --------------------------------------------------------------------
library(tidyverse)
library(here)
```

Dear Antonio, David, Mathieu and Maguelonne,

Thank you to consider the manuscript entitled "Urban and socio-economic correlates of property price evolution: Application to Dublin's costal area" in EnGeoData 2020. We would also like to take this opportunity to express our thanks to the reviewers for the positive feedback and helpful comments for correction or modification.

We believe have resulted in an improved revised manuscript, which you will find uploaded alongside this document. The manuscript has been revised to address the reviewer comments, which are appended alongside our responses to this letter.

# Reviewer \#1

\RC{Page 1 : « 10389 houses » in the Abstract and « 10395 properties » in the Method section and « 10387 Properties » in Discussion section ?}

\RC{Page 2 : « an eXtreme Gradient Boosting (XGBoost) regression model was calculated to examine the extent to which house prices rely on urban and on socio-economic features ». Could you explain wether you worked on the XGBoost algorithm or not ? What was your contribution to develop/adapt/or parameterize this algorithm ?}

\RC{I suppose that Urban features importance is computed by the XGBoost model (Table I). Could you explain this stage and comment results ? Embassy importance is detected thanks to statistical computations. Is it possible to comment and validate it by qualitative interviews with experts ?}

\RC{Page 4 : Similarly, I couldn’t understand how Socio-Economic features importance is computed (Table II). Could you explain this stage and comment results ? Religion importance is detected thanks to statistical computations. Is it possible to comment and validate it by qualitative interviews with experts ?}

\RC{« Embassies are in general located in the most expensive areas of cities ». Are the ambassies installed in expensive areas or are the areas expensive because of their embassies ?}

\RC{Globally, it is difficult to evaluate the added-value of this research work :

  - Is it a proposal of a new regression model ? If yes, you should better explain XGBoost and its originality
  - Is it a experiment of an existing regression model applied to housing prices ? If yes, you should comment the results on the 20% of the original dataset : are the price predictions relevant ? Is this approach better than existing ones for price prediction ? Number of parameters ? Computation speed ? etc.}

\RC{What was not possible before and what is possible now thanks to this work ?}

\RC{Typos : Page 1 : « 10395 property »}

# Reviewer \#2

\RC{This paper presents machine learning experiments to estimate price changes in the real estate market.

The main method is the XGBoost algorithm. Perhaps more methods could have been tried, including some simple baseline to put the presented results into perspective.

The focus of the paper in on the contribution of the proposed features.

There in not much novelty, but the experiments are well executed and the results are convincing. Real data is used for training and testing.}

# Reviewer \#3

\RC{The topic of the paper fits well with the special session main topics. Below some suggestions of improvement that can be made to this contribution.}

\RC{Concerning the following statement: “techniques used usually include artificial neural networks due to the volatility of the market”, even though it is backed up by bibliographic references, it would be worth further explaining it. The same applies to the following statement "While these methods are efficient in a non-restricted space, they have limitations when they are used in coastal areas.", explain why.}

\RC{There are some issues related to the data acquisition process and the possible presence of errors. You noted that errors may occur when the data is being filed, could you provide some calculated information about the quality of the data used based on missing values, incoherent values, etc. Also, property addresses were geocoded relying on OSM geocoder, can you estimate accuracy of the process? The quality of OSM data is usually weak, what are the sources reported on the OSM urban features you extracted (source key: https://wiki.openstreetmap.org/wiki/Key:source)?}

\RC{Input data is not available in the github repository given by the authors, can we suppose it is due to restrictions to share the data, if so, this should be said clearly.}

\RC{Add URL to point to the Central Statistics Office and All-Island Research Observatory where data is available.}

\RC{explicit acronym "SD".}

### 2020-08-18 manuscript/manuscript.Rmd
reviewer changes

@@
 ---
-title: Urban and socio-economic correlates of property price evolution\:\ Application to Dublin's costal area
+title: Urban and socio-economic correlates of property prices in Dublin's area
 affiliation:
@@
 However, the housing market is influenced by multiple factors.
-In this paper, the urban and socio-economic structure of an area is used to predict the price of 10389 houses sold in 2018 in the city of Dublin.
-More precisely, the direct distance from each house to 160 urban features taken from OpenStreetMap is calculated, and an extreme gradient boosting linear regression performed.
-Using these features, the model explains 47\% of the housing price variance.
+In this paper, the urban and socio-economic structure of an area is used to predict the price of 10387 properties sold in 2018 in the city of Dublin.
+More precisely, the direct distance from each property to 160 urban features taken from OpenStreetMap is calculated, and an extreme gradient boosting linear regression performed.
+Using these features, the model explains 45\% of the housing price variance.
 The most important features in this model are the proximity to an embassy and to a grassland.
-In addition, the results of a population census from 2016 are also used to correlate with the price of houses.
+In addition, the results of a population census from 2016 are also used to correlate with the price of properties.
 From this census, 48 features are used as the input of a gradient boost linear regression model.
-In all, the socio-economic features are explaining 42\% of the housing price variance as well.
-The most important socio-economic features are; the density of houses having more than eight or more rooms in the area, the density of young children, and the density of individuals reporting that they have no religion.
+In all, the socio-economic features are explaining 43\% of the housing price variance as well.
+The density of individuals reporting that they are not providing unpaid personal help for a friend or family member as well as individuals reporting that they have no religion are the most important socio-economic feratures.
 By taking into account either urban or socio-economic features, it is possible to accurately estimate housing prices and to predict their evolution.
@@
 
-# output: word_document
-
 header-includes:
   - \usepackage{graphicx}
+  - \usepackage{flushend}
 ---
 
-(ref:distrib-plot) Overall (A) and geographic (B) distribution of the houses prices density in the city of Dublin in 2018.
-The geographic distribution was obtained by using a generalized additive model with soap film smooth parameter to take into account the influence of coastal boundaries.
-Geographic estimates outside of the 95\% Confidence Interval were removed.
-(ref:osm-features-xgb) Property price prediction accuracy (A) and Property price prediction error (B) using urban features with XGBoost.
-(ref:osm-features-table) Urban features importance (higher than 1\%).
-(ref:census-features-xgb) Property price prediction accuracy (A) and Property price prediction error (B) using socio-economic features with XGBoost.
-(ref:census-features-table) Socio-economic features importance (higher than 1\%).
+(ref:distrib-plot) Overall distribution of the property prices density in the city of Dublin in 2018
+(ref:distrib-gps) Geographic distribution of the property prices density in the city of Dublin in 2018 using a generalized additive model with soap film smooth parameter to take into account the influence of coastal boundaries.
+Estimates outside of the 95\% Confidence Interval were removed.
+(ref:osm-model-comparison) Accuracy of XGBoost models using urban features according the type of gradient boosting method and output objective.
+(ref:osm-features-xgb) Property price prediction accuracy (A) and prediction error (B) using urban features with XGBoost.
+(ref:osm-features-table) Top 10 most important urban features contributing to the property price prediction.
+(ref:census-model-comparison) Accuracy of XGBoost models using socio-economic features according the type of gradient boosting method and output objective.
+(ref:census-features-xgb) Property price prediction accuracy (A) and prediction error (B) using socio-economic features with XGBoost.
+(ref:census-features-table) Top 10 most important socio-economic features contributing to the property price prediction.
 
@@
 # options ----------------------------------------------------------------------
-set.seed(123)
-options(scipen=999)
+base::set.seed(123)
+base::options(
+  scipen = 999,
+  knitr.kable.NA = " - "
+  )
 knitr::opts_chunk$set(
@@
   error = FALSE, 
-  echo = FALSE
+  echo = FALSE,
+  cache = TRUE,
+  fig.retina = 3
   )
@@
 
-# functions --------------------------------------------------------------------
+# sources ----------------------------------------------------------------------
 source(here::here("script/functions_project.R"))
 source(here::here("script/osm_dublin_features.R"))
+source(here::here("script/census_dublin_features.R"))
+
+# data -------------------------------------------------------------------------
+geocode_dublin_2018 <- here::here("data/ppr_dublin_2018.csv") %>% 
+  readr::read_csv()
+
+house_dublin_2018 <- geocode_dublin_2018 %>% 
+  dplyr::filter(!grepl('apartment|Apartment|APART|APT|Apt', Address))
+
+final_dublin_2018 <- here::here("data/data_dublin_2018_osm_features.rds") %>% 
+  readr::read_rds()
 
@@
     year_fact = as.factor(as.character(year))
-  )
+  ) %>% 
+  dplyr::filter(year == 2018)
 
@@
 # dublin map -------------------------------------------------------------------
-# boundaries_dublin <-
-#   osmdata::opq(bbox = 'Dublin, Ireland') %>%
-#   osmdata::add_osm_feature(key = 'admin_level', value = '7') %>%
-#   osmdata::osmdata_sf() %>%
-#   osmdata::unique_osmdata()
-# 
-# bb <- st_bbox(boundaries_dublin$osm_multipolygons$geometry)
-# # Defining a buffer
-# buffer <- 0.1
-# p_big <- rbind(
-#   c(bb[1] - buffer, bb[2] - buffer),
-#   c(bb[1] - buffer, bb[4] + buffer),
-#   c(bb[3] + buffer, bb[4] + buffer),
-#   c(bb[3] + buffer, bb[2] - buffer),
-#   c(bb[1] - buffer, bb[2] - buffer)
-#   )
-# # Putting the coordinates into a squared polygon object
-# pol_dublin <- st_polygon(list(p_big)) %>% st_geometry
-# # Providing the SRID (here, unprojected lon/lat)
-# st_crs(pol_dublin) <- 4326
-# 
-# # get Dublin coast line ------------------------------------------------------
-# coast_dublin <- opq(bbox = st_bbox(pol_dublin)) %>%
-#   add_osm_feature(key = 'natural', value = 'coastline') %>%
-#   osmdata_sf %>% unique_osmdata
-# 
-# blade_dublin <- coast_dublin$osm_lines$geometry %>% st_union %>% st_line_merge
-# 
-# ls <- st_linestring(blade_dublin[[1]][[1]])
-# 
-# dublin_multipol <- st_split(st_geometry(pol_dublin), st_geometry(ls))
-# 
-# readr::write_rds(dublin_multipol, here::here("data/dublin_multipol.rds"))
 dublin_multipol <- readr::read_rds(here::here("data/dublin_multipol.rds"))
@@
 However, it is difficult to evaluate the influence of such external factors on property valuation [@clapp2002predicting].
-By using machine learning algorithms, prediction of house prices has become more and more accurate.
+By using machine learning algorithms, prediction of property prices has become more and more accurate.
 Techniques used usually include artificial neural networks due to the volatility of the market [@limsombunchai2004house; @feng2015comparing; @selim2009determinants].
+Indeed, artificial neural network techniques can quickly adjust to volatility changes, therefore incorporating sharp increases or decreases in their predictions [@aragones2007neural].
+However, the model obtained from most artificial neural networks are hardly interpretable [@rudin2019stop].
 
 In this paper, an alternative machine learning algorithm is used to identify urban and socio-economic correlates of property price evolution in the area of Dublin, Ireland in 2018.
-Due to the high number of potential features, an eXtreme Gradient Boosting (XGBoost) regression model is believed to obtain the best results with low parameter customisation and high speed.
+Due to the high number of potential features, an eXtreme Gradient Boosting (XGBoost) regression model is believed to obtain the best results with low parameter customisation and high speed [@chen2016xgboost].
+Contrary to most artificial neural network techniques, XGBoost is an interpretation-focused method which automatically provides estimates of feature importance from a trained predictive model.
 
@@
 
-Because of the evolution of urban planning and the evolution of socio-economic features measured by the 2016 census, only houses sold in 2018 were processed to avoid potential temporal incoherences.
-In 2018, a total of 10395 property in the area of Dublin were sold, and these constitute the database used to identify the urban and socio-economic correlates of property prices.
+Because of the evolution of urban planning and the evolution of socio-economic features measured by the 2016 census, only properties sold in 2018 were processed to avoid potential temporal incoherence.
+The total amount of property sold in 2018 in the county Dublin which appears in the Property Price Register is `r nrow(geocode_dublin_2018)`.
+From these, all property corresponding to apartments were removed (`r scales::percent(1-(nrow(house_dublin_2018)/nrow(geocode_dublin_2018)), accuracy = 0.1)` of the original dataset).
+Finally, unsuccessfully geocoded properties were also removed from the original dataset (`r scales::percent(1-(nrow(final_dublin_2018)/nrow(house_dublin_2018)), accuracy = 0.1)`) to obtain the `r nrow(final_dublin_2018)` of the final dataset used to identify the urban and socio-economic correlates of property prices.
 
-In order to evaluate the spacial distribution of the property sold, property addresses were geocoded (i.e., converted to latitude and longitude) using the OpenStreetMap API.
+In order to evaluate the spacial distribution of the property sold, property addresses were geocoded (i.e., converted to latitude and longitude) using the Nominatim API [@nominatim2020] to OpenStreetMap data.
 OpenStreetMap is a collaborative project which aims to create and provide access to free editable maps of the world [@haklay2008openstreetmap].
@@
 While these methods are efficient in a non-restricted space, they have limitations when they are used in coastal areas.
-To deal with the influence of boundaries on price estimation, a generalised additive model using soap film smoothing has been used [@wood2008soap].
+This problem is known as "finite area smoothing" and occurs when predictions from a model are approximated across geographical barriers, such as irregular coastal shapes, which would lead to poorly predicted values [@ramsay2002spline].
+To deal with the influence of geographical boundaries on price estimation, a generalised additive model using soap film smoothing has been used [@wood2008soap].
 
@@
 Among these features, only 160 are available or relevant to the Dublin Area.
+Urban features are obtained using the R package osmdata [@R-osmdata] which queries the Overpass API [@wiki2010overpass].
 The distance between each property and the closest point corresponds to each of the 160 urban features.
@@
 
-In addition to the distance to every urban feature, each house sold in 2018 was related to 48 socio-economic features of the small area including the house, as measured by the Irish census 2016 [@cso2020].
-The results of Irish 2016 census consultation is accessible through both the Central Statistics Office and All-Island Research Observatory.
+In addition to the distance to every urban feature, each property sold in 2018 was related to 48 socio-economic features of the small area including the property, as measured by the Irish census 2016.
+The results of Irish 2016 census consultation is accessible through both the Central Statistics Office [@cso2020] and All-Island Research Observatory [@airo2020].
 The data obtained can be mapped over small area boundaries which are fractions of Irish Electoral Division map.
@@
 
-## Gradient Boosting Regressions
+## eXtreme Gradient Boosting
 
-Lastly, an eXtreme Gradient Boosting (XGBoost) regression model was calculated to examine the extent to which house prices rely on urban and on socio-economic features.
-The XGBoost algorithm optimizes the prediction accuracy by performing iterative least-squares regressions (set to 100 iterations as the best trade-off between accuracy and speed), thereby minimizing the root mean squared error [@friedman2001greedy; @chen2016xgboost].
+Lastly, an eXtreme Gradient Boosting (XGBoost) model was calculated to examine the extent to which property prices rely on urban and on socio-economic features.
+The XGBoost algorithm optimizes the prediction accuracy by performing boosting iterations (set to 100 iterations as the best trade-off between accuracy and speed), thereby minimizing the root mean squared error when the model's objective is a continuous output [@friedman2001greedy].
+The R package "xgboost" (v.1.1.1.1) is used to implement the XGBoost algorithm [@chen2016xgboost].
+With this version, two types of booster are available: tree based and linear regression based.
+Different types of learning task can also be defined according to the distribution of the output.
+In the case of a continuous output, they include regression with squared loss, regression with squared log loss, gamma regression with log-link, and tweedie regression with log-link.
+
 The original dataset has been randomly split to 80% for training and 20% for testing the models.
-Models accuracy is estimated using the coefficient of determination ($R^2$), the Root Mean Square Error ($RMSE$) and the Mean Absolute Error ($MAE$) of the predicted property price values of the test dataset.
+The prediction accuracy is estimated using the coefficient of determination ($R^2$), the Root Mean Square Error ($RMSE$) and the Mean Absolute Error ($MAE$) of the predicted property price values of the test dataset.
+
+Both tree and linear regression based boosting model are used for every of the continuous output learning task in order to compare their accuracy with urban and socio-economic features.
+As no specific weight or bias in the models has been hypothesized, the parameters of the XGBoost function have been set to the default value.
+Then, the most accurate model is used to identify the features that are contributing the most to the prediction of property prices.
+This importance is a measure of the improvement in accuracy brought by a feature based on its coefficient's absolute magnitude.
 
@@
 
-In 2018, the average price of a sold property in Dublin was `r scales::dollar(round(mean_dat, 0), suffix = "", prefix = "€")` (SD = `r scales::dollar(round(sd_dat, 0), suffix = "", prefix = "€")`).
+In 2018, the average price of a sold property in Dublin was `r scales::dollar(round(mean_dat, 0), suffix = "", prefix = "€")` with a standard deviation of `r scales::dollar(round(sd_dat, 0), suffix = "", prefix = "€")`.
 In order to remove potential human errors and outliers, prices higher or lower than 1 SD were removed from the original dataset.
 
-```{r distrib-plot, fig.env='figure*', fig.height=3, out.width='0.98\\textwidth', fig.cap="(ref:distrib-plot)"}
-price_density <- data_dublin_geocoded %>%
+```{r distrib-plot, fig.height=3, out.width="0.98\\columnwidth", fig.pos ="!h", fig.cap="(ref:distrib-plot)"}
+data_dublin_geocoded %>%
   dplyr::filter(year == 2018) %>%
@@
   theme(
-    text = element_text(family = "serif", size = 6),
+    text = element_text(family = "serif", size = 12),
     axis.text.y = element_blank(),
@@
     legend.position = "none"
-  ) +
-  ggtitle("A")
+  )
+```
 
-map_density <- gam_pred %>%
+```{r distrib-gps, fig.height=5, out.width="0.98\\columnwidth", fig.pos ="!h", fig.cap="(ref:distrib-gps)"}
+gam_pred %>%
   dplyr::filter(value > 200000) %>%
@@
   theme(
-    text = element_text(family = "serif", size = 6),
+    text = element_text(family = "serif", size = 12),
     legend.position = "none"
-  ) +
-  ggtitle("B")
-
-price_density + map_density
+  )
 ```
 
-The density of housing prices reveals a unimodal distribution of property prices (Figure \@ref(fig:distrib-plot)A).
-In addition, the geographical distribution of property prices indicates a higher valuation in the South-West of the city as well as on the coast line (Figure \@ref(fig:distrib-plot)B).
+The density of housing prices reveals a unimodal distribution of property prices (Figure \@ref(fig:distrib-plot)).
+In addition, the geographical distribution of property prices indicates a higher valuation in the South-West of the city as well as on the coast line (Figure \@ref(fig:distrib-gps)).
 
@@
   as.matrix()
+```
 
+In order to compare the possibilities of the different type of parameters in the XGBoost models,
+all combinations of tree and linear regression based boosters with all possible output objectives (i.e squared error, squared log error, gamma, and tweedie) are displayed Table \@ref(tab:osm-model-comparison).
+
+```{r osm-model-comparison, results="asis"}
+comparison_xgboost <- 
+  tidyr::expand_grid(
+    booster_parameter = c("gbtree", "gblinear"), 
+    objective_parameter = c("reg:squarederror", "reg:squaredlogerror", "reg:gamma", "reg:tweedie")
+  ) %>%
+  dplyr::rowwise() %>%
+  dplyr::summarise(xgboost_all_parameters(booster_parameter, objective_parameter)) %>% 
+  tidyr::pivot_wider(names_from = .metric, values_from = .estimate) %>% 
+  dplyr::mutate(
+    objective = stringr::str_sub(objective, 5),
+    rmse = scales::dollar(rmse, suffix = "", prefix = "€"),
+    mae = scales::dollar(mae, suffix = "", prefix = "€")
+  )
+
+comparison_xgboost %>% 
+  dplyr::select(-booster) %>% 
+  knitr::kable(
+    "latex",
+    caption = "(ref:osm-model-comparison)",
+    booktabs = TRUE,
+    linesep = "",
+    col.names = c("Objective", "$RMSE$", "$R^2$", "$MAE$"),
+    escape = FALSE
+  ) %>%
+  kableExtra::kable_styling(font_size = 8, latex_options = "HOLD_position") %>% 
+  kableExtra::pack_rows(index = c("Tree based" = 4, "Linear regression based" = 4)) %>% 
+  kableExtra::footnote("Missing value ares due to incompatibilities between data and methods.")
+```
+
+The model comparison reveals that tree based boosting models are more accurate than linear regression based model.
+In addition, within the possible objectives of the tree based boosting models, the tweedie regression with log-link appears to obtain the highest $R^2$ value as well as the lowest $RMSE$ and $MAE$ prediction errors.
+
+```{r osm-best-model}
 xgmodel <- xgboost::xgboost(
+  # https://xgboost.readthedocs.io/en/latest/parameter.html --------------------
   data = train_tbl_x,
   label = train_tbl_y,
-  objective = "reg:squarederror",
+  # Parameters for Tree Booster ------------------------------------------------
+  booster = "gbtree",
+  learning_rate = 0.3,
+  min_split_loss = 0,
+  max_depth = 6,
+  min_child_weight = 1,
+  max_delta_step = 0,
+  subsample = 1,
+  sampling_method = "uniform",
+  colsample_bytree = 1,
+  colsample_bylevel = 1,
+  colsample_bynode = 1,
+  reg_lambda = 1,
+  reg_alpha = 0,
+  tree_method = "auto", # Choices: auto, exact, approx, hist, gpu_hist
+  # scale_pos_weight = 1,
+  refresh_leaf = 1,
+  process_type = "default", # Choices: default, update
+  grow_policy = "depthwise", # Choices: depthwise, lossguide
+  max_leaves = 0,
+  # Parameters for Linear Booster ----------------------------------------------
+  # booster = "gblinear",
+  # reg_lambda = 0,
+  # reg_alpha = 0,
+  # updater = "shotgun", # Choices: shotgun, coord_descent
+  # feature_selector = "cyclic", # Choices: cyclic, shuffle
+  # Learning Task Parameters ---------------------------------------------------
+  objective = "reg:tweedie", 
+  # Choices (reg only): squarederror, squaredlogerror, gamma, tweedie
   eval_metric = "rmse",
+  # Choices: rmse, rmsle, mae, mphe, logloss, error, error@t, merror, ...
+  base_score = 0.5,
   verbose = 0,
@@
 
-Using the XGBoost algorithm, the 160 urban features are explaining `r scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"])` of the property price variance (`r lm_result$statistic$modelfit$r2`) with a $RMSE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "rmse"], suffix = "", prefix = "€")` and a $MAE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "mae"], suffix = "", prefix = "€")`.
+Using the XGBoost algorithm with a tree based booster and a tweedie regression with log-link learning objective, the 160 urban features are explaining `r scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"])` of the property price variance (`r lm_result$statistic$modelfit$r2`) with a $RMSE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "rmse"], suffix = "", prefix = "€")` and a $MAE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "mae"], suffix = "", prefix = "€")`.
 
-```{r osm-features-xgb, fig.cap="(ref:osm-features-xgb)", fig.height=8, out.width='0.98\\columnwidth'}
+The overall correlation of the predicted property prices with urban features is shown in Figure \@ref(fig:osm-features-xgb)A.
+It can be noticed that property prices situated at the low and the high end of the range are the most difficult to predict (Figure \@ref(fig:osm-features-xgb)B).
+A possible explanation is the absence of distinctive and recurrent patterns in urban features for these properties.
+However, the prices higher than €300,000 are potentially driving down the prediction accuracy due to outliers.
+
+```{r osm-features-xgb, fig.cap="(ref:osm-features-xgb)", fig.height=7, fig.pos ="!h", out.width="0.98\\columnwidth"}
 result_accuracy <- result %>%
@@
   theme(
-    text = element_text(family = "serif"),
+    text = element_text(family = "serif", size = 12),
     legend.position = "none"
@@
   theme(
-    text = element_text(family = "serif")
+    text = element_text(family = "serif", size = 12)
   ) +
@@
 
-The overall correlation of the predicted property prices with urban features is shown in Figure \@ref(fig:osm-features-xgb)A.
-It can be noticed that property prices situated at the low and the high end of the range are the most difficult to predict (Figure \@ref(fig:osm-features-xgb)B).
-A possible explanation is the absence of distinctive and recurrent patterns in urban features for these houses.
-However, the prices higher than €300,000 are potentially driving down the prediction accuracy due to outliers.
-
-```{r osm-features-table, results="asis"}
-table_osm <- xgb.importance(model = xgmodel) %>%
-  dplyr::filter(Gain > 0.01) %>%
+```{r features-table}
+table_osm <- xgboost::xgb.importance(model = xgmodel) %>%
+  dplyr::slice_head(n = 10) %>% 
+  #dplyr::filter(Gain > 0.01) %>%
   dplyr::arrange(desc(Gain)) %>%
@@
   tidyr::separate(Feature, c("Feature Category", "Feature Type"), sep = "_")
+```
 
+By analysing their importance (Table \@ref(tab:osm-features-table)), the most relevant geographical features to predict housing prices are the distance to an embassy (`r table_osm$Importance[1]`) and the distance to natural grasslands such as parks and gardens (`r table_osm$Importance[2]`).
+
+```{r osm-features-table, results="asis"}
 table_osm %>% 
@@
   ) %>%
-  kableExtra::kable_styling(font_size = 8, full_width = TRUE)
+  kableExtra::kable_styling(
+    font_size = 8, 
+    full_width = TRUE, 
+    latex_options = "HOLD_position"
+  )
 ```
-
-By analysing their importance (Table \@ref(tab:osm-features-table)), the most relevant geographical features to predict housing prices are the distance to an embassy (`r table_osm$Importance[1]`) and the distance to natural grasslands such as parks and gardens (`r table_osm$Importance[2]`).
 
@@
 ```{r census-features}
-# census dublin features
-source(here::here("script/census_dublin_features.R"))
-
 train_tbl_y <- train_dublin_2018_features %>%
@@
   as.matrix()
+```
 
+Similarly to urban features, the accuracy of different XGBoost models using socio-economic features to predict property prices are displayed Table \@ref(tab:census-model-comparison).
+
+```{r census-model-comparison, results="asis"}
+comparison_xgboost <- 
+  tidyr::expand_grid(
+    booster_parameter = c("gbtree", "gblinear"), 
+    objective_parameter = c("reg:squarederror", "reg:squaredlogerror", "reg:gamma", "reg:tweedie")
+  ) %>%
+  dplyr::rowwise() %>%
+  dplyr::summarise(xgboost_all_parameters(booster_parameter, objective_parameter)) %>% 
+  tidyr::pivot_wider(names_from = .metric, values_from = .estimate) %>% 
+  dplyr::mutate(
+    objective = stringr::str_sub(objective, 5),
+    rmse = scales::dollar(rmse, suffix = "", prefix = "€"),
+    mae = scales::dollar(mae, suffix = "", prefix = "€")
+  )
+
+comparison_xgboost %>% 
+  dplyr::select(-booster) %>% 
+  knitr::kable(
+    "latex",
+    caption = "(ref:census-model-comparison)",
+    booktabs = T,
+    linesep = "",
+    col.names = c("Objective", "$RMSE$", "$R^2$", "$MAE$"),
+    escape = FALSE
+  ) %>%
+  kableExtra::kable_styling(font_size = 8, latex_options = "HOLD_position") %>% 
+  pack_rows(index = c("Tree based" = 4, "Linear regression based" = 4)) %>% 
+  kableExtra::footnote("Missing value ares due to incompatibilities between data and methods.")
+```
+
+The comparison of accuracy metrics reveals that tree based boosting models obtain better results than linear regression based boosting models.
+Within these tree based boosting models, the gamma regression with log-link is the best model as it obtains the highest $R^2$ and lowest $RMSE$ values.
+
+```{r census-best-model}
 xgmodel <- xgboost::xgboost(
+  # https://xgboost.readthedocs.io/en/latest/parameter.html --------------------
   data = train_tbl_x,
   label = train_tbl_y,
-  objective = "reg:squarederror",
+  # Parameters for Tree Booster ------------------------------------------------
+  booster = "gbtree",
+  learning_rate = 0.3,
+  min_split_loss = 0,
+  max_depth = 6,
+  min_child_weight = 1,
+  max_delta_step = 0,
+  subsample = 1,
+  sampling_method = "uniform",
+  colsample_bytree = 1,
+  colsample_bylevel = 1,
+  colsample_bynode = 1,
+  reg_lambda = 1,
+  reg_alpha = 0,
+  tree_method = "auto", # Choices: auto, exact, approx, hist, gpu_hist
+  # scale_pos_weight = 1,
+  refresh_leaf = 1,
+  process_type = "default", # Choices: default, update
+  grow_policy = "depthwise", # Choices: depthwise, lossguide
+  max_leaves = 0,
+  # Parameters for Linear Booster ----------------------------------------------
+  # booster = "gblinear",
+  # reg_lambda = 0,
+  # reg_alpha = 0,
+  # updater = "shotgun", # Choices: shotgun, coord_descent
+  # feature_selector = "cyclic", # Choices: cyclic, shuffle
+  # Learning Task Parameters ---------------------------------------------------
+  objective = "reg:gamma", 
+  # Choices (reg only): squarederror, squaredlogerror, gamma, tweedie
   eval_metric = "rmse",
+  # Choices: rmse, rmsle, mae, mphe, logloss, error, error@t, merror, ...
+  base_score = 0.5,
   verbose = 0,
@@
 
-Using the XGBoost algorithm, the 48 socio-economic features explain `r scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"])` of the property price variance (`r lm_result$statistic$modelfit$r2`) with a $RMSE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "rmse"], suffix = "", prefix = "€")` and a $MAE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "mae"], suffix = "", prefix = "€")`.
-
-```{r census-features-xgb, fig.cap="(ref:census-features-xgb)", fig.height=8, out.width='0.98\\columnwidth'}
+```{r census-features-xgb, fig.cap="(ref:census-features-xgb)", fig.height=7, fig.pos ="!h", out.width="0.98\\columnwidth"}
 result_accuracy <- result %>%
@@
   theme_apa() +
-  theme(text = element_text(family = "serif")) +
+  theme(text = element_text(family = "serif", size = 12)) +
   ggtitle("B")
@@
 
+Using the XGBoost algorithm with a tree based booster and a gamma regression with log-link learning objective, the 48 socio-economic features explain `r scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"])` of the property price variance (`r lm_result$statistic$modelfit$r2`) with a $RMSE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "rmse"], suffix = "", prefix = "€")` and a $MAE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "mae"], suffix = "", prefix = "€")`.
+
 The overall correlation of the predicted property prices with socio-economic features is shown in Figure \@ref(fig:census-features-xgb)A.
@@
 
-```{r census-features-table, results="asis"}
+```{r}
 table_census <- xgb.importance(model = xgmodel) %>%
-  dplyr::filter(Gain > 0.01) %>%
+  dplyr::slice_head(n = 10) %>% 
+  #dplyr::filter(Gain > 0.01) %>%
   dplyr::arrange(desc(Gain)) %>%
@@
   )
+```
 
+According to the analysis of the relative importance of socio-economic features (Table \@ref(tab:census-features-table)), the most important socio-economic features is the proportion individual not providing regular unpaid personal help for a friend or family member with a long-term illness, health problem or disability (`r table_census$Importance[1]`).
+It also appears that the proportion of large properties (i.e., houses with eight or more rooms) in the small area containing the property (`r table_census$Importance[2]`) as well as the proportion of people reporting that they have no religion (`r table_census$Importance[3]`), influence the model.
+
+```{r census-features-table, results="asis"}
 table_census %>%
@@
   ) %>%
-  kableExtra::kable_styling(font_size = 8)
+  kableExtra::kable_styling(
+    font_size = 8, 
+    full_width = TRUE, 
+    latex_options = "HOLD_position"
+  )
 ```
-
-According to the analysis of the relative importance of socio-economic features (Table \@ref(tab:census-features-table)), the most important socio-economic features are the proportion of large houses (i.e., houses with eight or more rooms) in the small area containing the property (`r table_census$Importance[1]`).
-It also appears that areas having a high proportion of young children (`r table_census$Importance[3]`), as well as the proportion of people reporting that they have no religion (`r table_census$Importance[2]`), influence the model.
 
@@
 The main reason is the structure of the market, mainly made of residential properties and few apartment buildings.
-In this study, the prices of 10387 properties sold in 2018 in the area of Dublin were analysed from the Property Services Regulatory Authority database in order to identify the potential factors correlating with these prices.
-The enforcement of a housing price registry has been shown as being an essential tool to regulate the housing market [@tomson2016property].
+In this study, the prices of `r nrow(data_dublin_2018_features)` properties sold in 2018 in the area of Dublin were analysed from the Property Services Regulatory Authority dataset in order to identify the potential factors correlating with these prices.
+The enforcement of a housing price registry has been shown as being an essential tool to regulate the housing market [@kitchin2013making].
 However, a major limitation of this study is the absence of characteristics for the property itself.
 Indeed, prices are mainly determined by factors such as the size of the property or the number of rooms.
+In addition, while some features appear to be more important than others in their contribution to property prices, these results are not causal.
+Thus, features' importance can be either the cause or the consequence of property prices in a neighborhood.
 
 Despite the absence of property characteristics, urban and socio-economic features were used to identify spatial correlates to the housing prices.
-Results revealed that proximity to embassies and grasslands is a driver of house prices.
+Results revealed that proximity to embassies and grasslands is a driver of property prices.
 Embassies are in general located in the most expensive areas of cities which are also the most aesthetically pleasing and the most secure part of cities.
 In addition, the density of embassies and properties allocated to embassy staff reduces the density of property available on the market and drives their prices up.
-The distance to a park or a natural area is also a very important factor for house prices.
-Results also revealed that the density of large houses in the area and the proportion of individuals reporting having no religion are very important.
-Again, the size of the property sold was not included in the database; however, the higher the density of large houses in the area, the higher the probability for the property to be a large house.
-The second most important feature is the density of young children.
-While the literature is not conclusive of the link between wealth and family size, future research could investigate the mediation effect of house size between family size and property price.
+The distance to a park or a natural area is also a very important factor for property prices.
+Results also revealed that the density of individual not providing regular unpaid personal help in the area is the most important socio-economic feature to predict property prices.
+This feature is related to the multiple factors such as the average age of the household or the economic status of the household as they are related to general health [@smith1999healthy].
+Large properties in the area and the proportion of individuals reporting having no religion are also relatively important.
+Again, the size of the property sold was not included in the dataset; however, the higher the density of large properties in the area, the higher the probability for the property to be a large property.
 Finally, the relation with the expression of religious belief, or more precisely its absence, is a very interesting feature.
@@
 Whereas some studies reveal a positive correlation between religion and income [@guiso2003people; @elgin2013religion] others reveals no evidence of such relationship [@de2008does].
-Here, it appears that prices of houses are higher in areas with a high density of people indicating that they have no religion.
+Here, it appears that prices of properties are higher in areas with a high density of people indicating that they have no religion.
 
@@
 The evolution of housing prices is a pressing issue in most European capitals and specially in Dublin.
-Given their significant increase, houses are less and less affordable for individuals.
-By preforming a feature analysis with urban and socio-economic features, it is possible to evaluate and predict the potential price of a house.
-Indeed features such as the presence of embassies or parks are criteria that influence significantly the price of houses.
+Given their significant increase, properties are less and less affordable for individuals.
+By preforming a feature analysis with urban and socio-economic features, it is possible to evaluate and predict the potential price of a property.
+Indeed features such as the presence of embassies or parks are criteria that influence significantly the price of properties.
 Similarly, the characteristics of inhabitants in the area such as religion, health and age is correlated to the evolution of housing prices.
-These results allow an understanding of why some areas have higher prices than others.
+These results allow an understanding of why some areas have higher prices than others which is a relevant information not only for real estate agents in charge of property valuations but also to buyers in order to approximate the real value of a property.
 
@@
 
-The authors would like to thank the developers of the following R packages used to process, analyse, display and report data: `r papaja::cite_r(file = "r-references.bib")`.
+The authors would like to thank the developers of the following R packages used to process, analyse, display and report data: `r papaja::cite_r(file = "bib/r-references.bib")`.
 

### 2020-08-18 manuscript/manuscript.tex
reviewer changes

--- 2020-08-06 dsaa_application_manuscript/dsaa_application_manuscript.tex
@@
 \hypersetup{
-            pdftitle={Urban and socio-economic correlates of property price evolution:~Application to Dublin's costal area},
+            pdftitle={Urban and socio-economic correlates of property price in Dublin's area},
             pdfkeywords={Property price, Housing Market, Feature Analysis, Machine Learning, Geocoding},
@@
 \usepackage{graphicx}
+\usepackage{flushend}
 \usepackage{booktabs}
@@
 % Do not put math or special symbols in the title.
-\title{Urban and socio-economic correlates of property price evolution:~Application to Dublin's costal area}
+\title{Urban and socio-economic correlates of property price in Dublin's area}
 
@@
 However, it is difficult to evaluate the influence of such external factors on property valuation (Clapp, Kim, and Gelfand 2002).
-By using machine learning algorithms, prediction of house prices has become more and more accurate.
+By using machine learning algorithms, prediction of property prices has become more and more accurate.
 Techniques used usually include artificial neural networks due to the volatility of the market (Limsombunchai 2004; Feng and Jones 2015; Selim 2009).
+Indeed, artificial neural network techniques can quickly adjust to volatility changes, therefore incorporating sharp increases or decreases in their predictions (Aragonés, Blanco, and Estévez 2007).
+However, the model obtained from most artificial neural networks are hardly interpretable (Rudin 2019).
 
 In this paper, an alternative machine learning algorithm is used to identify urban and socio-economic correlates of property price evolution in the area of Dublin, Ireland in 2018.
-Due to the high number of potential features, an eXtreme Gradient Boosting (XGBoost) regression model is believed to obtain the best results with low parameter customisation and high speed.
+Due to the high number of potential features, an eXtreme Gradient Boosting (XGBoost) regression model is believed to obtain the best results with low parameter customisation and high speed (Chen and Guestrin 2016).
+Contrary to most artificial neural network techniques, XGBoost is an interpretation-focused method which automatically provides estimates of feature importance from a trained predictive model.
 
@@
 Since 2010, under the Irish Property Services (Regulation) Act, all individuals acquiring a property in Ireland have to declare it to Property Services Regulatory Authority (PSRA).
-They must provide details such as the date of sale, the price and the address of all residential properties purchased in Ireland as declared to the Revenue Commissioners for stamp duty purposes (Authority 2020).
+They must provide details such as the date of sale, the price and the address of all residential properties purchased in Ireland as declared to the Revenue Commissioners for stamp duty purposes (Property Services Regulatory Authority 2020).
 Data is filed electronically by persons doing the conveyancing of the property on behalf of the purchaser, and it must be noted that errors may occur when the data is being filed.
 
-Because of the evolution of urban planning and the evolution of socio-economic features measured by the 2016 census, only houses sold in 2018 were processed to avoid potential temporal incoherences.
-In 2018, a total of 10395 property in the area of Dublin were sold, and these constitute the database used to identify the urban and socio-economic correlates of property prices.
+Because of the evolution of urban planning and the evolution of socio-economic features measured by the 2016 census, only properties sold in 2018 were processed to avoid potential temporal incoherence.
+The total amount of property sold in 2018 in the county Dublin which appears in the Property Price Register is 18665.
+From these, all property corresponding to apartments were removed (10.3\% of the original dataset).
+Finally, unsuccessfully geocoded properties were also removed from the original dataset (38.0\%) to obtain the 10389 of the final dataset used to identify the urban and socio-economic correlates of property prices.
 
-In order to evaluate the spacial distribution of the property sold, property addresses were geocoded (i.e., converted to latitude and longitude) using the OpenStreetMap API.
+In order to evaluate the spacial distribution of the property sold, property addresses were geocoded (i.e., converted to latitude and longitude) using the Nominatim API (Hoffmann 2020) to OpenStreetMap data.
 OpenStreetMap is a collaborative project which aims to create and provide access to free editable maps of the world (Haklay and Weber 2008).
@@
 While these methods are efficient in a non-restricted space, they have limitations when they are used in coastal areas.
-To deal with the influence of boundaries on price estimation, a generalised additive model using soap film smoothing has been used (Wood, Bravington, and Hedley 2008).
+This problem is known as ``finite area smoothing'' and occurs when predictions from a model are approximated across geographical barriers, such as irregular coastal shapes, which would lead to poorly predicted values (Ramsay 2002).
+To deal with the influence of geographical boundaries on price estimation, a generalised additive model using soap film smoothing has been used (Wood, Bravington, and Hedley 2008).
 
@@
 
-OpenStreetMap combines information about more than 400 urban features including road information and building information to categorize features such as amenities, leisure or tourism structures (OpenStreetMap 2020).
+OpenStreetMap combines information about more than 400 urban features including road information and building information to categorize features such as amenities, leisure or tourism structures (OpenStreetMap 2020a).
 Among these features, only 160 are available or relevant to the Dublin Area.
+Urban features are obtained using the R package osmdata (Padgham et al.
+2017) which queries the Overpass API (OpenStreetMap 2020b).
 The distance between each property and the closest point corresponds to each of the 160 urban features.
@@
 
-In addition to the distance to every urban feature, each house sold in 2018 was related to 48 socio-economic features of the small area including the house, as measured by the Irish census 2016 (Office 2020).
-The results of Irish 2016 census consultation is accessible through both the Central Statistics Office and All-Island Research Observatory.
+In addition to the distance to every urban feature, each property sold in 2018 was related to 48 socio-economic features of the small area including the property, as measured by the Irish census 2016.
+The results of Irish 2016 census consultation is accessible through both the Central Statistics Office (Central Statistics Office 2020) and All-Island Research Observatory (All-Island Research Observatory 2020).
 The data obtained can be mapped over small area boundaries which are fractions of Irish Electoral Division map.
@@
 
-\hypertarget{gradient-boosting-regressions}{%
-\subsection{Gradient Boosting Regressions}\label{gradient-boosting-regressions}}
+\hypertarget{extreme-gradient-boosting}{%
+\subsection{eXtreme Gradient Boosting}\label{extreme-gradient-boosting}}
 
-Lastly, an eXtreme Gradient Boosting (XGBoost) regression model was calculated to examine the extent to which house prices rely on urban and on socio-economic features.
-The XGBoost algorithm optimizes the prediction accuracy by performing iterative least-squares regressions (set to 100 iterations as the best trade-off between accuracy and speed), thereby minimizing the root mean squared error (Friedman 2001; Chen and Guestrin 2016).
+Lastly, an eXtreme Gradient Boosting (XGBoost) model was calculated to examine the extent to which property prices rely on urban and on socio-economic features.
+The XGBoost algorithm optimizes the prediction accuracy by performing boosting iterations (set to 100 iterations as the best trade-off between accuracy and speed), thereby minimizing the root mean squared error when the model's objective is a continuous output (Friedman 2001).
+The R package ``xgboost'' (v.1.1.1.1) is used to implement the XGBoost algorithm (Chen and Guestrin 2016).
+With this version, two types of booster are available: tree based and linear regression based.
+Different types of learning task can also be defined according to the distribution of the output.
+In the case of a continuous output, they include regression with squared loss, regression with squared log loss, gamma regression with log-link, and tweedie regression with log-link.
+
 The original dataset has been randomly split to 80\% for training and 20\% for testing the models.
-Models accuracy is estimated using the coefficient of determination (\(R^2\)), the Root Mean Square Error (\(RMSE\)) and the Mean Absolute Error (\(MAE\)) of the predicted property price values of the test dataset.
+The prediction accuracy is estimated using the coefficient of determination (\(R^2\)), the Root Mean Square Error (\(RMSE\)) and the Mean Absolute Error (\(MAE\)) of the predicted property price values of the test dataset.
+
+Both tree and linear regression based boosting model are used for every of the continuous output learning task in order to compare their accuracy with urban and socio-economic features.
+As no specific weight or bias in the models has been hypothesized, the parameters of the XGBoost function have been set to the default value.
+Then, the most accurate model is used to identify the features that are contributing the most to the prediction of property prices.
+This importance is a measure of the improvement in accuracy brought by a feature based on its coefficient's absolute magnitude.
 
@@
 
-In 2018, the average price of a sold property in Dublin was €330,364 (SD = €180,448).
+In 2018, the average price of a sold property in Dublin was €384,892 with a standard deviation of €175,476.
 In order to remove potential human errors and outliers, prices higher or lower than 1 SD were removed from the original dataset.
 
-\begin{figure*}
-\includegraphics[width=0.98\textwidth]{dsaa_application_manuscript_files/figure-latex/distrib-plot-1} \caption{Overall (A) and geographic (B) distribution of the houses prices density in the city of Dublin in 2018.
-The geographic distribution was obtained by using a generalized additive model with soap film smooth parameter to take into account the influence of coastal boundaries.
-Geographic estimates outside of the 95\% Confidence Interval were removed.}\label{fig:distrib-plot}
-\end{figure*}
+\begin{figure}[!h]
+\includegraphics[width=0.98\columnwidth]{manuscript_files/figure-latex/distrib-plot-1} \caption{Overall distribution of the property prices density in the city of Dublin in 2018}\label{fig:distrib-plot}
+\end{figure}
 
-The density of housing prices reveals a unimodal distribution of property prices (Figure \ref{fig:distrib-plot}A).
-In addition, the geographical distribution of property prices indicates a higher valuation in the South-West of the city as well as on the coast line (Figure \ref{fig:distrib-plot}B).
+\begin{figure}[!h]
+\includegraphics[width=0.98\columnwidth]{manuscript_files/figure-latex/distrib-gps-1} \caption{Geographic distribution of the property prices density in the city of Dublin in 2018 using a generalized additive model with soap film smooth parameter to take into account the influence of coastal boundaries.
+Estimates outside of the 95\% Confidence Interval were removed.}\label{fig:distrib-gps}
+\end{figure}
+
+The density of housing prices reveals a unimodal distribution of property prices (Figure \ref{fig:distrib-plot}).
+In addition, the geographical distribution of property prices indicates a higher valuation in the South-West of the city as well as on the coast line (Figure \ref{fig:distrib-gps}).
 
@@
 
-Using the XGBoost algorithm, the 160 urban features are explaining 44\% of the property price variance (\(F(1, 2076) = 1,629.56\), \(p < .001\)) with a \(RMSE\) of €127,783 and a \(MAE\) of €85,115.25.
+In order to compare the possibilities of the different type of parameters in the XGBoost models,
+all combinations of tree and linear regression based boosters with all possible output objectives (i.e squared error, squared log error, gamma, and tweedie) are displayed Table \ref{tab:osm-model-comparison}.
 
-\begin{figure}
-\includegraphics[width=0.98\columnwidth]{dsaa_application_manuscript_files/figure-latex/osm-features-xgb-1} \caption{Property price prediction accuracy (A) and Property price prediction error (B) using urban features with XGBoost.}\label{fig:osm-features-xgb}
-\end{figure}
+\begin{table}[H]
+
+\caption{\label{tab:osm-model-comparison}Accuracy of XGBoost models using urban features according the type of gradient boosting method and output objective.}
+\centering
+\fontsize{8}{10}\selectfont
+\begin{tabular}[t]{llrl}
+\toprule
+Objective & $RMSE$ & $R^2$ & $MAE$\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Tree based}}\\
+\hspace{1em}squarederror & €127,783 & 0.44 & €85,115\\
+\hspace{1em}squaredlogerror & €426,989 & - & €391,670\\
+\hspace{1em}gamma & €129,052 & 0.43 & €87,039\\
+\hspace{1em}tweedie & €126,626 & 0.45 & €85,073\\
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Linear regression based}}\\
+\hspace{1em}squarederror & €150,833 & 0.21 & €111,788\\
+\hspace{1em}squaredlogerror & €406,947 & 0.00 & €369,713\\
+\hspace{1em}gamma & €180,313 & 0.06 & €136,059\\
+\hspace{1em}tweedie & €172,903 & 0.08 & €130,148\\
+\bottomrule
+\multicolumn{4}{l}{\textit{Note: }}\\
+\multicolumn{4}{l}{Missing value ares due to incompatibilities between data and methods.}\\
+\end{tabular}
+\end{table}
+
+The model comparison reveals that tree based boosting models are more accurate than linear regression based model.
+In addition, within the possible objectives of the tree based boosting models, the tweedie regression with log-link appears to obtain the highest \(R^2\) value as well as the lowest \(RMSE\) and \(MAE\) prediction errors.
+
+Using the XGBoost algorithm with a tree based booster and a tweedie regression with log-link learning objective, the 160 urban features are explaining 45\% of the property price variance (\(F(1, 2076) = 1,690.80\), \(p < .001\)) with a \(RMSE\) of €126,626 and a \(MAE\) of €85,073.07.
 
@@
 It can be noticed that property prices situated at the low and the high end of the range are the most difficult to predict (Figure \ref{fig:osm-features-xgb}B).
-A possible explanation is the absence of distinctive and recurrent patterns in urban features for these houses.
+A possible explanation is the absence of distinctive and recurrent patterns in urban features for these properties.
 However, the prices higher than €300,000 are potentially driving down the prediction accuracy due to outliers.
 
-\begin{table}
+\begin{figure}[!h]
+\includegraphics[width=0.98\columnwidth]{manuscript_files/figure-latex/osm-features-xgb-1} \caption{Property price prediction accuracy (A) and prediction error (B) using urban features with XGBoost.}\label{fig:osm-features-xgb}
+\end{figure}
 
-\caption{\label{tab:osm-features-table}Urban features importance (higher than 1\%).}
+By analysing their importance (Table \ref{tab:osm-features-table}), the most relevant geographical features to predict housing prices are the distance to an embassy (15.6\%) and the distance to natural grasslands such as parks and gardens (4.8\%).
+
+\begin{table}[H]
+
+\caption{\label{tab:osm-features-table}Top 10 most important urban features contributing to the property price prediction.}
 \centering
@@
 \midrule
-amenity & embassy & 17.7\%\\
-natural & grassland & 6.7\%\\
-route & bus & 2.3\%\\
-power & line & 2.2\%\\
+amenity & embassy & 15.6\%\\
+natural & grassland & 4.8\%\\
+power & line & 2.4\%\\
+route & bus & 2.4\%\\
+amenity & bar & 2.0\%\\
+boundary & political & 1.9\%\\
+place & island & 1.6\%\\
 boundary & administrative & 1.5\%\\
-place & island & 1.5\%\\
-barrier & wall & 1.5\%\\
-cycleway & share & 1.4\%\\
-route & ferry & 1.3\%\\
-cutting & yes & 1.1\%\\
-amenity & bar & 1.1\%\\
-barrier & yes & 1.1\%\\
-place & locality & 1.1\%\\
-boundary & political & 1.1\%\\
-route & road & 1.1\%\\
-highway & path & 1.1\%\\
-tunnel & yes & 1.0\%\\
-boundary & postal & 1.0\%\\
+highway & residential & 1.4\%\\
+barrier & wall & 1.4\%\\
 \bottomrule
@@
 
-By analysing their importance (Table \ref{tab:osm-features-table}), the most relevant geographical features to predict housing prices are the distance to an embassy (17.7\%) and the distance to natural grasslands such as parks and gardens (6.7\%).
-
 \hypertarget{correlates-with-socio-economic-features}{%
@@
 
-Using the XGBoost algorithm, the 48 socio-economic features explain 42\% of the property price variance (\(F(1, 2076) = 1,500.87\), \(p < .001\)) with a \(RMSE\) of €131,270 and a \(MAE\) of €86,891.85.
+Similarly to urban features, the accuracy of different XGBoost models using socio-economic features to predict property prices are displayed Table \ref{tab:census-model-comparison}.
 
-\begin{figure}
-\includegraphics[width=0.98\columnwidth]{dsaa_application_manuscript_files/figure-latex/census-features-xgb-1} \caption{Property price prediction accuracy (A) and Property price prediction error (B) using socio-economic features with XGBoost.}\label{fig:census-features-xgb}
+\begin{table}[H]
+
+\caption{\label{tab:census-model-comparison}Accuracy of XGBoost models using socio-economic features according the type of gradient boosting method and output objective.}
+\centering
+\fontsize{8}{10}\selectfont
+\begin{tabular}[t]{llrl}
+\toprule
+Objective & $RMSE$ & $R^2$ & $MAE$\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Tree based}}\\
+\hspace{1em}squarederror & €131,270 & 0.42 & €86,892\\
+\hspace{1em}squaredlogerror & €426,989 & - & €391,670\\
+\hspace{1em}gamma & €129,221 & 0.43 & €86,949\\
+\hspace{1em}tweedie & €130,927 & 0.42 & €86,707\\
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Linear regression based}}\\
+\hspace{1em}squarederror & €146,781 & 0.26 & €107,890\\
+\hspace{1em}squaredlogerror & €406,108 & 0.01 & €368,877\\
+\hspace{1em}gamma & €182,432 & 0.08 & €136,441\\
+\hspace{1em}tweedie & €164,439 & 0.13 & €122,259\\
+\bottomrule
+\multicolumn{4}{l}{\textit{Note: }}\\
+\multicolumn{4}{l}{Missing value ares due to incompatibilities between data and methods.}\\
+\end{tabular}
+\end{table}
+
+The comparison of accuracy metrics reveals that tree based boosting models obtain better results than linear regression based boosting models.
+Within these tree based boosting models, the gamma regression with log-link is the best model as it obtains the highest \(R^2\) and lowest \(RMSE\) values.
+
+\begin{figure}[!h]
+\includegraphics[width=0.98\columnwidth]{manuscript_files/figure-latex/census-features-xgb-1} \caption{Property price prediction accuracy (A) and prediction error (B) using socio-economic features with XGBoost.}\label{fig:census-features-xgb}
 \end{figure}
+
+Using the XGBoost algorithm with a tree based booster and a gamma regression with log-link learning objective, the 48 socio-economic features explain 43\% of the property price variance (\(F(1, 2076) = 1,554.37\), \(p < .001\)) with a \(RMSE\) of €129,221 and a \(MAE\) of €86,949.15.
 
@@
 
-\begin{table}
+According to the analysis of the relative importance of socio-economic features (Table \ref{tab:census-features-table}), the most important socio-economic features is the proportion individual not providing regular unpaid personal help for a friend or family member with a long-term illness, health problem or disability (32.8\%).
+It also appears that the proportion of large properties (i.e., houses with eight or more rooms) in the small area containing the property (19.1\%) as well as the proportion of people reporting that they have no religion (4.2\%), influence the model.
 
-\caption{\label{tab:census-features-table}Socio-economic features importance (higher than 1\%).}
+\begin{table}[H]
+
+\caption{\label{tab:census-features-table}Top 10 most important socio-economic features contributing to the property price prediction.}
 \centering
 \fontsize{8}{10}\selectfont
-\begin{tabular}[t]{lll}
+\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X}
 \toprule
@@
 \midrule
-housing rooms & 8 or more Rooms & 28.2\%\\
-population & Age 0 - 14 & 4.5\%\\
-religion & No Religion & 4.4\%\\
-population & Age 80 Plus & 3.6\%\\
-general health & Very Good & 3.5\%\\
-housing rooms & 6 Rooms & 3.2\%\\
-housing tenure & Owner Occupier with Mortgage & 2.5\%\\
-disabilty age group & Persons with a disability aged 65 Plus & 2.3\%\\
-housing rooms & 5 Rooms & 2.2\%\\
-housing tenure & Social Rented & 2.2\%\\
-housing rooms & 4 Rooms & 2.1\%\\
-religion & Other Catholic & 2.0\%\\
-housing rooms & 7 Rooms & 1.9\%\\
-housing rooms & 3 Rooms & 1.8\%\\
-religion & No Answer & 1.8\%\\
-housing tenure & Private Rented & 1.6\%\\
-population & Age 45 - 64 & 1.5\%\\
-population & Age 25 - 44 & 1.5\%\\
-housing tenure & Owner Occupier No Mortgage & 1.5\%\\
-disabilty age group & Persons with a disability aged 0 - 14 & 1.5\%\\
-disabilty age group & Persons with a disability aged 25 - 44 & 1.4\%\\
-housing rooms & 1 Room & 1.4\%\\
-carers & Provides No Care & 1.3\%\\
-religion & Roman Catholic & 1.3\%\\
-general health & Good & 1.3\%\\
-population & Female & 1.3\%\\
-general health & Fair & 1.3\%\\
-disabilty age group & Total persons with a disability & 1.2\%\\
-general health & Bad & 1.2\%\\
-religion & Other Religion & 1.2\%\\
-disabilty age group & Persons with a disability aged 45 - 64 & 1.2\%\\
-carers & Total Care Providers & 1.2\%\\
-population & Age 15 - 24 & 1.1\%\\
-carers & 20-49 hours unpaid PW & 1.1\%\\
-housing tenure & Rented Free of Rent & 1.0\%\\
+carers & Provides No Care & 32.8\%\\
+housing rooms & 8 or more Rooms & 19.1\%\\
+religion & No Religion & 4.2\%\\
+population & Age 0 - 14 & 3.2\%\\
+housing rooms & 6 Rooms & 2.1\%\\
+carers & 1-19 hours unpaid PW & 1.9\%\\
+housing tenure & Social Rented & 1.8\%\\
+general health & Very Good & 1.7\%\\
+religion & Other Catholic & 1.6\%\\
+general health & Good & 1.5\%\\
 \bottomrule
-\end{tabular}
+\end{tabu}
 \end{table}
-
-According to the analysis of the relative importance of socio-economic features (Table \ref{tab:census-features-table}), the most important socio-economic features are the proportion of large houses (i.e., houses with eight or more rooms) in the small area containing the property (28.2\%).
-It also appears that areas having a high proportion of young children (4.4\%), as well as the proportion of people reporting that they have no religion (4.5\%), influence the model.
 
@@
 The main reason is the structure of the market, mainly made of residential properties and few apartment buildings.
-In this study, the prices of 10387 properties sold in 2018 in the area of Dublin were analysed from the Property Services Regulatory Authority database in order to identify the potential factors correlating with these prices.
-The enforcement of a housing price registry has been shown as being an essential tool to regulate the housing market (Tomson 2016).
+In this study, the prices of 10387 properties sold in 2018 in the area of Dublin were analysed from the Property Services Regulatory Authority dataset in order to identify the potential factors correlating with these prices.
+The enforcement of a housing price registry has been shown as being an essential tool to regulate the housing market (Kitchin 2013).
 However, a major limitation of this study is the absence of characteristics for the property itself.
 Indeed, prices are mainly determined by factors such as the size of the property or the number of rooms.
+In addition, while some features appear to be more important than others in their contribution to property prices, these results are not causal.
+Thus, features' importance can be either the cause or the consequence of property prices in a neighborhood.
 
 Despite the absence of property characteristics, urban and socio-economic features were used to identify spatial correlates to the housing prices.
-Results revealed that proximity to embassies and grasslands is a driver of house prices.
+Results revealed that proximity to embassies and grasslands is a driver of property prices.
 Embassies are in general located in the most expensive areas of cities which are also the most aesthetically pleasing and the most secure part of cities.
 In addition, the density of embassies and properties allocated to embassy staff reduces the density of property available on the market and drives their prices up.
-The distance to a park or a natural area is also a very important factor for house prices.
-Results also revealed that the density of large houses in the area and the proportion of individuals reporting having no religion are very important.
-Again, the size of the property sold was not included in the database; however, the higher the density of large houses in the area, the higher the probability for the property to be a large house.
-The second most important feature is the density of young children.
-While the literature is not conclusive of the link between wealth and family size, future research could investigate the mediation effect of house size between family size and property price.
+The distance to a park or a natural area is also a very important factor for property prices.
+Results also revealed that the density of individual not providing regular unpaid personal help in the area is the most important socio-economic feature to predict property prices.
+This feature is related to the multiple factors such as the average age of the household or the economic status of the household as they are related to general health (Smith 1999).
+Large properties in the area and the proportion of individuals reporting having no religion are also relatively important.
+Again, the size of the property sold was not included in the dataset; however, the higher the density of large properties in the area, the higher the probability for the property to be a large property.
 Finally, the relation with the expression of religious belief, or more precisely its absence, is a very interesting feature.
@@
 2013) others reveals no evidence of such relationship (De La O and Rodden 2008).
-Here, it appears that prices of houses are higher in areas with a high density of people indicating that they have no religion.
+Here, it appears that prices of properties are higher in areas with a high density of people indicating that they have no religion.
 
@@
 \begin{cslreferences}
+\leavevmode\hypertarget{ref-airo2020}{}%
+All-Island Research Observatory. 2020. ``AIRO Data Store.'' June 1, 2020. \url{http://airo.maynoothuniversity.ie/datastore}.
+
 \leavevmode\hypertarget{ref-antonakakis2016dynamic}{}%
@@
 
+\leavevmode\hypertarget{ref-aragones2007neural}{}%
+Aragonés, José R, Carlos Blanco, and Pablo García Estévez.
+2007.
+``Neural Network Volatility Forecasts.'' \emph{Intelligent Systems in Accounting, Finance \& Management: International Journal} 15 (3-4): 107--21.
+
 \leavevmode\hypertarget{ref-R-papaja}{}%
 Aust, Frederik, and Marius Barth. 2020. \emph{papaja: Create APA Manuscripts with R Markdown}. \url{https://github.com/crsh/papaja}.
-
-\leavevmode\hypertarget{ref-rppr2020}{}%
-Authority, Property Services Regulatory. 2020. ``Residential Property Price Register.'' June 1, 2020. \url{https://propertypriceregister.ie}.
 
@@
 ``Spatial Dependence, Housing Submarkets, and House Price Prediction.'' \emph{The Journal of Real Estate Finance and Economics} 35 (2): 143--60.
+
+\leavevmode\hypertarget{ref-cso2020}{}%
+Central Statistics Office. 2020. ``Census 2016 Small Area Population Statistics.'' June 1, 2020. \url{https://www.cso.ie/en/census/census2016reports/census2016smallareapopulationstatistics/}.
 
@@
 
+\leavevmode\hypertarget{ref-nominatim2020}{}%
+Hoffmann, Sarah. 2020. ``Nominatim: Open-Source Geocoding with Openstreetmap Data.'' June 1, 2020. \url{https://nominatim.org/}.
+
+\leavevmode\hypertarget{ref-kitchin2013making}{}%
+Kitchin, Rob. 2013. ``Making Informed Decisions on Future Housing Policy.'' \emph{Housing Ireland: A Journal for Irish Housing Professionals}, no. 3: 36--41.
+
 \leavevmode\hypertarget{ref-R-yardstick}{}%
@@
 
-\leavevmode\hypertarget{ref-cso2020}{}%
-Office, Central Statistics. 2020. ``Census 2016 Small Area Population Statistics.'' June 1, 2020. \url{http://census.cso.ie/sapmap}.
+\leavevmode\hypertarget{ref-openstreetmap2020}{}%
+OpenStreetMap. 2020a. ``Map Features (Wiki).'' June 1, 2020. \url{https://wiki.openstreetmap.org/wiki/Map_Features}.
 
-\leavevmode\hypertarget{ref-openstreetmap2020}{}%
-OpenStreetMap. 2020. ``Map Features (Wiki).'' June 1, 2020. \url{https://wiki.openstreetmap.org/wiki/Map_Features}.
+\leavevmode\hypertarget{ref-wiki2010overpass}{}%
+---------. 2020b. ``Overpass API.'' June 1, 2020. \url{https://wiki.openstreetmap.org/wiki/Overpass_API}.
 
@@
 
+\leavevmode\hypertarget{ref-rppr2020}{}%
+Property Services Regulatory Authority. 2020. ``Residential Property Price Register.'' June 1, 2020. \url{https://propertypriceregister.ie}.
+
+\leavevmode\hypertarget{ref-ramsay2002spline}{}%
+Ramsay, Tim. 2002. ``Spline Smoothing over Difficult Regions.'' \emph{Journal of the Royal Statistical Society: Series B (Statistical Methodology)} 64 (2): 307--19.
+
 \leavevmode\hypertarget{ref-R-base}{}%
 R Core Team. 2020. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.
+
+\leavevmode\hypertarget{ref-rudin2019stop}{}%
+Rudin, Cynthia. 2019. ``Stop Explaining Black Box Machine Learning Models for High Stakes Decisions and Use Interpretable Models Instead.'' \emph{Nature Machine Intelligence} 1 (5): 206--15.
 
@@
 
+\leavevmode\hypertarget{ref-smith1999healthy}{}%
+Smith, James P. 1999. ``Healthy Bodies and Thick Wallets: The Dual Relation Between Health and Economic Status.'' \emph{Journal of Economic Perspectives} 13 (2): 145--66.
+
 \leavevmode\hypertarget{ref-R-tmap}{}%
 Tennekes, Martijn. 2018. ``tmap: Thematic Maps in R.'' \emph{Journal of Statistical Software} 84 (6): 1--39. \url{https://doi.org/10.18637/jss.v084.i06}.
-
-\leavevmode\hypertarget{ref-tomson2016property}{}%
-Tomson, AIVAR.
-2016.
-``Property Sales Register as a Tool to Improve the Quality of Valuation and Market Research: Lessons Learned from Selected Countries.'' In \emph{World Bank Conference on Land and Poverty, Washington, Dc, March}, 14--18.
 

### 2020-08-18 review/revision_letter.Rmd
reviewer changes

@@
 manuscript  : "Paper 147"
+
+bibliography: "C:/Users/dupred/OneDrive/Projects/DSAA_2020/manuscript/bib/references.bib"
 
@@
 library(here)
+library(janitor)
 ```
@@
 
+\RC{Urban and socio-economic correlates of property price evolution: Application to Dublin’s costal area
+
+Hypothesis : By taking into account either urban or socio-economic features, it is possible to accurately estimate housing prices and to predict their evolution.
+
+Proposal : By using machine learning algorithms, prediction of house prices has become more and more accurate.
+In this paper, an alternative machine learning algorithm is used to identify urban and socio-economic correlates of property price evolution in the area of Dublin, Ireland in 2018.}
+
+Many thanks for your review and for your comments, we will address each of them as indicated in our answers.
+
 \RC{Page 1 : « 10389 houses » in the Abstract and « 10395 properties » in the Method section and « 10387 Properties » in Discussion section ?}
+
+The figures have been corrected with the appropriate number of observation in our data set (10387 properties). In addition the terms house/s has been replaced by property/ies where applicable.
 
@@
 
+No modifications of the current version of the XGBoost R-package algorithm has been applied to obtain our results. The hyperparameters are all defaults as precised in the manuscript:
+
+> As no specific weight or bias in the models has been hypothesized, the parameters of the XGBoost function have been set to the default value.
+
 \RC{I suppose that Urban features importance is computed by the XGBoost model (Table I).
@@
 Embassy importance is detected thanks to statistical computations.
-Is it possible to comment and validate it by qualitative interviews with experts ?}
-
-\RC{Page 4 : Similarly, I couldn’t understand how Socio-Economic features importance is computed (Table II).
+Is it possible to comment and validate it by qualitative interviews with experts ?
+Page 4 : Similarly, I couldn’t understand how Socio-Economic features importance is computed (Table II).
 Could you explain this stage and comment results ?
@@
 
+Indeed the importance is obtained from the XGBoost model.
+It is a measure of the improvement in accuracy brought by a feature based on its coefficient's absolute magnitude.
+Compared to the other features included, the importance a specific feature corresponds to the proportion of its contribution to the overall accuracy of the model.
+This information as been added to the "eXtreme Gradient Boosting" section as follow:
+
+> Then, the most accurate model is used to identify the features that are contributing the most to the prediction of property prices.
+This importance is a measure of the improvement in accuracy brought by a feature based on its coefficient's absolute magnitude.
+
+Our explanations on the importance of certain features to the model are interpretative and, therefore, we would like to keep these interpretations in the discussion section.
+We agree that it would be easier for the reader to find them right after each result section, however, we think that it is important to keep separated objective results and subjective interpretations.
+
+The idea of qualitative interviews is very interesting. However, we would like to keep this submission as a quantitative study. Interviews will be considered in longer manuscripts.
+
 \RC{« Embassies are in general located in the most expensive areas of cities ». Are the ambassies installed in expensive areas or are the areas expensive because of their embassies ?}
+
+The question of correlation vs.
+causation is highly relevant.
+Because linear models are based on correlations, we want to avoid any interpretation related to causation.
+As a consequence, the following sentence has been added to the document discussion:
+
+> In addition, while some features appear to be more important than others in their contribution to property prices, these results are not causal.
+Thus, features' importance can be either the cause or the consequence of property prices in a neighborhood.
 
@@
 - Is it a proposal of a new regression model ? If yes, you should better explain XGBoost and its originality
+
 - Is it a experiment of an existing regression model applied to housing prices ?
-If yes, you should comment the results on the 20% of the original dataset : are the price predictions relevant ?
+If yes, you should comment the results on the $20\%$ of the original dataset : are the price predictions relevant ?
 Is this approach better than existing ones for price prediction ?
@@
 Computation speed ?
-etc.}
+etc.
 
-\RC{What was not possible before and what is possible now thanks to this work ?}
+What was not possible before and what is possible now thanks to this work ?}
+
+As indicated in the title, the added value of the paper is about identifying urban and socio-economic correlates of property price evolution.
+These correlated are obtained on the results of the test dataset.
+The price predictions is relevant to real estate agents in charge of property valuations and to buyers to approximate the real value of a property.
+This added value has been added in the conclusion as follow:
+
+> These results allow an understanding of why some areas have higher prices than others which is a relevant information not only for real estate agents in charge of property valuations but also to buyers in order to approximate the real value of a property.
 
 \RC{Typos : Page 1 : « 10395 property »}
+
+Many thanks for identifying this typo. It has been corrected in the new submission of the manuscript.
 
@@
 
+Thank you for your review and comments.
+In order to compare the accuracy of different variations of the XGBoost algorithm, the result of 8 models are now presented table I and III.
+The models are comparing XGBoost algorithms using tree based or linear regression booster.
+In addition each booster type is calculated according different learning objective output: squared error, squared log error, gamma, and tweedie.
+
 # Reviewer \#3
@@
 
+We thank you for your thoughtful and thorough review and we believe your input has greatly improved the manuscript.
+
 \RC{Concerning the following statement: “techniques used usually include artificial neural networks due to the volatility of the market”, even though it is backed up by bibliographic references, it would be worth further explaining it.
 The same applies to the following statement "While these methods are efficient in a non-restricted space, they have limitations when they are used in coastal areas.", explain why.}
+
+A precision on the use of artificial neural networks techniques has been added to the manuscript as follow:
+
+> Indeed, artificial neural network techniques can quickly adjust to volatility changes, therefore incorporating sharp increases or decreases in their predictions [@aragones2007neural].
+However, the model obtained from most artificial neural networks are hardly interpretable [@rudin2019stop].
+
+The limitation of classic models applied to coastal areas is presented as follow:
+
+> This problem is known as "finite area smoothing" and occurs when predictions from a model are approximated across geographical barriers, such as irregular coastal shapes, which would lead to poorly predicted values [@ramsay2002spline].
 
@@
 
+```{r}
+geocode_dublin_2018 <- here::here("data/ppr_dublin_2018.csv") %>% 
+  readr::read_csv()
+
+house_dublin_2018 <- geocode_dublin_2018 %>% 
+  janitor::clean_names() %>% 
+  dplyr::filter(!grepl('apartment|Apartment|APART|APT|Apt', address))
+
+final_dublin_2018 <- here::here("data/data_dublin_2018_osm_features.rds") %>% 
+  readr::read_rds()
+```
+
+The total amount of property sold in 2018 in the county Dublin which appears in the Property Price Register is `r nrow(geocode_dublin_2018)`.
+From these, all property corresponding to apartments were removed (`r scales::percent(1-(nrow(house_dublin_2018)/nrow(geocode_dublin_2018)), accuracy = 0.1)` of the original dataset).
+Finally, unsuccessfully geocoded properties were also removed from the original dataset (`r scales::percent(1-(nrow(final_dublin_2018)/nrow(house_dublin_2018)), accuracy = 0.1)`) to obtain the `r nrow(final_dublin_2018)` of the final dataset used to identify the urban and socio-economic correlates of property prices.
+
+These information have been added as it in the new version of the manuscript.
+
+Urban features are obtained using the R package osmdata which queries the Overpass API. This information as also been added to the manuscript.
+
 \RC{Input data is not available in the github repository given by the authors, can we suppose it is due to restrictions to share the data, if so, this should be said clearly.}
+
+The github repository has been update to ensure the full reproducibility of the results.
+The geocoded dataset has been added as well as the scripts to obtain the batch geocoding of the properties, to access to the Overpass API for the distance calculation with the urban features and the links to download and process the census data.
 
@@
 
+the URLs for the Central Statistics Office and the All-Island Research Observatory have been added to the manuscript.
+
+> The results of Irish 2016 census consultation is accessible through both the Central Statistics Office [@cso2020] and All-Island Research Observatory [@airo2020].
+
 \RC{explicit acronym "SD".}
+
+The acronym SD has been explicitly introduced in the manuscript.
+
+# References

### 2020-08-18 script/distance_osm_feature.R

@@ -1,0 +1,99 @@
+################################################################################
+#                        Distance to OSM Features                              #
+################################################################################
+
+# libraries --------------------------------------------------------------------
+library(tidyverse)
+library(osmdata)
+library(sp)
+library(sf)
+
+# options ----------------------------------------------------------------------
+# https://wiki.openstreetmap.org/wiki/Overpass_API
+# osmdata::get_overpass_url()
+osmdata::set_overpass_url('https://overpass-api.de/api/interpreter')
+
+# data -------------------------------------------------------------------------
+geocode_dublin_2018 <- here::here("data/geocode_dublin_2018.rds") %>% 
+  readr::read_rds()
+
+# case with 1 OSM feature ######################################################
+# convert data to sf class
+data_sf <- st_as_sf(geocode_dublin_2018, coords = c("lng", "lat"))
+st_crs(data_sf) <- st_crs(4326) # assign crs
+data_sf <- st_transform(data_sf, crs = 32721) # transform
+
+# OSM feature ------------------------------------------------------------------
+variable_osm <- osmdata::opq("Dublin, Ireland") %>%
+  osmdata::add_osm_feature(key = "leisure", value = "park")
+
+# variable_osm %>%
+#   osmdata::osmdata_sp() %>%
+#   magrittr::use_series(osm_points) %>%
+#   sp::plot()
+
+variable_sf <- variable_osm %>% 
+  osmdata::osmdata_sf() %>%
+  magrittr::use_series(osm_points) %>% 
+  magrittr::use_series(geometry) %>% 
+  sf::st_sf()
+
+st_crs(variable_sf) <- st_crs(4326) # assign crs
+variable_sf <- st_transform(variable_sf, crs = 32721) # transform
+variable_sf <- st_combine(variable_sf)
+
+# distance to OSM feature ------------------------------------------------------
+# by_element = TRUE returns the shorter distance
+geocode_dublin_2018$variable <- st_distance(x = data_sf, y = variable_sf)
+
+# Case with all OSM features ###################################################
+# manual list OSM features -----------------------------------------------------
+list_osm_features <- file.path(data_path, "list_osm_features.rds") %>% 
+  readr::read_rds() %>% 
[…]

### 2020-08-18 script/functions_project.R

@@ -1,7 +1,32 @@
+################################################################################
+geocode_batch <- function(address){
+  # for osm geocoding ----------------------------------------------------------
+  res <-
+    jsonlite::fromJSON(
+      gsub(
+        "\\@addr\\@",
+        gsub("\\s+", "\\%20", address),
+        'http://nominatim.openstreetmap.org/search/@addr@?format=json&addressdetails=0&limit=1'
+      )
+    )
+  # for pickpoint geocoding ----------------------------------------------------
+  # res <- 
+  #   prettymapr::geocode(
+  #     location = address, 
+  #     source = "pickpoint", 
+  #     key = config::get("pickpoint")$key
+  #   )
+  if (length(res) > 0) {
+    dplyr::select(res, display_name, lat, lon)
+  } else {
+    data.frame(display_name = NA, lat = NA, lon = NA)
+  }
+}
+
+################################################################################
 gam_density <- function (x, view = NULL, cond = list(), n.grid = 40, too.far = 0, 
                     col = NA, color = "heat", contour.col = NULL, se = -1, type = "link", 
-                    plot.type = "persp", zlim = NULL, nCol = 50, ...) 
-{
+                    plot.type = "persp", zlim = NULL, nCol = 50, ...) {
   fac.seq <- function(fac, n.grid) {
     fn <- length(levels(fac))
     gn <- n.grid
@@ -124,18 +149,97 @@ gam_density <- function (x, view = NULL, cond = list(), n.grid = 40, too.far = 0
   dat <- as.data.frame(z)
   rownames(dat) <- m1
   colnames(dat) <- m2
-  
   dat %>%
     tibble::rownames_to_column("lng") %>%
-    tidyr::gather(lat,value, -lng) %>%
-    dplyr::mutate(lng = as.numeric(lng)) %>%
-    dplyr::mutate(lat = as.numeric(lat))
+    tidyr::gather(lat, value, -lng) %>%
+    dplyr::mutate(
+      lng = as.numeric(lng),
+      lat = as.numeric(lat)
+    )
 }
-
+################################################################################
 read_join <- function(file_name){
   file.path(data_path, "aero_data", file_name) %>% 
     readr::read_csv(col_types = list(GEOG_ID = col_character())) %>% 
     dplyr::filter(GEOG_ID %in% list_geogid)
 }
-
+################################################################################
[…]

### 2020-08-18 script/geocode_dublin_2018.R

@@ -1,0 +1,40 @@
+################################################################################
+#                         Dublin Property Geocoding                            #
+################################################################################
+
+# libraries --------------------------------------------------------------------
+library(here)
+library(tidyverse)
+library(utils)
+library(janitor)
+library(jsonlite)
+library(config)
+library(prettymapr)
+
+# sources ----------------------------------------------------------------------
+source(here::here("script/functions_project.R"))
+
+# download PPR -----------------------------------------------------------------
+utils::download.file(
+  url = "https://www.propertypriceregister.ie/website/npsra/ppr/npsra-ppr.nsf/Downloads/PPR-2018-Dublin.csv/$FILE/PPR-2018-Dublin.csv",
+  destfile = here::here("data/ppr_dublin_2018.csv")
+)
+
+# geocode addresses ------------------------------------------------------------
+geocode_dublin_2018 <- here::here("data/ppr_dublin_2018.csv") %>% 
+  readr::read_csv() %>% 
+  dplyr::rename(
+    date_of_sale = `Date of Sale (dd/mm/yyyy)`,
+    price = `Price (<U+0080>)`
+  ) %>%
+  dplyr::mutate(price = readr::parse_number(price)) %>%
+  dplyr::mutate(date_of_sale = as.Date(date_of_sale, format = "%d/%m/%Y")) %>%
+  dplyr::mutate(full_address = paste(Address, County, "Ireland")) %>%
+  tibble::rowid_to_column("sale_id") %>%
+  dplyr::mutate(year = lubridate::year(date_of_sale)) %>% 
+  janitor::clean_names() %>% 
+  dplyr::filter(!grepl('apartment|Apartment|APART|APT|Apt', address)) %>% 
+  dplyr::rowwise() %>% 
+  dplyr::mutate(geocode_batch(address)) %>%
+  dplyr::filter(between(lat, 50, 53.7) & between(lng, -7, -5.5)) %>% #gps artefact
+  readr::write_rds(here::here("data/geocode_dublin_2018.rds"))

### 2020-08-18 temp.R

@@ -1,0 +1,53 @@
+################################################################################
+#                      Property Distance to OSM Feature                        #
+################################################################################
+
+data_sf <- st_as_sf(data_dublin_2018, coords = c("lng", "lat"))
+st_crs(data_sf) <- st_crs(4326) # assign crs
+data_sf <- st_transform(data_sf, crs = 32721) # transform
+
+list_osm_features <- readr::read_rds(here::here("data/list_osm_features.rds")) %>% 
+  dplyr::filter(Value != "*") %>% 
+  dplyr::filter(!stringr::str_detect(Value, "/|[(]|User Defined|Number|Date|Name|see opening_hours|[:digit:]")) %>% 
+  dplyr::filter(!stringr::str_detect(Key, ":"))
+
+for (i in 1:nrow(list_osm_features)) {
+  Key <- list_osm_features[i,"Key"]
+  Value <- list_osm_features[i,"Value"]
+  variable_name <- paste(Key, Value, sep = "_")
+  
+  paste(i, variable_name) %>% print()
+  # obtain gps coordinates of geographic variables (e.g., cycle lanes)
+  
+  variable_osm <- osmdata::opq('Dublin, Ireland') %>%
+    osmdata::add_osm_feature(key = Key, value = Value)
+  
+  # variable_osm %>% 
+  #   osmdata::osmdata_sp() %>% 
+  #   magrittr::use_series(osm_points) %>% 
+  #   sp::plot()
+  
+  variable_sf <- variable_osm %>% 
+    osmdata::osmdata_sf() %>%
+    magrittr::use_series(osm_lines) %>% 
+    magrittr::use_series(geometry) %>% 
+    st_sf()
+  
+  if (nrow(variable_sf) == 0) {
+    
+    data_dublin_2018 <- data_dublin_2018 %>% 
+      dplyr::mutate(!!variable_name := NA)
+    
+  } else {
+    
+    st_crs(variable_sf) <- st_crs(4326) # assign crs
+    variable_sf <- st_transform(variable_sf, crs = 32721) # transform
+    variable_sf <- st_combine(variable_sf)
+    
+    # distance
+    data_dublin_2018 <- data_dublin_2018 %>% 
+      dplyr::mutate(!!variable_name := as.numeric(st_distance(x = data_sf, y = variable_sf)))
+    
+  }
+  Sys.sleep(5)
+}

### 2020-08-18 manuscript/manuscript.Rmd
paper submission

@@
 library(rticles)
+library(scales)
 library(tidyverse)
@@
 # data -------------------------------------------------------------------------
-geocode_dublin_2018 <- here::here("data/ppr_dublin_2018.csv") %>% 
-  readr::read_csv()
+n_raw <- here::here("data/ppr_dublin_2018.csv") %>% 
+  readr::read_csv() %>% 
+  nrow()
 
-house_dublin_2018 <- geocode_dublin_2018 %>% 
-  dplyr::filter(!grepl('apartment|Apartment|APART|APT|Apt', Address))
+n_house <- here::here("data/ppr_dublin_2018.csv") %>% 
+  readr::read_csv() %>% 
+  dplyr::filter(!grepl('apartment|Apartment|APART|APT|Apt', Address)) %>% 
+  nrow()
 
-final_dublin_2018 <- here::here("data/data_dublin_2018_osm_features.rds") %>% 
+n_final <- here::here("data/data_dublin_2018_features.rds") %>% 
+  readr::read_rds() %>% 
+  nrow()
+
+# all dublin features ----------------------------------------------------------
+data_dublin_2018_features <- here::here("data/data_dublin_2018_features.rds") %>%
   readr::read_rds()
 
-# data path --------------------------------------------------------------------
-data_path <- here::here() %>% 
-  dirname() %>% 
-  dirname() %>% 
-  paste0("/data/property_price_ireland_data")
-
-# all price registry------------------------------------------------------------
-data_dublin_geocoded <- file.path(data_path, "data_dublin_geocoded_clean.rds") %>% 
-  readr::read_rds() %>%
-  janitor::clean_names() %>%
-  dplyr::mutate(
-    year_fact = as.factor(as.character(year))
-  ) %>% 
-  dplyr::filter(year == 2018)
-
-mean_dat <- mean(data_dublin_geocoded$price)
-sd_dat <- sd(data_dublin_geocoded$price)
-
-data_dublin_geocoded <- data_dublin_geocoded %>%
-  dplyr::filter(price > mean_dat - sd_dat)
-
-# all dublin features ----------------------------------------------------------
-data_dublin_2018_features <- data_path %>%
-  file.path("data_dublin_2018_osm_features.rds") %>%
-  readr::read_rds() %>% 
-  janitor::remove_empty(which = "cols") %>% 
-  janitor::clean_names() %>%
-  dplyr::mutate(geog_id = stringr::str_remove(geogid, "[A]"))
-
-list_geogid <- unique(data_dublin_2018_features$geog_id)
-
-data_dublin_2018_features <- c(
-  "carers.csv",
-  "disabilty_age_group.csv", 
-  "general_health.csv",
-  "housing_occupancy.csv",
-  "housing_rooms.csv",
-  "housing_tenure.csv",
-  "housing_type.csv",
-  "population.csv",
-  "religion.csv"
-  ) %>%
-  purrr::map(read_join) %>% 
-  purrr::reduce(dplyr::left_join) %>% 
-  dplyr::inner_join(data_dublin_2018_features, by = c("GEOG_ID" = "geog_id"))
-
 # train test split -------------------------------------------------------------
-train_index <- sample(1:nrow(data_dublin_2018_features), 0.8 * nrow(data_dublin_2018_features))
-test_index <- setdiff(1:nrow(data_dublin_2018_features), train_index)
+train_index <- sample(1:n_final, 0.8 * n_final)
+test_index <- setdiff(1:n_final, train_index)
 
@@
 Data is filed electronically by persons doing the conveyancing of the property on behalf of the purchaser, and it must be noted that errors may occur when the data is being filed.
-
 Because of the evolution of urban planning and the evolution of socio-economic features measured by the 2016 census, only properties sold in 2018 were processed to avoid potential temporal incoherence.
-The total amount of property sold in 2018 in the county Dublin which appears in the Property Price Register is `r nrow(geocode_dublin_2018)`.
-From these, all property corresponding to apartments were removed (`r scales::percent(1-(nrow(house_dublin_2018)/nrow(geocode_dublin_2018)), accuracy = 0.1)` of the original dataset).
-Finally, unsuccessfully geocoded properties were also removed from the original dataset (`r scales::percent(1-(nrow(final_dublin_2018)/nrow(house_dublin_2018)), accuracy = 0.1)`) to obtain the `r nrow(final_dublin_2018)` of the final dataset used to identify the urban and socio-economic correlates of property prices.
+The total amount of property sold in 2018 in the county Dublin which appears in the Property Price Register is `r n_raw`.
+From these, all property corresponding to apartments were removed (`r scales::percent(1-(n_house/n_raw), accuracy = 0.1)` of the original dataset).
+Finally, unsuccessfully geocoded properties were also removed from the original dataset (`r scales::percent(1-(n_final/n_house), accuracy = 0.1)`) to obtain the `r n_final` of the final dataset used to identify the urban and socio-economic correlates of property prices.
 
@@
 
+```{r descriptives}
+data_mean <- data_dublin_2018_features$price %>% 
+  mean(na.rm = TRUE) %>% round(0) %>% scales::dollar(suffix = "", prefix = "€")
+data_sd <- data_dublin_2018_features$price %>% 
+  sd(na.rm = TRUE) %>% round(0) %>% scales::dollar(suffix = "", prefix = "€")
+```
+
+In 2018, the average price of a sold property in Dublin was `r data_mean` with a standard deviation of `r data_sd` (Figure \@ref(fig:distrib-plot)).
+
 ```{r gam-model}
@@
 # knots ------------------------------------------------------------------------
-# boundaries_dublin <- 
-#   osmdata::opq(bbox = 'Dublin, Ireland') %>%
-#   osmdata::add_osm_feature(key = 'admin_level', value = '7') %>%
-#   osmdata::osmdata_sf() %>% 
-#   osmdata::unique_osmdata()
-# 
-# bb <- st_bbox(boundaries_dublin$osm_multipolygons$geometry)
-# 
-# readr::write_rds(bb, here::here("data/bb.rds"))
-
 bb <- readr::read_rds(here::here("data/bb.rds"))
@@
   price ~ s(lng, lat, bs = "so", xt = list(bnd = bound)), 
-  data = dplyr::filter(data_dublin_geocoded, year == 2018), 
+  data = data_dublin_2018_features, 
   method = "REML", 
@@
 
-gam_pred <- gam_density(gam_dublin_gps, too.far = 0.05, n.grid = 30)
+gam_pred <- gam_density(gam_dublin_gps, too.far = 0.05, n.grid = 100)
 ```
 
-In 2018, the average price of a sold property in Dublin was `r scales::dollar(round(mean_dat, 0), suffix = "", prefix = "€")` with a standard deviation of `r scales::dollar(round(sd_dat, 0), suffix = "", prefix = "€")`.
-In order to remove potential human errors and outliers, prices higher or lower than 1 SD were removed from the original dataset.
-
-```{r distrib-plot, fig.height=3, out.width="0.98\\columnwidth", fig.pos ="!h", fig.cap="(ref:distrib-plot)"}
-data_dublin_geocoded %>%
-  dplyr::filter(year == 2018) %>%
+```{r distrib-plot, fig.height=4, out.width="0.98\\columnwidth", fig.pos ="!h", fig.cap="(ref:distrib-plot)"}
+data_dublin_2018_features %>%
   magrittr::use_series(price) %>%
@@
   )) +
-  scale_x_continuous("Price", labels = scales::dollar_format(suffix = "", prefix = "€")) +
+  scale_x_continuous("Price", labels = dollar_format(suffix = "", prefix = "€")) +
   scale_y_continuous("Density") +
@@
   dplyr::filter(value > 200000) %>%
-  ggplot() +
-  geom_raster(aes(x = lng, y = lat, fill = value)) +
-  geom_sf(data = dublin_sea, fill = "light blue") +
+  ggplot(aes(x = lng, y = lat)) + 
+  geom_raster(aes(fill = value), na.rm = TRUE) +
+  geom_contour(aes(z = value), color = "black", na.rm = TRUE) +
+  geom_sf(data = dublin_sea, fill = "light blue", inherit.aes = FALSE) +
   scale_fill_gradient(
@@
     breaks = seq(53.2, 53.5, 0.1)) +
-  theme_apa() +
+  theme_classic() +
   theme(
     text = element_text(family = "serif", size = 12),
-    legend.position = "none"
+    legend.position = "none",
+    panel.background = element_rect(fill = "grey50")
   )
@@
 
-The density of housing prices reveals a unimodal distribution of property prices (Figure \@ref(fig:distrib-plot)).
+The density of housing prices reveals a unimodal distribution of property prices.
 In addition, the geographical distribution of property prices indicates a higher valuation in the South-West of the city as well as on the coast line (Figure \@ref(fig:distrib-gps)).
@@
   select(-.estimator) %>% 
-  mutate(.estimate = round(.estimate, 2))
+  mutate(.estimate = case_when(
+    .metric == "rsq" ~ scales::percent(.estimate, accuracy = 0.1),
+    TRUE ~ scales::dollar(.estimate, suffix = "", prefix = "€")
+  )) %>% 
+  dplyr::pull(.estimate, .metric)
 
@@
 
-Using the XGBoost algorithm with a tree based booster and a tweedie regression with log-link learning objective, the 160 urban features are explaining `r scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"])` of the property price variance (`r lm_result$statistic$modelfit$r2`) with a $RMSE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "rmse"], suffix = "", prefix = "€")` and a $MAE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "mae"], suffix = "", prefix = "€")`.
+Using the XGBoost algorithm with a tree based booster and a tweedie regression with log-link learning objective, the 160 urban features are explaining `r metrics_result["rsq"]` of the property price variance (`r lm_result$statistic$modelfit$r2`) with a $RMSE$ of `r metrics_result["rmse"]` and a $MAE$ of `r metrics_result["mae"]`.
 
@@
     color = "black",
-    size = 0.1
+    size = 0.1,
+    na.rm = TRUE
   ) +
@@
   theme(
-    text = element_text(family = "serif", size = 12),
-    legend.position = "none"
+    text = element_text(family = "serif", size = 12)
   ) +
-  labs(
-    title = "A",
-    subtitle = latex2exp::TeX(paste(
-      "$R^2 =$", 
-      scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"]),
-      ";",
-      "$RMSE =$",
-      scales::dollar(
-        metrics_result$.estimate[metrics_result$.metric == "rmse"], 
-        suffix = "", 
-        prefix = "€"
-      ),
-      ";",
-      "$MAE =$",
-      scales::dollar(
-        metrics_result$.estimate[metrics_result$.metric == "mae"], 
-        suffix = "", 
-        prefix = "€"
-      )
-      )
-    )
-  )
+  ggtitle("A")
 
@@
     fullrange = TRUE,
-    size = 0.1
+    size = 0.1,
+    na.rm = TRUE
   ) +
@@
     latex_options = "HOLD_position"
-  )
+  ) %>% 
+  kableExtra::column_spec(2, width = "10em")
 ```
@@
   kableExtra::kable_styling(font_size = 8, latex_options = "HOLD_position") %>% 
-  pack_rows(index = c("Tree based" = 4, "Linear regression based" = 4)) %>% 
+  kableExtra::pack_rows(index = c("Tree based" = 4, "Linear regression based" = 4)) %>% 
   kableExtra::footnote("Missing value ares due to incompatibilities between data and methods.")
@@
   theme_apa() +
-  theme(
-    text = element_text(family = "serif", size = 12),
-    legend.position = "none"
-  ) +
-    labs(
-    title = "A",
-    subtitle = latex2exp::TeX(paste(
-      "$R^2 =$", 
-      scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"]),
-      ";",
-      "$RMSE =$",
-      scales::dollar(
-        metrics_result$.estimate[metrics_result$.metric == "rmse"], 
-        suffix = "", 
-        prefix = "€"
-      ),
-      ";",
-      "$MAE =$",
-      scales::dollar(
-        metrics_result$.estimate[metrics_result$.metric == "mae"], 
-        suffix = "", 
-        prefix = "€"
-      )
-      )
-    )
-  )
+  theme(text = element_text(family = "serif", size = 12)) +
+  ggtitle("A")
 
@@
 
-Using the XGBoost algorithm with a tree based booster and a gamma regression with log-link learning objective, the 48 socio-economic features explain `r scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"])` of the property price variance (`r lm_result$statistic$modelfit$r2`) with a $RMSE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "rmse"], suffix = "", prefix = "€")` and a $MAE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "mae"], suffix = "", prefix = "€")`.
+Using the XGBoost algorithm with a tree based booster and a gamma regression with log-link learning objective, the 48 socio-economic features explain `r metrics_result["rsq"]` of the property price variance (`r lm_result$statistic$modelfit$r2`) with a $RMSE$ of `r metrics_result["rmse"]` and a $MAE$ of `r metrics_result["mae"]`.
 
@@
     latex_options = "HOLD_position"
-  )
+  ) %>% 
+  kableExtra::column_spec(2, width = "10em") 
 ```
@@
 The main reason is the structure of the market, mainly made of residential properties and few apartment buildings.
-In this study, the prices of `r nrow(data_dublin_2018_features)` properties sold in 2018 in the area of Dublin were analysed from the Property Services Regulatory Authority dataset in order to identify the potential factors correlating with these prices.
+In this study, the prices of `r n_final` properties sold in 2018 in the area of Dublin were analysed from the Property Services Regulatory Authority dataset in order to identify the potential factors correlating with these prices.
 The enforcement of a housing price registry has been shown as being an essential tool to regulate the housing market [@kitchin2013making].

### 2020-08-18 manuscript/manuscript.tex
paper submission

@@
 \hypersetup{
-            pdftitle={Urban and socio-economic correlates of property price in Dublin's area},
+            pdftitle={Urban and socio-economic correlates of property prices in Dublin's area},
             pdfkeywords={Property price, Housing Market, Feature Analysis, Machine Learning, Geocoding},
@@
 % Do not put math or special symbols in the title.
-\title{Urban and socio-economic correlates of property price in Dublin's area}
+\title{Urban and socio-economic correlates of property prices in Dublin's area}
 
@@
 Data is filed electronically by persons doing the conveyancing of the property on behalf of the purchaser, and it must be noted that errors may occur when the data is being filed.
-
 Because of the evolution of urban planning and the evolution of socio-economic features measured by the 2016 census, only properties sold in 2018 were processed to avoid potential temporal incoherence.
@@
 From these, all property corresponding to apartments were removed (10.3\% of the original dataset).
-Finally, unsuccessfully geocoded properties were also removed from the original dataset (38.0\%) to obtain the 10389 of the final dataset used to identify the urban and socio-economic correlates of property prices.
+Finally, unsuccessfully geocoded properties were also removed from the original dataset (38.0\%) to obtain the 10387 of the final dataset used to identify the urban and socio-economic correlates of property prices.
 
@@
 
-In 2018, the average price of a sold property in Dublin was €384,892 with a standard deviation of €175,476.
-In order to remove potential human errors and outliers, prices higher or lower than 1 SD were removed from the original dataset.
+In 2018, the average price of a sold property in Dublin was €394,093 with a standard deviation of €170,219 (Figure \ref{fig:distrib-plot}).
 
@@
 
-The density of housing prices reveals a unimodal distribution of property prices (Figure \ref{fig:distrib-plot}).
+The density of housing prices reveals a unimodal distribution of property prices.
 In addition, the geographical distribution of property prices indicates a higher valuation in the South-West of the city as well as on the coast line (Figure \ref{fig:distrib-gps}).
@@
 
-Using the XGBoost algorithm with a tree based booster and a tweedie regression with log-link learning objective, the 160 urban features are explaining 45\% of the property price variance (\(F(1, 2076) = 1,690.80\), \(p < .001\)) with a \(RMSE\) of €126,626 and a \(MAE\) of €85,073.07.
+Using the XGBoost algorithm with a tree based booster and a tweedie regression with log-link learning objective, the 160 urban features are explaining 44.9\% of the property price variance (\(F(1, 2076) = 1,690.80\), \(p < .001\)) with a \(RMSE\) of €126,626 and a \(MAE\) of €85,073.
 
@@
 \fontsize{8}{10}\selectfont
-\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X}
+\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright\arraybackslash}p{10em}>{\raggedright}X}
 \toprule
@@
 \multicolumn{4}{l}{\textbf{Linear regression based}}\\
-\hspace{1em}squarederror & €146,781 & 0.26 & €107,890\\
-\hspace{1em}squaredlogerror & €406,108 & 0.01 & €368,877\\
-\hspace{1em}gamma & €182,432 & 0.08 & €136,441\\
-\hspace{1em}tweedie & €164,439 & 0.13 & €122,259\\
+\hspace{1em}squarederror & €146,985 & 0.25 & €108,141\\
+\hspace{1em}squaredlogerror & €406,180 & 0.01 & €368,972\\
+\hspace{1em}gamma & €171,736 & 0.11 & €128,489\\
+\hspace{1em}tweedie & €174,236 & 0.10 & €130,269\\
 \bottomrule
@@
 
-Using the XGBoost algorithm with a tree based booster and a gamma regression with log-link learning objective, the 48 socio-economic features explain 43\% of the property price variance (\(F(1, 2076) = 1,554.37\), \(p < .001\)) with a \(RMSE\) of €129,221 and a \(MAE\) of €86,949.15.
+Using the XGBoost algorithm with a tree based booster and a gamma regression with log-link learning objective, the 48 socio-economic features explain 42.8\% of the property price variance (\(F(1, 2076) = 1,554.37\), \(p < .001\)) with a \(RMSE\) of €129,221 and a \(MAE\) of €86,949.
 

### 2020-08-18 temp.R

@@ -1,53 +1,59 @@
-################################################################################
-#                      Property Distance to OSM Feature                        #
-################################################################################
-
-data_sf <- st_as_sf(data_dublin_2018, coords = c("lng", "lat"))
-st_crs(data_sf) <- st_crs(4326) # assign crs
-data_sf <- st_transform(data_sf, crs = 32721) # transform
-
-list_osm_features <- readr::read_rds(here::here("data/list_osm_features.rds")) %>% 
-  dplyr::filter(Value != "*") %>% 
-  dplyr::filter(!stringr::str_detect(Value, "/|[(]|User Defined|Number|Date|Name|see opening_hours|[:digit:]")) %>% 
-  dplyr::filter(!stringr::str_detect(Key, ":"))
-
-for (i in 1:nrow(list_osm_features)) {
-  Key <- list_osm_features[i,"Key"]
-  Value <- list_osm_features[i,"Value"]
-  variable_name <- paste(Key, Value, sep = "_")
-  
-  paste(i, variable_name) %>% print()
-  # obtain gps coordinates of geographic variables (e.g., cycle lanes)
-  
-  variable_osm <- osmdata::opq('Dublin, Ireland') %>%
-    osmdata::add_osm_feature(key = Key, value = Value)
-  
-  # variable_osm %>% 
-  #   osmdata::osmdata_sp() %>% 
-  #   magrittr::use_series(osm_points) %>% 
-  #   sp::plot()
-  
-  variable_sf <- variable_osm %>% 
-    osmdata::osmdata_sf() %>%
-    magrittr::use_series(osm_lines) %>% 
-    magrittr::use_series(geometry) %>% 
-    st_sf()
-  
-  if (nrow(variable_sf) == 0) {
-    
-    data_dublin_2018 <- data_dublin_2018 %>% 
-      dplyr::mutate(!!variable_name := NA)
-    
-  } else {
-    
-    st_crs(variable_sf) <- st_crs(4326) # assign crs
-    variable_sf <- st_transform(variable_sf, crs = 32721) # transform
-    variable_sf <- st_combine(variable_sf)
-    
-    # distance
-    data_dublin_2018 <- data_dublin_2018 %>% 
-      dplyr::mutate(!!variable_name := as.numeric(st_distance(x = data_sf, y = variable_sf)))
-    
+function (object, show = c("estimate", "se"), contour = TRUE, 
+          contour_col = "black", n_contour = NULL, xlab, ylab, 
[…]

### 2020-08-19 manuscript/manuscript.Rmd
final submission

@@
   )
+
 knitr::opts_chunk$set(
@@
   echo = FALSE,
-  cache = TRUE,
   fig.retina = 3
@@
 Beside their inherent characteristics such as size, design and materials [@bourassa2007spatial; @liu2013spatial], property valuation is dictated by multiple external factors explaining their spatial autocorrelation [@basu1998analysis].
-Some of these factors are macroeconomic [@antonakakis2016dynamic], related to the evolution of population general wealth while others are related to the characteristics of the neighbourhood [@dubin1998predicting].
-These characteristics include the urban and socio economic structure of the neighbourhood [@goodman1998housing].
+Some of these factors are macroeconomic [@antonakakis2016dynamic], related to the evolution of population general wealth while others are related to the characteristics of the neighborhood [@dubin1998predicting].
+These characteristics include the urban and socio economic structure of the neighborhood [@goodman1998housing].
 However, it is difficult to evaluate the influence of such external factors on property valuation [@clapp2002predicting].
@@
 Indeed, artificial neural network techniques can quickly adjust to volatility changes, therefore incorporating sharp increases or decreases in their predictions [@aragones2007neural].
-However, the model obtained from most artificial neural networks are hardly interpretable [@rudin2019stop].
+However, the model obtained from most artificial neural networks are difficult to interpret [@rudin2019stop].
 
 In this paper, an alternative machine learning algorithm is used to identify urban and socio-economic correlates of property price evolution in the area of Dublin, Ireland in 2018.
-Due to the high number of potential features, an eXtreme Gradient Boosting (XGBoost) regression model is believed to obtain the best results with low parameter customisation and high speed [@chen2016xgboost].
+Due to the high number of potential features, an eXtreme Gradient Boosting (XGBoost) regression model is believed to obtain the best results with low parameter customization and high speed [@chen2016xgboost].
 Contrary to most artificial neural network techniques, XGBoost is an interpretation-focused method which automatically provides estimates of feature importance from a trained predictive model.
@@
 Because of the evolution of urban planning and the evolution of socio-economic features measured by the 2016 census, only properties sold in 2018 were processed to avoid potential temporal incoherence.
-The total amount of property sold in 2018 in the county Dublin which appears in the Property Price Register is `r n_raw`.
-From these, all property corresponding to apartments were removed (`r scales::percent(1-(n_house/n_raw), accuracy = 0.1)` of the original dataset).
-Finally, unsuccessfully geocoded properties were also removed from the original dataset (`r scales::percent(1-(n_final/n_house), accuracy = 0.1)`) to obtain the `r n_final` of the final dataset used to identify the urban and socio-economic correlates of property prices.
+In 2018, the Property Price Register references `r n_raw` properties sold in Dublin's area.
+Among these properties, those corresponding to apartments have been removed from the original dataset in order to evaluate only houses (`r scales::percent(1-(n_house/n_raw), accuracy = 0.1)` of the properties).
+Finally, some properties have been unsuccessfully geocoded because of mistakes on their address or because they have not been found (`r scales::percent(1-(n_final/n_house), accuracy = 0.1)` of the properties).
+These properties have also been removed from the original dataset to obtain a final dataset of `r n_final` properties.
 
@@
 While these methods are efficient in a non-restricted space, they have limitations when they are used in coastal areas.
-This problem is known as "finite area smoothing" and occurs when predictions from a model are approximated across geographical barriers, such as irregular coastal shapes, which would lead to poorly predicted values [@ramsay2002spline].
-To deal with the influence of geographical boundaries on price estimation, a generalised additive model using soap film smoothing has been used [@wood2008soap].
+This problem is known as "finite area smoothing" and occurs when predictions from a model are approximated across geographical barriers, such as irregular coastal shapes, which can lead to poorly predicted values [@ramsay2002spline].
+To deal with the influence of geographical boundaries on price estimation, a generalized additive model using soap film smoothing has been used [@wood2008soap].
 
@@
 In addition to the distance to every urban feature, each property sold in 2018 was related to 48 socio-economic features of the small area including the property, as measured by the Irish census 2016.
-The results of Irish 2016 census consultation is accessible through both the Central Statistics Office [@cso2020] and All-Island Research Observatory [@airo2020].
+The results of Irish 2016 census consultation is accessible through both the Central Statistics Office [@cso2020] and the All-Island Research Observatory [@airo2020] websites.
 The data obtained can be mapped over small area boundaries which are fractions of Irish Electoral Division map.
@@
 Both tree and linear regression based boosting model are used for every of the continuous output learning task in order to compare their accuracy with urban and socio-economic features.
-As no specific weight or bias in the models has been hypothesized, the parameters of the XGBoost function have been set to the default value.
-Then, the most accurate model is used to identify the features that are contributing the most to the prediction of property prices.
-This importance is a measure of the improvement in accuracy brought by a feature based on its coefficient's absolute magnitude.
+As no specific weight or bias in the models have been hypothesized, parameters of the XGBoost algorithms have been kept as default values.
+Next, the most accurate model is used to identify the features that are contributing most to the prediction of property prices.
+This contribution (also called "importance") is a measure of the improvement in accuracy brought by a feature.
+It is based on the absolute magnitude of feature's coefficient.
 
@@
   scale_y_continuous("Density") +
-  scale_color_gradient(
-    low = '#3a53a4', 
-    high = '#ed1c24', 
-    ) +
+  scale_color_gradient(low = '#3a53a4', high = '#ed1c24') +
   theme_apa() +
   theme(
-    text = element_text(family = "serif", size = 12),
+    text = element_text(family = "serif", size = 14),
     axis.text.y = element_blank(),
@@
   theme(
-    text = element_text(family = "serif", size = 12),
+    text = element_text(family = "serif", size = 14),
     legend.position = "none",
@@
 
-```{r osm-features-xgb, fig.cap="(ref:osm-features-xgb)", fig.height=7, fig.pos ="!h", out.width="0.98\\columnwidth"}
+```{r osm-features-xgb, fig.cap="(ref:osm-features-xgb)", fig.height=9, fig.pos ="!h", out.width="0.98\\columnwidth"}
 result_accuracy <- result %>%
@@
   theme(
-    text = element_text(family = "serif", size = 12)
+    text = element_text(family = "serif", size = 14)
   ) +
@@
 
-By analysing their importance (Table \@ref(tab:osm-features-table)), the most relevant geographical features to predict housing prices are the distance to an embassy (`r table_osm$Importance[1]`) and the distance to natural grasslands such as parks and gardens (`r table_osm$Importance[2]`).
+By analyzing their importance (Table \@ref(tab:osm-features-table)), the most relevant geographical features to predict housing prices are the distance to an embassy (`r table_osm$Importance[1]`) and the distance to natural grasslands such as parks and gardens (`r table_osm$Importance[2]`).
 
@@
 
-```{r census-features-xgb, fig.cap="(ref:census-features-xgb)", fig.height=7, fig.pos ="!h", out.width="0.98\\columnwidth"}
+```{r census-features-xgb, fig.cap="(ref:census-features-xgb)", fig.height=9, fig.pos ="!h", out.width="0.98\\columnwidth"}
 result_accuracy <- result %>%
@@
   theme_apa() +
-  theme(text = element_text(family = "serif", size = 12)) +
+  theme(text = element_text(family = "serif", size = 14)) +
   ggtitle("A")
@@
   theme_apa() +
-  theme(text = element_text(family = "serif", size = 12)) +
+  theme(text = element_text(family = "serif", size = 14)) +
   ggtitle("B")
@@
 The main reason is the structure of the market, mainly made of residential properties and few apartment buildings.
-In this study, the prices of `r n_final` properties sold in 2018 in the area of Dublin were analysed from the Property Services Regulatory Authority dataset in order to identify the potential factors correlating with these prices.
+In this study, the prices of `r n_final` properties sold in 2018 in the area of Dublin from the Property Services Regulatory Authority dataset were analyzed in order to identify the potential factors correlating with these prices.
 The enforcement of a housing price registry has been shown as being an essential tool to regulate the housing market [@kitchin2013making].
@@
 In addition, while some features appear to be more important than others in their contribution to property prices, these results are not causal.
-Thus, features' importance can be either the cause or the consequence of property prices in a neighborhood.
+Thus, a feature's importance can be either the cause or the consequence of property prices in a neighborhood
 
@@
 Similarly, the characteristics of inhabitants in the area such as religion, health and age is correlated to the evolution of housing prices.
-These results allow an understanding of why some areas have higher prices than others which is a relevant information not only for real estate agents in charge of property valuations but also to buyers in order to approximate the real value of a property.
+These results allow an understanding of why some areas have higher prices than others which is relevant information not only for real estate agents in charge of property valuations but also for buyers in order to estimate the real value of a property.
 
@@
 
-The authors would like to thank the developers of the following R packages used to process, analyse, display and report data: `r papaja::cite_r(file = "bib/r-references.bib")`.
+The authors would like to thank the developers of the following R packages used to process, analyze, display and report data: `r papaja::cite_r(file = "bib/r-references.bib")`.
 

### 2020-08-19 manuscript/manuscript.tex
final submission

@@
 Beside their inherent characteristics such as size, design and materials (Bourassa, Cantoni, and Hoesli 2007; Liu 2013), property valuation is dictated by multiple external factors explaining their spatial autocorrelation (Basu and Thibodeau 1998).
-Some of these factors are macroeconomic (Antonakakis and Floros 2016), related to the evolution of population general wealth while others are related to the characteristics of the neighbourhood (Dubin 1998).
-These characteristics include the urban and socio economic structure of the neighbourhood (Goodman and Thibodeau 1998).
+Some of these factors are macroeconomic (Antonakakis and Floros 2016), related to the evolution of population general wealth while others are related to the characteristics of the neighborhood (Dubin 1998).
+These characteristics include the urban and socio economic structure of the neighborhood (Goodman and Thibodeau 1998).
 However, it is difficult to evaluate the influence of such external factors on property valuation (Clapp, Kim, and Gelfand 2002).
@@
 Indeed, artificial neural network techniques can quickly adjust to volatility changes, therefore incorporating sharp increases or decreases in their predictions (Aragonés, Blanco, and Estévez 2007).
-However, the model obtained from most artificial neural networks are hardly interpretable (Rudin 2019).
+However, the model obtained from most artificial neural networks are difficult to interpret (Rudin 2019).
 
 In this paper, an alternative machine learning algorithm is used to identify urban and socio-economic correlates of property price evolution in the area of Dublin, Ireland in 2018.
-Due to the high number of potential features, an eXtreme Gradient Boosting (XGBoost) regression model is believed to obtain the best results with low parameter customisation and high speed (Chen and Guestrin 2016).
+Due to the high number of potential features, an eXtreme Gradient Boosting (XGBoost) regression model is believed to obtain the best results with low parameter customization and high speed (Chen and Guestrin 2016).
 Contrary to most artificial neural network techniques, XGBoost is an interpretation-focused method which automatically provides estimates of feature importance from a trained predictive model.
@@
 Because of the evolution of urban planning and the evolution of socio-economic features measured by the 2016 census, only properties sold in 2018 were processed to avoid potential temporal incoherence.
-The total amount of property sold in 2018 in the county Dublin which appears in the Property Price Register is 18665.
-From these, all property corresponding to apartments were removed (10.3\% of the original dataset).
-Finally, unsuccessfully geocoded properties were also removed from the original dataset (38.0\%) to obtain the 10387 of the final dataset used to identify the urban and socio-economic correlates of property prices.
+In 2018, the Property Price Register references 18665 properties sold in Dublin's area.
+Among these properties, those corresponding to apartments have been removed from the original dataset in order to evaluate only houses (10.3\% of the properties).
+Finally, some properties have been unsuccessfully geocoded because of mistakes on their address or because they have not been found (38.0\% of the properties).
+These properties have also been removed from the original dataset to obtain a final dataset of 10387 properties.
 
@@
 While these methods are efficient in a non-restricted space, they have limitations when they are used in coastal areas.
-This problem is known as ``finite area smoothing'' and occurs when predictions from a model are approximated across geographical barriers, such as irregular coastal shapes, which would lead to poorly predicted values (Ramsay 2002).
-To deal with the influence of geographical boundaries on price estimation, a generalised additive model using soap film smoothing has been used (Wood, Bravington, and Hedley 2008).
+This problem is known as ``finite area smoothing'' and occurs when predictions from a model are approximated across geographical barriers, such as irregular coastal shapes, which can lead to poorly predicted values (Ramsay 2002).
+To deal with the influence of geographical boundaries on price estimation, a generalized additive model using soap film smoothing has been used (Wood, Bravington, and Hedley 2008).
 
@@
 In addition to the distance to every urban feature, each property sold in 2018 was related to 48 socio-economic features of the small area including the property, as measured by the Irish census 2016.
-The results of Irish 2016 census consultation is accessible through both the Central Statistics Office (Central Statistics Office 2020) and All-Island Research Observatory (All-Island Research Observatory 2020).
+The results of Irish 2016 census consultation is accessible through both the Central Statistics Office (Central Statistics Office 2020) and the All-Island Research Observatory (All-Island Research Observatory 2020) websites.
 The data obtained can be mapped over small area boundaries which are fractions of Irish Electoral Division map.
@@
 Both tree and linear regression based boosting model are used for every of the continuous output learning task in order to compare their accuracy with urban and socio-economic features.
-As no specific weight or bias in the models has been hypothesized, the parameters of the XGBoost function have been set to the default value.
-Then, the most accurate model is used to identify the features that are contributing the most to the prediction of property prices.
-This importance is a measure of the improvement in accuracy brought by a feature based on its coefficient's absolute magnitude.
+As no specific weight or bias in the models have been hypothesized, parameters of the XGBoost algorithms have been kept as default values.
+Next, the most accurate model is used to identify the features that are contributing most to the prediction of property prices.
+This contribution (also called ``importance'') is a measure of the improvement in accuracy brought by a feature.
+It is based on the absolute magnitude of feature's coefficient.
 
@@
 \multicolumn{4}{l}{\textbf{Linear regression based}}\\
-\hspace{1em}squarederror & €150,833 & 0.21 & €111,788\\
-\hspace{1em}squaredlogerror & €406,947 & 0.00 & €369,713\\
-\hspace{1em}gamma & €180,313 & 0.06 & €136,059\\
-\hspace{1em}tweedie & €172,903 & 0.08 & €130,148\\
+\hspace{1em}squarederror & €150,925 & 0.21 & €111,810\\
+\hspace{1em}squaredlogerror & €406,947 & 0.00 & €369,709\\
+\hspace{1em}gamma & €183,987 & 0.05 & €138,978\\
+\hspace{1em}tweedie & €175,950 & 0.07 & €132,838\\
 \bottomrule
@@
 
-By analysing their importance (Table \ref{tab:osm-features-table}), the most relevant geographical features to predict housing prices are the distance to an embassy (15.6\%) and the distance to natural grasslands such as parks and gardens (4.8\%).
+By analyzing their importance (Table \ref{tab:osm-features-table}), the most relevant geographical features to predict housing prices are the distance to an embassy (15.6\%) and the distance to natural grasslands such as parks and gardens (4.8\%).
 
@@
 \multicolumn{4}{l}{\textbf{Linear regression based}}\\
-\hspace{1em}squarederror & €146,985 & 0.25 & €108,141\\
-\hspace{1em}squaredlogerror & €406,180 & 0.01 & €368,972\\
-\hspace{1em}gamma & €171,736 & 0.11 & €128,489\\
-\hspace{1em}tweedie & €174,236 & 0.10 & €130,269\\
+\hspace{1em}squarederror & €146,997 & 0.25 & €108,172\\
+\hspace{1em}squaredlogerror & €406,126 & 0.01 & €368,916\\
+\hspace{1em}gamma & €185,183 & 0.08 & €138,255\\
+\hspace{1em}tweedie & €173,179 & 0.10 & €129,087\\
 \bottomrule
@@
 The main reason is the structure of the market, mainly made of residential properties and few apartment buildings.
-In this study, the prices of 10387 properties sold in 2018 in the area of Dublin were analysed from the Property Services Regulatory Authority dataset in order to identify the potential factors correlating with these prices.
+In this study, the prices of 10387 properties sold in 2018 in the area of Dublin from the Property Services Regulatory Authority dataset were analyzed in order to identify the potential factors correlating with these prices.
 The enforcement of a housing price registry has been shown as being an essential tool to regulate the housing market (Kitchin 2013).
@@
 
-The authors would like to thank the developers of the following R packages used to process, analyse, display and report data: R (Version 4.0.2; R Core Team 2020) and the R-packages \emph{dplyr} (Version 1.0.1; Wickham, François, et al.
+The authors would like to thank the developers of the following R packages used to process, analyze, display and report data: R (Version 4.0.2; R Core Team 2020) and the R-packages \emph{dplyr} (Version 1.0.1; Wickham, François, et al.
 2020), \emph{forcats} (Version 0.5.0; Wickham 2020), \emph{ggplot2} (Version 3.3.2; Wickham 2016), \emph{here} (Version 0.1; Müller 2017), \emph{kableExtra} (Version 1.1.0; Zhu 2019), \emph{latex2exp} (Version 0.4.0; Meschiari 2015), \emph{lubridate} (Version 1.7.9; Grolemund and Wickham 2011), \emph{lwgeom} (Version 0.2.5; Pebesma 2020), \emph{magrittr} (Version 1.5; Bache and Wickham 2014), \emph{mgcv} (Version 1.8.31; Wood 2011, 2004, 2003; Wood et al.

### 2020-08-19 review/revision_letter.Rmd
final submission

@@
   cache.extra = knitr::rand_seed,
-  cache = TRUE, 
   message = FALSE, 
@@
 
-> As no specific weight or bias in the models has been hypothesized, the parameters of the XGBoost function have been set to the default value.
+> As no specific weight or bias in the models have been hypothesized, parameters of the XGBoost algorithms have been kept as default values.
 
@@
 
-> Then, the most accurate model is used to identify the features that are contributing the most to the prediction of property prices.
-This importance is a measure of the improvement in accuracy brought by a feature based on its coefficient's absolute magnitude.
+> Next, the most accurate model is used to identify the features that are contributing most to the prediction of property prices.
+This contribution (also called "importance") is a measure of the improvement in accuracy brought by a feature.
+It is based on the absolute magnitude of feature's coefficient.
 
@@
 > In addition, while some features appear to be more important than others in their contribution to property prices, these results are not causal.
-Thus, features' importance can be either the cause or the consequence of property prices in a neighborhood.
+Thus, a feature's importance can be either the cause or the consequence of property prices in a neighborhood.
 
@@
 
-> These results allow an understanding of why some areas have higher prices than others which is a relevant information not only for real estate agents in charge of property valuations but also to buyers in order to approximate the real value of a property.
+> These results allow an understanding of why some areas have higher prices than others which is relevant information not only for real estate agents in charge of property valuations but also for buyers in order to estimate the real value of a property.
 
@@
 > Indeed, artificial neural network techniques can quickly adjust to volatility changes, therefore incorporating sharp increases or decreases in their predictions [@aragones2007neural].
-However, the model obtained from most artificial neural networks are hardly interpretable [@rudin2019stop].
+However, the model obtained from most artificial neural networks are difficult to interpret [@rudin2019stop].
 
@@
 
-> This problem is known as "finite area smoothing" and occurs when predictions from a model are approximated across geographical barriers, such as irregular coastal shapes, which would lead to poorly predicted values [@ramsay2002spline].
+> This problem is known as "finite area smoothing" and occurs when predictions from a model are approximated across geographical barriers, such as irregular coastal shapes, which can lead to poorly predicted values [@ramsay2002spline].
 
@@
 ```{r}
-geocode_dublin_2018 <- here::here("data/ppr_dublin_2018.csv") %>% 
-  readr::read_csv()
+n_raw <- here::here("data/ppr_dublin_2018.csv") %>% 
+  readr::read_csv() %>% 
+  nrow()
 
-house_dublin_2018 <- geocode_dublin_2018 %>% 
-  janitor::clean_names() %>% 
-  dplyr::filter(!grepl('apartment|Apartment|APART|APT|Apt', address))
+n_house <- here::here("data/ppr_dublin_2018.csv") %>% 
+  readr::read_csv() %>% 
+  dplyr::filter(!grepl('apartment|Apartment|APART|APT|Apt', Address)) %>% 
+  nrow()
 
-final_dublin_2018 <- here::here("data/data_dublin_2018_osm_features.rds") %>% 
-  readr::read_rds()
+n_final <- here::here("data/data_dublin_2018_features.rds") %>% 
+  readr::read_rds() %>% 
+  nrow()
 ```
 
-The total amount of property sold in 2018 in the county Dublin which appears in the Property Price Register is `r nrow(geocode_dublin_2018)`.
-From these, all property corresponding to apartments were removed (`r scales::percent(1-(nrow(house_dublin_2018)/nrow(geocode_dublin_2018)), accuracy = 0.1)` of the original dataset).
-Finally, unsuccessfully geocoded properties were also removed from the original dataset (`r scales::percent(1-(nrow(final_dublin_2018)/nrow(house_dublin_2018)), accuracy = 0.1)`) to obtain the `r nrow(final_dublin_2018)` of the final dataset used to identify the urban and socio-economic correlates of property prices.
+In 2018, the Property Price Register references `r n_raw` properties sold in Dublin's area.
+Among these properties, those corresponding to apartments have been removed from the original dataset in order to evaluate only houses (`r scales::percent(1-(n_house/n_raw), accuracy = 0.1)` of the properties).
+Finally, some properties have been unsuccessfully geocoded because of mistakes on their address or because they have not been found (`r scales::percent(1-(n_final/n_house), accuracy = 0.1)` of the properties).
+These properties have also been removed from the original dataset to obtain a final dataset of `r n_final` properties.
 
@@
 
-> The results of Irish 2016 census consultation is accessible through both the Central Statistics Office [@cso2020] and All-Island Research Observatory [@airo2020].
+> The results of Irish 2016 census consultation is accessible through both the Central Statistics Office [@cso2020] and the All-Island Research Observatory [@airo2020] websites.
 

### 2020-08-19 script/census_dublin_features.R

@@ -1,68 +1,69 @@
-census_dublin_features <- c(
-carers <- c(              
-  "% Provides No Care",                          
-  "% 1-19 hours unpaid PW",                     
-  "% 20-49 hours unpaid PW",                     
-  "% 50+ hours unpaid PW",                      
-  "% Total Care Providers"                      
-),
-disabilty_age_group <- c(
-  "% Persons with a disability aged 0 - 14",      
-  "% Persons with a disability aged 15 - 44",    
-  "% Persons with a disability aged 25 - 44",   
-  "% Persons with a disability aged 45 - 64",    
-  "% Persons with a disability aged 65 Plus",   
-  "% Total persons with a disability" 
-),
-general_health <- c(
-  "% Very Good",                                
-  "% Good",                                     
-  "% Fair",                                     
-  "% Bad",                                       
-  "% Very Bad"
-),
-housing_occupancy <- c(
-  "% Occupied/HS With ususal Residents",         
-  "% Unoccupied/HS Without Ususal Residents" 
-),
-housing_rooms <- c(
-  "% 1 Room (Households)",                      
-  "% 2 Rooms (Households)",                      
-  "% 3 Rooms (Households)",                     
-  "% 4 Rooms (Households)",                      
-  "% 5 Rooms (Households)",                     
-  "% 6 Rooms (Households)",                      
-  "% 7 Rooms (Households)",                     
-  "% 8 or more Rooms (Households)"
-),
-housing_tenure <- c(
-  "% Owner Occupier with Mortgage (Households)",
-  "% Owner Occupier No Mortgage (Households)",   
-  "% Private Rented",                           
-  "% Social Rented",                             
-  "% Rented Free of Rent (Households)" 
-),                          
-housing_type <- c(
-  "% House/Bungalow",                           
-  "% Flat/Apartment/BedSit",                     
-  "% Caravan/Mobile home/Temperory" 
-),          
-population <- c(
-  "% Male",                                     
-  "% Female",                                  
-  "% Age 0 - 14",                                   
-  "% Age 15 - 24",                                 
-  "% Age 25 - 44",                                  
-  "% Age 45 - 64",                                  
-  "% Age 65 Plus",                                    
-  "% Age 15- 64",                                  
-  "% Age 80 Plus"
-),
-religion <- c(
-  "% Roman Catholic",                                 
-  "% Other Catholic",                                   
-  "% Other Religion",                                  
[…]

### 2020-08-19 script/functions_project.R

@@ -158,12 +158,6 @@ gam_density <- function (x, view = NULL, cond = list(), n.grid = 40, too.far = 0
     )
 }
 ################################################################################
-read_join <- function(file_name){
-  file.path(data_path, "aero_data", file_name) %>% 
-    readr::read_csv(col_types = list(GEOG_ID = col_character())) %>% 
-    dplyr::filter(GEOG_ID %in% list_geogid)
-}
-################################################################################
 xgboost_all_parameters <- function(booster_parameter, objective_parameter){
   if (booster_parameter == "gbtree") {
     xgboost::xgboost(

### 2020-08-20 manuscript/manuscript.Rmd
final submission

@@
 
-keywords: ["Property price", "Housing Market", "Feature Analysis", "Machine Learning", "Geocoding"]
+keywords: ["Property Price", "Housing Market", "Feature Analysis", "Machine Learning", "Geocoding"]
 abstract: |

### 2020-08-20 manuscript/manuscript.tex
final submission

@@
             pdftitle={Urban and socio-economic correlates of property prices in Dublin's area},
-            pdfkeywords={Property price, Housing Market, Feature Analysis, Machine Learning, Geocoding},
+            pdfkeywords={Property Price, Housing Market, Feature Analysis, Machine Learning, Geocoding},
             pdfborder={0 0 0},
@@
 \begin{IEEEkeywords}
-Property price; Housing Market; Feature Analysis; Machine Learning; Geocoding
+Property Price; Housing Market; Feature Analysis; Machine Learning; Geocoding
 \end{IEEEkeywords}
@@
 \multicolumn{4}{l}{\textbf{Linear regression based}}\\
-\hspace{1em}squarederror & €150,925 & 0.21 & €111,810\\
-\hspace{1em}squaredlogerror & €406,947 & 0.00 & €369,709\\
-\hspace{1em}gamma & €183,987 & 0.05 & €138,978\\
-\hspace{1em}tweedie & €175,950 & 0.07 & €132,838\\
+\hspace{1em}squarederror & €150,847 & 0.21 & €111,786\\
+\hspace{1em}squaredlogerror & €406,939 & 0.00 & €369,708\\
+\hspace{1em}gamma & €177,238 & 0.07 & €133,973\\
+\hspace{1em}tweedie & €171,480 & 0.08 & €129,138\\
 \bottomrule
@@
 \multicolumn{4}{l}{\textbf{Linear regression based}}\\
-\hspace{1em}squarederror & €146,997 & 0.25 & €108,172\\
-\hspace{1em}squaredlogerror & €406,126 & 0.01 & €368,916\\
-\hspace{1em}gamma & €185,183 & 0.08 & €138,255\\
-\hspace{1em}tweedie & €173,179 & 0.10 & €129,087\\
+\hspace{1em}squarederror & €146,963 & 0.25 & €108,151\\
+\hspace{1em}squaredlogerror & €406,213 & 0.02 & €369,023\\
+\hspace{1em}gamma & €183,008 & 0.08 & €136,445\\
+\hspace{1em}tweedie & €174,731 & 0.10 & €130,483\\
 \bottomrule
@@
 
-The authors would like to thank the developers of the following R packages used to process, analyze, display and report data: R (Version 4.0.2; R Core Team 2020) and the R-packages \emph{dplyr} (Version 1.0.1; Wickham, François, et al.
+The authors would like to thank the developers of the following R packages used to process, analyze, display and report data: R (Version 4.0.2; R Core Team 2020) and the R-packages \emph{dplyr} (Version 1.0.2; Wickham, François, et al.
 2020), \emph{forcats} (Version 0.5.0; Wickham 2020), \emph{ggplot2} (Version 3.3.2; Wickham 2016), \emph{here} (Version 0.1; Müller 2017), \emph{kableExtra} (Version 1.1.0; Zhu 2019), \emph{latex2exp} (Version 0.4.0; Meschiari 2015), \emph{lubridate} (Version 1.7.9; Grolemund and Wickham 2011), \emph{lwgeom} (Version 0.2.5; Pebesma 2020), \emph{magrittr} (Version 1.5; Bache and Wickham 2014), \emph{mgcv} (Version 1.8.31; Wood 2011, 2004, 2003; Wood et al.
@@
 2020), \emph{OpenStreetMap} (Version 0.3.4; Fellows and JMapViewer library by Jan Peter Stotz 2019), \emph{osmdata} (Version 0.1.3; Padgham et al.
-2017), \emph{papaja} (Version 0.1.0.9997; Aust and Barth 2020), \emph{patchwork} (Version 1.0.1; Pedersen 2019), \emph{purrr} (Version 0.3.4; Henry and Wickham 2020), \emph{readr} (Version 1.3.1; Wickham, Hester, and Francois 2018), \emph{sf} (Version 0.9.5; Pebesma 2018), \emph{stringr} (Version 1.4.0; Wickham 2019), \emph{tibble} (Version 3.0.3; Müller and Wickham 2020), \emph{tidyr} (Version 1.1.0; Wickham and Henry 2020), \emph{tidyverse} (Version 1.3.0; Wickham, Averick, et al.
+2017), \emph{papaja} (Version 0.1.0.9997; Aust and Barth 2020), \emph{patchwork} (Version 1.0.1; Pedersen 2019), \emph{purrr} (Version 0.3.4; Henry and Wickham 2020), \emph{readr} (Version 1.3.1; Wickham, Hester, and Francois 2018), \emph{sf} (Version 0.9.5; Pebesma 2018), \emph{stringr} (Version 1.4.0; Wickham 2019), \emph{tibble} (Version 3.0.3; Müller and Wickham 2020), \emph{tidyr} (Version 1.1.1; Wickham and Henry 2020), \emph{tidyverse} (Version 1.3.0; Wickham, Averick, et al.
 2019), \emph{tmap} (Version 3.1; Tennekes 2018), \emph{xgboost} (Version 1.1.1.1; Chen et al.

### 2020-08-20 temp.R

@@ -1,59 +1,0 @@
-function (object, show = c("estimate", "se"), contour = TRUE, 
-          contour_col = "black", n_contour = NULL, xlab, ylab, 
-          title = NULL, subtitle = NULL, caption = NULL, response_range = NULL, 
-          ...) 
-{
-  smooth_vars <- names(object)[3:4]
-  show <- match.arg(show)
-  if (isTRUE(identical(show, "estimate"))) {
-    guide_title <- "Effect"
-    plot_var <- "est"
-    guide_limits <- if (is.null(response_range)) {
-      c(-1, 1) * max(abs(object[[plot_var]]))
-    }
-    else {
-      response_range
-    }
-  }
-  else {
-    guide_title <- "Std. err."
-    plot_var <- "se"
-    guide_limits <- range(object[["se"]])
-  }
-  plt <- ggplot(object, aes_string(x = smooth_vars[1], y = smooth_vars[2])) + 
-    geom_raster(mapping = aes_string(fill = plot_var))
-  if (isTRUE(contour)) {
-    plt <- plt + geom_contour(mapping = aes_string(z = plot_var), 
-                              colour = contour_col, bins = n_contour)
-  }
-  if (missing(xlab)) {
-    xlab <- smooth_vars[1L]
-  }
-  if (missing(xlab)) {
-    xlab <- smooth_vars[1L]
-  }
-  if (missing(ylab)) {
-    ylab <- smooth_vars[2L]
-  }
-  if (is.null(title)) {
-    title <- unique(object[["smooth"]])
-  }
-  if (all(!is.na(object[["by_variable"]]))) {
-    spl <- strsplit(title, split = ":")
-    title <- spl[[1L]][[1L]]
-    if (is.null(subtitle)) {
-      by_var <- as.character(unique(object[["by_variable"]]))
-      subtitle <- paste0("By: ", by_var, "; ", 
-                         unique(object[[by_var]]))
-    }
-  }
-  plt <- plt + labs(x = xlab, y = ylab, title = title, subtitle = subtitle, 
-                    caption = caption)
-  plt <- plt + scale_fill_distiller(palette = "RdBu", 
-                                    type = "div", limits = guide_limits)
-  plt <- plt + guides(fill = guide_colourbar(title = guide_title, 
-                                             direction = "vertical", barheight = grid::unit(0.25, 
-                                                                                            "npc")))
-  plt <- plt + theme(legend.position = "right")
-  plt
-}

### 2020-09-04 presentation/slides.Rmd
desactivate renv

---
title: "Urban and socio-economic correlates of property prices in Dublin's area"
subtitle: |
"The 7th IEEE International Conference on Data Science and Advanced Analytics"
"Special Session: Environmental and Geo-spatial Data Analytics (EnGeoData)"
author: "Damien Dupré"
date: "Dublin City University - October 6th, 2020"
output:
  xaringan::moon_reader:
    css: ["default", "metropolis", "metropolis-fonts", "css/custom_design.css"]
    lib_dir: libs
    nature:
      beforeInit: "libs/cols_macro.js"
      highlightStyle: zenburn
      highlightLines: true
      countIncrementalSlides: false
params:
  eval_rights: FALSE
---

```{r setup, include = FALSE}
# general options --------------------------------------------------------------
options(scipen = 999)
set.seed(123)
# chunk options ----------------------------------------------------------------
knitr::opts_chunk$set(
  cache.extra = knitr::rand_seed, 
  message = FALSE, 
  warning = FALSE, 
  error = FALSE, 
  echo = FALSE,
  cache = FALSE,
  comment = "", 
  fig.align = "center", 
  fig.retina = 3
  )
# libraries --------------------------------------------------------------------
library(tidyverse)
library(fontawesome)
```

layout: true

<div class="custom-footer"><span>DSAA EnGeoData - Dupré (2020) </span></div>

---

# My Journey into Data Science

#### Developement of the DynEmo Facial Expression Database (Master)

* Dynamic and spontaneous emotions
* Assessed with self-reports and by observers

#### Analysis of Emotional User Experience of Innovative Tech. (Industrial PhD)

* Understand users' acceptance of technologies from their emotional response
* Based on multivariate self-reports

#### Evaluation of Emotions from Facial and Physiological Measures (Industrial PostDoc)

* Applications to marketing, sports and automotive industries
* Dynamic changes with trend extraction techniques (2 patents)

#### Performance Prediction using Machine Learning (Academic PostDoc)

* Application to sport analytics
* Big Data treatment (> 1 million users with activities recorded in the past 5 years)

---
class: inverse, mline, center, middle

# 1. A Closer Look at Property Prices

---

# Market Characteristics

An housing market is endemic to a country in a specific context.

Countries are very different in term of:
- Styles
- Materials
- Legislations
- Urban features
- Population characteristics
#https://www.visualcapitalist.com/mapped-the-countries-with-the-highest-housing-bubble-risks/

![Plot title. ](slides_insertimage_1.png)

---
class: inverse, mline, left, middle

<img class="circle" src="https://github.com/damien-dupre.png" width="250px"/>

# Thanks for your attention, find me at...

[`r fa(name = "twitter")` @damien_dupre](http://twitter.com/damien_dupre)  
[`r fa(name = "github")` @damien-dupre](http://github.com/damien-dupre)  
[`r fa(name = "link")` damien-datasci-blog.netlify.app](https://damien-datasci-blog.netlify.app)  
[`r fa(name = "paper-plane")` damien.dupre@dcu.ie](mailto:damien.dupre@dcu.ie)
